### 2019-10-31 analysis/reports/statevars.Rmd
try this

---
title: "Max N and S"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(drake)
library(ggplot2)

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache-dat.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)

loadd(all_sv, cache = cache)

```
```{r s and n plot}

statevars_plot <- ggplot(data = all_sv, aes(x = s0, y = n0, color = singletons)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_d(end = .5) +
  geom_vline(xintercept = 62) +
  geom_hline(yintercept = 15000)

statevars_plot

```

```{r s and n plot facetted}

statevars_plot_facetted <- statevars_plot +
  facet_wrap(vars(dat), scales = "free")

statevars_plot_facetted
```