### 2019-11-14 analysis/reports/cross_dataset.Rmd
run more fia

---
title: "Between-dataset comparisons"
output: 
  github_document
params:
  on_hpg: [FALSE]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

if(params$on_hpg) {

all_di <- list() 
datasets <- c("bbs", "fia_short", "gentry", "mcdb", "misc_abund_short", "portal_plants")

#datasets <- c("fia", "gentry", "mcdb", "misc_abund_short", "portal_plants")

for(i in 1:length(datasets)) {

  this_dataset = datasets[i]
  if(this_dataset == "misc_abund_short") {
    cache_loc = "miscabund"
  } else if(this_dataset == "fia_short") {

    cache_loc = "fia"
  } else {
    cache_loc = this_dataset
  }

  ## Set up the cache and config
  db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", paste0("drake-cache-", cache_loc, ".sqlite")))
  cache <- storr::storr_dbi("datatable", "keystable", db)

    all_di[[i]] <- readd(all_di_obs, cache = cache)

  all_di[[i]] <- filter(all_di[[i]], source == "observed")

  this_dat <- readd(paste0("dat_s_dat_", this_dataset), cache = cache, character_only = T)
  sv <- get_statevars(this_dat)
  all_di[[i]] <- left_join(all_di[[i]], sv, by = c("sim", "site", "source", "singletons", "dat"))

  DBI::dbDisconnect(db)
  rm(cache)
  rm(db)
}

all_di <- bind_rows(all_di)

write.csv(all_di, "all_di.csv", row.names = F)
} else {
  all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)
}
```

```{r overall percentile heatmap}

overall_heatmap_skew <- ggplot(data = all_di, aes(x = s0, y = n0, color = skew_percentile)) +
  geom_point(alpha = .2) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8, direction = -1) +
  facet_wrap(vars(singletons)) +
  ggtitle("Skewness Percentile") +
  theme(legend.position = "bottom")

overall_heatmap_simpson <- ggplot(data = all_di, aes(x = s0, y = n0, color = simpson_percentile)) +
  geom_point(alpha = .2) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8)+
  facet_wrap(vars(singletons)) +
  ggtitle("Simpson's Percentile")+
  theme(legend.position = "bottom")

overall_heatmap_skew_zoomed <- overall_heatmap_skew +
  ylim(0, 5000) +
  xlim(0, 250)+
  theme(legend.position = "none")

overall_heatmap_simpson_zoomed <- overall_heatmap_simpson +
  ylim(0, 5000) +
  xlim(0, 250)+
  theme(legend.position = "none")

gridExtra::grid.arrange(grobs = list(overall_heatmap_skew, overall_heatmap_skew_zoomed), nrow= 1)
gridExtra::grid.arrange(grobs = list(overall_heatmap_simpson, overall_heatmap_simpson_zoomed), nrow= 1)

```

Broken out by dataset

```{r dataset heatmaps, fig.width = 30, fig.height = 20}

dat_heatmap_skew <-  ggplot(data = filter(all_di, !singletons), aes(x = s0, y = n0, color = skew_percentile)) +
  geom_point(alpha = .4) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8, direction = -1)+
  facet_wrap(vars(dat), scales = "free") +
  ggtitle("Skew Percentile")+
  theme(legend.position = "bottom")
dat_heatmap_skew

dat_heatmap_simpson <-  ggplot(data = filter(all_di, !singletons), aes(x = s0, y = n0, color = simpson_percentile)) +
  geom_point(alpha = .4) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8)+
  facet_wrap(vars(dat), scales = "free") +
  ggtitle("Simpson's Percentile")+
  theme(legend.position = "bottom")
dat_heatmap_simpson
```

Against S0

```{r s0 percentile}

overall_s0_skew <- ggplot(data = filter(all_di, !singletons), aes(x = s0, y = skew_percentile)) +
  geom_point(alpha = .1) +
  geom_smooth(method = "lm", se = F, color = "pink") +
  theme_bw() +
  ggtitle("Skew percentile")
overall_s0_skew

dat_s0_skew <- overall_s0_skew +
  facet_wrap(vars(dat), scales = "free")
dat_s0_skew

overall_s0_simpson <- ggplot(data = filter(all_di, !singletons), aes(x = s0, y = simpson_percentile)) +
  geom_point(alpha = .1) +
  geom_smooth(method = "lm", se = F, color = "pink") +
  theme_bw() +
  ggtitle("Simpson percentile")
overall_s0_simpson

dat_s0_simpson <- overall_s0_simpson +
  facet_wrap(vars(dat), scales = "free")
dat_s0_simpson

```

Against nsamples

```{r nsamples}

skew_nsamples <- ggplot(data = filter(all_di, !singletons), aes(x = nsamples, y = skew_percentile)) +
  geom_point(alpha = .1) +
    geom_smooth(method = "lm", se = F, color = "pink") +

  theme_bw() +
  ggtitle("Skew percentile v. nsamples")

skew_nsamples

dat_skew_nsamples <- skew_nsamples +
  facet_wrap(vars(dat))

dat_skew_nsamples

simp_nsamples <- ggplot(data = filter(all_di, !singletons), aes(x = nsamples, y = simpson_percentile)) +
  geom_point(alpha = .1) + 
    geom_smooth(method = "lm", se = F, color = "pink") +

  theme_bw() +
  ggtitle("Simpson percentile v. nsamples")

simp_nsamples
```

```{r histograms, fig.height = 6, fig.width =6}

skewhists <- ggplot(data = all_di, aes(x = skew_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500)) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Skewness")
skewhists

simpsonhists <- ggplot(data = all_di, aes(x = simpson_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500)) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Simpson")
simpsonhists

```