### 2019-11-16 analysis/reports/cross_dataset.Rmd
fix some formatting

---
title: "Between-dataset comparisons"
output: 
  github_document
params:
  on_hpg: [FALSE]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

if(params$on_hpg) {

  all_di <- list() 
  datasets <- c("bbs", "fia_short", "gentry", "mcdb", "misc_abund_short", "portal_plants")

  #datasets <- c("fia", "gentry", "mcdb", "misc_abund_short", "portal_plants")

  for(i in 1:length(datasets)) {

    this_dataset = datasets[i]
    if(this_dataset == "misc_abund_short") {
      cache_loc = "miscabund"
    } else if(this_dataset == "fia_short") {

      cache_loc = "fia"
    } else {
      cache_loc = this_dataset
    }

    ## Set up the cache and config
    db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", paste0("drake-cache-", cache_loc, ".sqlite")))
    cache <- storr::storr_dbi("datatable", "keystable", db)

    all_di[[i]] <- readd(all_di_obs, cache = cache)

    all_di[[i]] <- filter(all_di[[i]], source == "observed")

    this_dat <- readd(paste0("dat_s_dat_", this_dataset), cache = cache, character_only = T)
    sv <- get_statevars(this_dat)
    all_di[[i]] <- left_join(all_di[[i]], sv, by = c("sim", "site", "source", "singletons", "dat"))

    DBI::dbDisconnect(db)
    rm(cache)
    rm(db)
  }

  all_di <- bind_rows(all_di)

  write.csv(all_di, "all_di.csv", row.names = F)
} else {
  all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)
}
```

```{r histograms, fig.height = 6, fig.width =6}

skewhists <- ggplot(data = all_di, aes(x = skew_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500)) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Skewness") +
  geom_vline(xintercept = 95, color = "red")
skewhists

simpsonhists <- ggplot(data = all_di, aes(x = simpson_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500)) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Simpson") +
    geom_vline(xintercept = 5, color = "red")

simpsonhists

```

Observations from these histograms:

* FIA really struggles to get 2500 samples
* Portal is the *most* concentrated of all the datasets.
* Then MCDB and misc, then gentry, then bbs. FIA is hard to place but looks idiosyncratic.
* Curious about how variation in %ile maps on to S and N, *particularly* for Gentry and BBS. 
* Skewness is less extreme than Simpson (more density in the in-between). 
* Singletons doesn't appear to be terrifically impactful. From now on filtering to *unaltered* vectors.
* I think it's best to filter to at least 2000 samples.

Some summary statistics:

```{r summary stats}

di_summary <- all_di %>%
  filter(nsamples >= 2000) %>%
  group_by(dat, singletons) %>%
  summarize(mean_skew = mean(skew_percentile, na.rm=T),
            sd_skew = sd(skew_percentile, na.rm = T),
            mean_simpson = mean(simpson_percentile, na.rm = T),
            sd_simpson = sd(simpson_percentile, na.rm = T),
            mean_nsamples = mean(nsamples),
            nsites = length(unique(site)),
            prop_skew_high = mean(skew_percentile >= 95),
            prop_simpson_low = mean(simpson_percentile <= 5)) %>%
  ungroup()

skew_summary_plot <- ggplot(data = di_summary, aes(x = singletons, y = mean_skew, group = singletons)) +
  geom_point() +
  geom_errorbar(aes(x = singletons, ymin = mean_skew - sd_skew, ymax = mean_skew + sd_skew)) +
  geom_label(aes(x = singletons, y = 10, label = signif(mean_skew, digits = 4))) +
  theme_bw() +
  facet_wrap(vars(dat)) +
  geom_hline(yintercept = 100, linetype = 2, color = "red") +
  ylim(0, 110) +
  ggtitle("Skewness, >= 2000 samples")
skew_summary_plot

simpson_summary_plot <- ggplot(data = di_summary, aes(x = singletons, y = mean_simpson, group = singletons)) +
  geom_point() +
  geom_errorbar(aes(x = singletons, ymin = mean_simpson - sd_simpson, ymax = mean_simpson + sd_simpson)) +
    geom_label(aes(x = singletons, y = 90, label = signif(mean_simpson, digits = 4))) +
  theme_bw() +
  facet_wrap(vars(dat)) +
  geom_hline(yintercept = 0, linetype = 2, color = "red") +
  ylim(-10, 100) +
  ggtitle("Simpson's, >= 2000 samples")
simpson_summary_plot

print(select(di_summary, dat, singletons, mean_nsamples, nsites, prop_skew_high, prop_simpson_low))

```

```{r s and n mapping, fig.width = 12, fig.height = 18}

all_di <- filter(all_di, !singletons)

skew_s_n_maps <- ggplot(data = all_di, aes(x = s0, y = n0, color = skew_percentile)) +
  geom_point(alpha = .1, size = .5) +
  geom_point(data = filter(all_di, nsamples >= 2000), alpha = .2) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free", ncol = 2) + 
  ggtitle("Skewness") +
  scale_color_viridis_c(option = "plasma", end= .75)

skew_s_n_maps

simpson_s_n_maps <- ggplot(data = all_di, aes(x = s0, y = n0, color = simpson_percentile)) +
  geom_point(alpha =  .1, size = .5) +
  geom_point(data = filter(all_di, nsamples >= 2000), alpha = .2) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free", ncol = 2)  + 
  ggtitle("Simpson") +
  scale_color_viridis_c(option = "plasma", end= .75, direction = -1)

simpson_s_n_maps

```

The less extreme (low skew, high simpson) vectors appear vaguely collected in the lower right for Gentry and BBS. Those are the regions with relatively high N/S, aka low average abundance, aka a relatively small feasible set.

Conversely, for MCDB and Misc, it almost looks like the very least skewed elements are way down at tiny abundances?

Do percentiles scan with S/N?

```{r avg abund}

all_di <- mutate(all_di, avg_abund = n0/s0)

skew_avgabund <- ggplot(data = all_di, aes(x = avg_abund, y = skew_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  ggtitle("Skew") +
  facet_wrap(vars(dat), scales = "free")

skew_avgabund

simpson_avgabund <- ggplot(data = all_di, aes(x = avg_abund, y = simpson_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  ggtitle("Simpson") +
  facet_wrap(vars(dat), scales = "free")

simpson_avgabund

```
There is perhaps a constraint, but that might be because you rarely get v high average abundances. You seem to rarely get low skewness or high Simpson at high average abundance, which scans with the heatmaps.

Do Simpson and skewness correspond?
```{r do skew and simp correlate}

ss_plot <- ggplot(data = all_di, aes(x= skew_percentile, y = simpson_percentile, color = dat)) +
  geom_point(alpha = .1) +
  scale_color_viridis_d(end = .8) +
  theme_bw() +
  ggtitle("Skew v Simp")

ss_plot

ss_facet_plot <- ss_plot +
  facet_wrap(vars(dat))

ss_facet_plot
```

There's a weak relationship between Simpson and skewness, but a lot of noise. They are not substitutable.

Why is Gentry so strangely bimodal? It's like a U, when all the others tend towards one end or the other.

Why are BBS and mammals less frequently squished than Portal?

Looking at range of skewness and Simpsons:

```{r ranges}

skew_range_hist <- ggplot(data = filter(all_di, nsamples > 2000), aes(x = skew_range, y = skew_percentile, color = dat)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "lm", se =F) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  facet_wrap(vars(dat), scales = "free") + 
  ggtitle("Skewness")

skew_range_hist

simpson_range_hist <- ggplot(data = filter(all_di, nsamples > 2000), aes(x = simpson_range, y = simpson_percentile, color = dat)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "lm", se =F) +

  theme_bw() +
  scale_color_viridis_d(end = .8) +
  facet_wrap(vars(dat), scales = "free") +
  ggtitle("Simpson")

simpson_range_hist

```

I added the ranges because I thought the percentiles might be constrained somewhat by the range of values represented in the feasible set. It doesn't look to me like there is a strong relationship between range and %ile.