### 2020-01-17 analysis/reports/regressions.Rmd
adding some glms, not sure that this is the right statistical appraoch

---
title: "Regressions"
output: 
  github_document
params:
  on_hpg: [FALSE]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

if(params$on_hpg) {

  all_di <- list() 
  datasets <- c("bbs", "fia_short", "gentry", "mcdb", "misc_abund_short", "portal_plants")

  #datasets <- c("fia", "gentry", "mcdb", "misc_abund_short", "portal_plants")

  for(i in 1:length(datasets)) {

    this_dataset = datasets[i]
    if(this_dataset == "misc_abund_short") {
      cache_loc = "miscabund"
    } else if(this_dataset == "fia_short") {

      cache_loc = "fia"
    } else {
      cache_loc = this_dataset
    }

    ## Set up the cache and config
    db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", paste0("drake-cache-", cache_loc, ".sqlite")))
    cache <- storr::storr_dbi("datatable", "keystable", db)

    all_di[[i]] <- readd(all_di_obs, cache = cache)

    all_di[[i]] <- filter(all_di[[i]], source == "observed")

    this_dat <- readd(paste0("dat_s_dat_", this_dataset), cache = cache, character_only = T)
    sv <- get_statevars(this_dat)
    all_di[[i]] <- left_join(all_di[[i]], sv, by = c("sim", "site", "source", "singletons", "dat"))

    DBI::dbDisconnect(db)
    rm(cache)
    rm(db)
  }

  all_di <- bind_rows(all_di)

  write.csv(all_di, "all_di.csv", row.names = F)
} else {
  all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)
}
```

```{r histograms, fig.height = 6, fig.width =6}

skewhists <- ggplot(data = all_di, aes(x = skew_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3, boundary = 0) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500), boundary = 0) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Skewness") +
  geom_vline(xintercept = 95, color = "red")
skewhists

simpsonhists <- ggplot(data = all_di, aes(x = simpson_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3, boundary = 0) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500), boundary = 0) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Simpson") +
  geom_vline(xintercept = 5, color = "red")

simpsonhists

```

Observations from these histograms:

* FIA really struggles to get 2500 samples
* Portal is the *most* concentrated of all the datasets.
* Then MCDB and misc, then gentry, then bbs. FIA is hard to place but looks idiosyncratic.
* Curious about how variation in %ile maps on to S and N, *particularly* for Gentry and BBS. 
* Skewness is less extreme than Simpson (more density in the in-between). 
* Singletons doesn't appear to be terrifically impactful. From now on filtering to *unaltered* vectors.
* I think it's best to filter to at least 2000 samples.

Some summary statistics:

```{r summary plots}

di_summary <- all_di %>%
  filter(nsamples >= 2000) %>%
  mutate(avg_n = n0 / s0,
         skew_pp = skew_percentile / 100,
         simp_pp = simpson_percentile/100)
```

### Avg abundance

```{r avg abundance regression}

avgn_skew <- ggplot(data = di_summary, aes(x = avg_n, y = skew_pp)) +
  geom_point() +
  stat_smooth(method = "glm",method.args = list(family = binomial)) + 
  facet_wrap(vars(dat), scales = "free") +
  theme_bw()

avgn_skew

summary(glm(di_summary, formula = skew_pp ~ avg_n * dat, family = binomial))

for(i in unique(di_summary$dat)) {
  this_dat <- filter(di_summary, dat==i)
  print(i)
  print(summary(glm(skew_pp ~ avg_n, family = binomial,
          data = this_dat)))
}

```