### 2020-02-07 analysis/reports/overlap.Rmd
rough overlap comparison

---
title: "Comparing within overlapping s0-n0 space"
output: 
  github_document
params:
  on_hpg: [FALSE]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

if(params$on_hpg) {

  all_di <- list() 
  datasets <- c("bbs", "fia_short", "gentry", "mcdb", "misc_abund_short", "portal_plants")

  #datasets <- c("fia", "gentry", "mcdb", "misc_abund_short", "portal_plants")

  for(i in 1:length(datasets)) {

    this_dataset = datasets[i]
    if(this_dataset == "misc_abund_short") {
      cache_loc = "miscabund"
    } else if(this_dataset == "fia_short") {

      cache_loc = "fia"
    } else {
      cache_loc = this_dataset
    }

    ## Set up the cache and config
    db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", paste0("drake-cache-", cache_loc, ".sqlite")))
    cache <- storr::storr_dbi("datatable", "keystable", db)

    all_di[[i]] <- readd(all_di_obs, cache = cache)

    all_di[[i]] <- filter(all_di[[i]], source == "observed")

    this_dat <- readd(paste0("dat_s_dat_", this_dataset), cache = cache, character_only = T)
    sv <- get_statevars(this_dat)
    all_di[[i]] <- left_join(all_di[[i]], sv, by = c("sim", "site", "source", "singletons", "dat"))

    DBI::dbDisconnect(db)
    rm(cache)
    rm(db)
  }

  all_di <- bind_rows(all_di)

  write.csv(all_di, "all_di.csv", row.names = F)
} else {
  all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)
}
```

## Where the datasets fall

```{r dataset space}
ggplot(data = all_di, aes(x = log(s0), y = log(n0), color = dat)) + 
  geom_point(alpha = .2) +
  scale_color_viridis_d() +
  theme_bw() +
  geom_point(data = filter(all_di, dat == "I don't exist!")) # for a legend with alpha = 1

ggplot(data = filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = dat)) + 
  geom_point(alpha = .5) +
  scale_color_viridis_d() +
  theme_bw() +
  xlim(2.75, 3.25) +
  ylim(4.5, 6) +
  geom_point(data = filter(all_di, dat == "I don't exist!")) # for a legend with alpha = 1 
ggplot(data = filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = dat)) + 
  geom_point(alpha = .5) +
  scale_color_viridis_d() +
  theme_bw() +
  xlim(2.75, 3.25) +
  ylim(4.5, 6) +
  facet_wrap(vars(dat)) +
  geom_point(data = filter(all_di, dat == "I don't exist!")) # for a legend with alpha = 1

overlap_di <- filter(all_di,
                     log(s0) > 2.75,
                     log(s0) < 3.25,
                     log(n0) > 4.5,
                     log(n0) < 6)
summary_overlap <- overlap_di %>%
  filter(singletons == F) %>%
  group_by(dat) %>%
  summarize(count = dplyr::n()) %>%
  ungroup() %>%
  mutate(prop = count / sum(count))

summary_overlap

```

The datasets overlap but do occupy broadly different regions of S*N space. The arm - which is the weirdest region of the percentile values - is, unfortunately, 100% one dataset (Gentry).

For the vast majority of S*N space here, I don't think we can really use dataset as a predictor of percentile, because of the way they occupy different parts of what we know to be a very important range of variation.

We *could* narrow in on the region of maximum overlap/minimum variation and make comparison based on the sites only within that.

I think there's a plausible comparison between bbs and misc_abund in the subset depicted above. You could find other regions of pairwise intersection between the other datasets, but I don't love going down that road (of maximum comparisons).

```{r overlap plots}

overlap_di <- overlap_di %>%
  filter(!singletons,
         dat %in% c("bbs", "misc_abund_short"))

ggplot(data = overlap_di, aes(x = log(s0), y = log(n0), color = log(n0 / s0))) +
  geom_point(alpha = .5) +
  scale_color_viridis_c(option = "plasma", end = .9) +
  facet_wrap(vars(dat)) +
  theme_bw()

ggplot(data = overlap_di, aes(x = skew_percentile)) + 
  geom_histogram(bins = 100) + 
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(data = overlap_di, aes(x = log(s0), y = log(n0), color = skew_percentile)) +
  geom_point(alpha = .5) +
  scale_color_viridis_c(option = "plasma", end = .9) +
  facet_wrap(vars(dat)) +
  theme_bw()

ggplot(data = overlap_di, aes(x = simpson_percentile)) + 
  geom_histogram(bins = 100) + 
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(data = overlap_di, aes(x = log(s0), y = log(n0), color = simpson_percentile)) +
  geom_point(alpha = .5) +
  scale_color_viridis_c(option = "plasma", end = .9, direction = -1) +
  facet_wrap(vars(dat)) +
  theme_bw()

ggplot(data = overlap_di, aes(x = log(n0/s0))) +
  geom_histogram() +
  facet_wrap(vars(dat), scales = "free_y") +
  theme_bw()

ggplot(data = overlap_di, aes(x = log(n0 / s0), y = skew_percentile, color = dat)) +
  geom_point(alpha = .5) +
  scale_color_viridis_d(end = .8) +
  theme_bw() +
  geom_smooth(method = "lm", se =F)

ggplot(data = overlap_di, aes(x = log(n0 / s0), y = simpson_percentile, color = dat)) +
  geom_point(alpha = .5) +
  scale_color_viridis_d(end = .8) +
  theme_bw() +
  geom_smooth(method = "lm", se =F)
```

This is a rough comparison all around, but I am not seeing a pronounced difference in behavior between the datasets.