### 2020-02-12 analysis/reports/nparts.Rmd
add count elements

---
title: "Nparts"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

```{r fxns}

brute_force_parts <- function(s, n, nsamples = 1000) {
  bunch_of_fs <- feasiblesads::sample_fs(s, n, nsamples)
  colnames(bunch_of_fs) <- 1:s

  bunch_of_fs <- bunch_of_fs %>%
    as.data.frame %>%
    distinct() %>%
    mutate(sim = dplyr::row_number()) %>%
    tidyr::gather(-sim, key = "species", value = "abund") %>%
    arrange(sim)
  return(max(bunch_of_fs$sim)) 
}

# count_parts <- function(s, n, p_table = NULL) {
#   
#   if(is.null(p_table)) {
#     p_table <- feasiblesads::fill_ps(s, n, F)
#   }
#   
#   k_table <- feasiblesads::fill_ks(s, n)
#   
#   
#   ns = 1:k_table[s, n+1] # this differs from `feasiblesads` because 0 is NOT an option
#   
#   n_parts = p_table[s - 1,
#                     n - (s * ns) + 1]
#   total_parts = gmp::sum.bigz(gmp::as.bigz(n_parts))
#   
#   return(as.integer(total_parts))
# }

```

```{r small range}
s = seq(2, 10, by = 2)
n = seq(3, 53, by = 5)

comms <- expand.grid("s" =s , "n" = n) %>%
  filter(n > s) %>%
  mutate(bf_n = NA,
         cp_n = NA)

for(i in 1:nrow(comms)) {
  comms$bf_n[i] <- brute_force_parts(comms$s[i], comms$n[i], nsamples = 1000)
  comms$cp_n[i] <- scadsanalysis::count_elements(comms$s[i], comms$n[i])
}

ggplot(data = comms, aes(x = bf_n, y = cp_n)) +
  geom_point() +
  geom_abline(intercept = 0, slope  = 1) +
  theme_bw()

```
So this function looks pretty good until the size of the fs approaches 2x nb samples.

```{r s and n}
ggplot(data = comms, aes(x = s, y = n, color = cp_n)) +
  geom_point() +
  theme_bw()

```