### 2020-02-12 analysis/reports/dat_report_template.Rmd
new report

---
title: "Dataset-level report"
output: 
  github_document
params:
  dataset_name: ["mcdb"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

this_dataset <- params$dataset_name

if(this_dataset == "misc_abund_short") {
  cache_loc = "miscabund"
} else if (this_dataset == "fia_short") {
  cache_loc = "fia" 
} else {
  cache_loc = this_dataset
}

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", paste0("drake-cache-", cache_loc, ".sqlite")))
cache <- storr::storr_dbi("datatable", "keystable", db)

dat <- readd(paste0("dat_s_dat_", this_dataset), cache = cache, character_only = T)

all_di <- readd(all_di_obs, cache = cache)

```
# This dataset is `r this_dataset`.

## State variables and change with singletons

Here is a plot of the distribution of S and N in this dataset, and S and N if we estimate the true number of species.

```{r statevars}

sv <- get_statevars(dat)

statevars_plot <- ggplot(data = sv, aes(x = s0, y = n0, color = singletons)) +
  geom_point(alpha = .5) +
  theme_bw() +
  facet_wrap(vars(singletons)) +
  scale_color_viridis_d(end = .5) +
  ggtitle("State variables")

statevars_plot
```

Here is a plot of how many species were added to each dataset by estimating the true number of species.

```{r sv change}
sv_singles <- filter(sv, singletons == TRUE) %>%
  select(-singletons) %>%
  rename(singletons_s0 = s0,
         singletons_n0 = n0)

sv_change <- filter(sv, singletons == FALSE) %>%
  select(-singletons) %>%
  left_join(sv_singles, by = c("site", "dat", "sim", "source")) %>%
  mutate(s0_change = singletons_s0 - s0,
         n0_change = singletons_n0 - n0) %>%
  select(-singletons_s0, -singletons_n0)

sv_change_plot <- ggplot(data = sv_change, aes(x = s0, y = n0, color = log(s0_change))) +
  geom_point(alpha = .5) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .75, direction = -1) +
  ggtitle("State variable change")

sv_change_plot
```

## Number of samples achieved

```{r number of samples}
# 
# site_samples <- all_di %>%
#   left_join(select(sv, -source, -sim), by = c("site", "singletons"))
```

#### The maximum number of samples achieved is `r max(all_di$nsamples)`.

```{r plot nb samples}
samples_hist <- ggplot(data = all_di, aes(x = nsamples)) +
  geom_histogram(binwidth = 10) +
  theme_bw() +
  facet_wrap(vars(singletons), nrow = 1) +
  ggtitle("Number of samples achieved")  +
  xlim(c(-11, max(all_di$nsamples) + 11))

samples_hist

sv_samples_plot <- ggplot(data = all_di, aes(x = s0, y = n0, color = nsamples)) +
  geom_point(alpha = .5) +
  theme_bw() +
  facet_wrap(vars(singletons), nrow = 1) +
  scale_color_viridis_c(option = "magma", end = .75, limits = c(0, max(all_di$nsamples))) +
  ggtitle("Number of samples achieved by state variables")
sv_samples_plot

```

## Position (%ile) of empirical SADs within FS

```{r empirical positions}

skew_plot <- ggplot(data = filter(all_di, source == "observed"), aes(x= singletons, y = skew_percentile)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Skewness percentile") +
  ylim(0, 100)

simpson_plot <- ggplot(data = filter(all_di, source == "observed"), aes(x= singletons, y = simpson_percentile)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Simpson percentile") +
  ylim(0, 100)

gridExtra::grid.arrange(grobs = list(skew_plot, simpson_plot), ncol = 2)
```

Here is a heatmap of an observed + sampled FS:

```{r example heatmap}

a_fs_name <- cached(cache = cache)[ which(substr(cached(cache = cache), 0, 3) == "fs_")[1]]

fs_samples <- readd(a_fs_name, cache = cache, character_only = TRUE)

fs_heatmap <- ggplot(data = fs_samples, aes(x = rank, y = abund, group = sim, color = source)) +
  geom_line(alpha = .01) +
  geom_line(data = filter(fs_samples, source == "observed")) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  ggtitle(a_fs_name)

fs_heatmap

```

```{r manip}

if(this_dataset == "macdb") {
  all_di_manip <- readd(all_di_manip, cache = cache)

  manip_skew_plot <- ggplot(data = all_di_manip, aes(x = treatment, y = skew_percentile, group = studyID)) +
    geom_line() +
    theme_bw()

  print(manip_skew_plot)

  manip_simpson_plot <- ggplot(data = all_di_manip, aes(x = treatment, y = simpson_percentile, group = studyID)) +
    geom_line() +
    theme_bw()

  print(manip_simpson_plot)

}

if(this_dataset == "portal_plants_manip") {
  all_di_manip <- readd(all_di_manip, cache = cache)

  manip_skew_plot <- ggplot(data = all_di_manip, aes(x = treatment, y = skew_percentile)) +
    geom_boxplot() +
    facet_wrap(vars(season, singletons)) +
    theme_bw()

  print(manip_skew_plot)

  manip_simpson_plot <- ggplot(data = all_di_manip, aes(x = treatment, y = simpson_percentile)) +
    geom_boxplot() +
    facet_wrap(vars(season, singletons)) +
    theme_bw()

  print(manip_simpson_plot)
}
```

```{r cleanup}

DBI::dbDisconnect(db)
rm(cache)
```