### 2020-02-18 analysis/reports/synthesis.Rmd
run this

---
title: "Synthesis report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)
```

# Datasets in S and N space

# Number of samples
- How many samples are we achieving
- What % of the FS is this?

#