### 2020-02-19 analysis/reports/synthesis.Rmd

---
title: "Synthesis report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples))
```

# Datasets in S and N space

Here is where our communities fall in S and N space:

```{r datasets in s and n space}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = dat)) +
  geom_point(alpha = .2) +
  theme_bw() +
  scale_color_viridis_d()

```

Here is how that translates into the size of the feasible set:

```{r size of fs}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = log_nparts)) +
  geom_point(alpha = .3) +
  theme_bw() +
  scale_color_viridis_c(option = "magma", end = .9)

```

Note that the color scale is log transformed, so the largest communities have e^`r max(all_di$log_nparts)`, or `r exp(max(all_di$log_nparts))`, elements in the feasible set!

Here is how the size of the feasible set maps on to N/S:

```{r nparts vs avgn}
ggplot(filter(all_di, singletons == F), aes(x = log(n0 / s0), y = log_nparts, color = log(s0))) +
  theme_bw() +
  geom_point(alpha = .2) +
  scale_color_viridis_c()

```

# Number of samples

Here is how many samples we are achieving:

```{r nsamples}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = log_nsamples)) +
  geom_point(alpha = .3) +
  theme_bw() +
  scale_color_viridis_c(option = "magma", end = .9)

ggplot(filter(all_di, singletons == F, nsamples <  max(nsamples)), aes(x = log(s0), y = log(n0), color = log_nsamples)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(option = "magma", end = .9)
```

Only in the very smallest communities do we get appreciably fewer than `r max(all_di$nsamples)` samples.

Here is how the number of samples we're getting compares to the size of the feasible set:

```{r nsamples vs nparts}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = log_nsamples - log_nparts)) +
  geom_point(alpha = .7) +
  theme_bw() +
  scale_color_viridis_c(option = "magma", end = .9, direction = -1) +
  theme(legend.position = "top")

ggplot(filter(all_di, singletons == F), aes(x = log_nsamples - log_nparts, y = log_nsamples)) +
  theme_bw() +
  geom_vline(xintercept = c(log(.95), log(.1))) +
  geom_point(alpha = .3)

```

The vertical lines are the 10% and 95% marks, from left to right. When we got relatively few samples (`r min(all_di$nsamples)`), we had found `r exp(filter(all_di, nsamples == min(nsamples))$log_nsamples -  filter(all_di, nsamples == min(nsamples))$log_nparts)` of that feasible set. And there's a clear negative relationship between the proportion of the FS that we've found and the number of samples we got, once we start getting about 10% of the feasible set. This makes sense; at that point we have a 1 in 10 chance of drawing one we've already seen. But for almost all cases, we're sampling only a miniscule proportion of the feasible set - median of `r median(exp(all_di$log_nsamples - all_di$log_nparts))`! - and get no duplicates.

The vastness of the FS poses a limit on our capacity to detect how unlikely a value is. We could conservatively round up: scoring in the 100th percentile, when we had 4000 samples, means a maximum of (1 / 4000 + tiny error) chance of getting that score?

#