### 2020-02-19 analysis/reports/cdf.Rmd
cdf exploring

---
title: "Digging into %ile and CDF fxns"
output: 
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

this_dataset <- "portal_plants"

if(this_dataset == "misc_abund_short") {
  cache_loc = "miscabund"
} else if (this_dataset == "fia_short") {
  cache_loc = "fia" 
} else {
  cache_loc = this_dataset
}

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", paste0("drake-cache-", cache_loc, ".sqlite")))
cache <- storr::storr_dbi("datatable", "keystable", db)

dat <- readd(paste0("dat_s_dat_", this_dataset), cache = cache, character_only = T)

a_fs <- readd(di_fs_1991_winter_FALSE, cache = cache)
```

Here is the distribution of skew values in this fs:

```{r raw dists}

ggplot(filter(a_fs, source == "sampled"), aes(x = skew)) + 
  geom_histogram() +
  theme_bw()

```

The percentile ranks as computed by `get_percentile` for the max and minimum sampled values are `r get_percentile(max(filter(a_fs, source == "sampled")$skew), filter(a_fs, source == "sampled")$skew)` and `r get_percentile(min(filter(a_fs, source == "sampled")$skew), filter(a_fs, source == "sampled")$skew)`.

```{r ecdf}

skew_ecdf <- ecdf(filter(a_fs, source == "sampled")$skew)

plot(skew_ecdf)

```

The ecdf function at hte minimum and maximum sampled valures evaluates to `r skew_ecdf(max(filter(a_fs, source == "sampled")$skew))` and `r skew_ecdf(min(filter(a_fs, source == "sampled")$skew))`.

```{r ecdf w infinities}

inf_ecdf <- ecdf(c(filter(a_fs, source == "sampled")$skew, -Inf, Inf))
plot(x = filter(a_fs, source == "sampled")$skew, y = inf_ecdf(filter(a_fs, source == "sampled")$skew))
```

The inf_ecdf function at the minimum and maximum sampled valures evaluates to `r inf_ecdf(max(filter(a_fs, source == "sampled")$skew))` and `r inf_ecdf(min(filter(a_fs, source == "sampled")$skew))`.

```{r cleanup}

DBI::dbDisconnect(db)
rm(cache)
```