### 2020-02-20 analysis/reports/data_cdf.Rmd
cdf plots, not sure what to make of them

---
title: "CDF v percentiles report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples)) %>%
  mutate(prop_found = exp(log_nsamples - log_nparts)) %>%
  mutate(found_all = prop_found ==1)
```

## Skewness
```{r skew cds}
ggplot(filter(all_di, singletons == F), aes(x = skew_cd, y =..count.. / sum(..count..))) +
  geom_histogram(bins = 100) +
  theme_bw() +
  geom_hline(yintercept = .01)

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = skew_cd)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "plasma", end = .9)

ggplot(filter(all_di, singletons == F), aes(x = log(s0/n0), y = skew_cd, color = prop_found)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .9, direction = -1) +
  theme(legend.position = "top")

all_di$many_found <- all_di$prop_found >= .5

ggplot(filter(all_di, singletons == F), aes(x = skew_cd)) +
  theme_bw() +
  geom_histogram(bins = 100) +
  facet_wrap(vars(many_found))

ggplot(filter(all_di, singletons == F), aes(x = log(n0 / s0), y = skew_cd, color = log(s0))) +
  geom_point(alpha = .2) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c()
```

## Simpson
```{r simpson cds}
ggplot(filter(all_di, singletons == F), aes(x = simpson_cd, y =..count.. / sum(..count..))) +
  geom_histogram(bins = 100) +
  theme_bw() +
  geom_hline(yintercept = .01)

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = simpson_cd)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "plasma", end = .9)

ggplot(filter(all_di, singletons == F), aes(x = log(s0/n0), y = simpson_cd, color = prop_found)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .9, direction = -1) +
  theme(legend.position = "top")

ggplot(filter(all_di, singletons == F), aes(x = simpson_cd)) +
  theme_bw() +
  geom_histogram(bins = 100) +
  facet_wrap(vars(many_found))

ggplot(filter(all_di, singletons == F), aes(x = log(n0 / s0), y = simpson_cd, color = log(s0))) +
  geom_point(alpha = .2) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c()
```