### 2020-03-18 analysis/reports/synthesis.Rmd

---
title: "Synthesis report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples)) %>%
  mutate(prop_found = exp(log_nsamples - log_nparts)) %>%
  mutate(found_all = prop_found==1)
```

# Datasets in S and N space

Here is where our communities fall in S and N space:

```{r datasets in s and n space}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = dat)) +
  geom_point(alpha = .8) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_d()

```

Here is how that translates into the size of the feasible set:

```{r size of fs}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = log_nparts)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

```

Note that the color scale is log transformed, so the largest communities have e^`r max(all_di$log_nparts)`, or `r exp(max(all_di$log_nparts))`, elements in the feasible set!

Here is how the size of the feasible set maps on to N/S. It increases with n0/s0 and s0.

```{r nparts vs avgn}
ggplot(filter(all_di, singletons == F), aes(x = log(n0 / s0), y = log_nparts, color = log(s0))) +
  theme_bw() +
  geom_point(alpha = .2) +
  theme(legend.position = "top") + scale_color_viridis_c()

```

# Number of samples

Here is how many samples we are achieving:

```{r nsamples}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = log_nsamples)) +
  geom_point(alpha = .3) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9)

```

Only in small communities do we get appreciably fewer than `r max(all_di$nsamples)` samples.

Here is how the number of samples we're getting compares to the size of the feasible set:

```{r nsamples vs nparts, fig.dim = c(4,3)}
ggplot(filter(all_di, singletons == F), aes(x = prop_found)) +
  theme_bw() +
  geom_histogram(bins = 100)

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = prop_found)) +
  geom_point(alpha = .7) +
  geom_point(data = filter(all_di, singletons == F, prop_found == 1), color = "grey") +
  theme_bw() +
  scale_color_viridis_c(option = "magma", end = .9, direction = -1) #+
#theme(legend.position = "top")

ggplot(filter(all_di, singletons == F), aes(x = prop_found, y =nsamples)) +
  theme_bw() +
  geom_vline(xintercept = c((.95), (.1))) +
  geom_point(alpha = .3) +
  geom_point(data = filter(all_di, singletons == F, prop_found == 1), color = "grey")

prop_found_ecdf <- ecdf(filter(all_di,singletons == F)$prop_found)

pf_df <- data.frame(prop_found = seq(0, 1, by = 0.001),
                    cdf = prop_found_ecdf(seq(0, 1, by = 0.001)))

ggplot(data = pf_df, aes(x = prop_found, y = cdf)) +
  geom_line() +
  theme_bw()

```

For about `r 100 * mean(filter(all_di, singletons == F)$prop_found == 1)`% of sites, we found all the elements of the FS. The *vast majority* of this is FIA - here is what happens if we take out FIA:

```{r prop found no fia, fig.dim = c(4,3)}
ggplot(filter(all_di, singletons == F, dat != "fia_short"), aes(x = prop_found)) +
  theme_bw() +
  geom_histogram(bins = 100)

ggplot(filter(all_di, singletons == F, dat != "fia_short"), aes(x = log(s0), y = log(n0), color = prop_found)) +
  geom_point(alpha = .7) +
  geom_point(data = filter(all_di, singletons == F, dat != "fia_short", prop_found == 1), color = "grey") +
  theme_bw() +
  scale_color_viridis_c(option = "magma", end = .9, direction = -1)# +
#theme(legend.position = "top")

ggplot(filter(all_di, singletons == F, dat != "fia_short"), aes(x = prop_found, y =nsamples)) +
  theme_bw() +
  geom_vline(xintercept = c((.95), (.1))) +
  geom_point(alpha = .3) +
  geom_point(data = filter(all_di, singletons == F, prop_found == 1, dat != "fia_short"), color = "grey")

prop_found_ecdf <- ecdf(filter(all_di,singletons == F, dat != "fia_short")$prop_found)

pf_df <- data.frame(prop_found = seq(0, 1, by = 0.001),
                    cdf = prop_found_ecdf(seq(0, 1, by = 0.001)))

ggplot(data = pf_df, aes(x = prop_found, y = cdf)) +
  geom_line() +
  theme_bw()

```

Without FIA, we find all the samples about `r 100 * mean(filter(all_di, singletons == F, dat != "fia_short")$prop_found == 1)`% of the time.

# Distribution of percentile values

## Skewness

Here is the overall distribution of skewness, and if we split based on whether we found all the samples:
```{r skew percentiles}
ggplot(filter(all_di, singletons == F), aes(x = skew_percentile, y =..count.. / sum(..count..))) +
  geom_histogram(bins = 100) +
  theme_bw() +
  geom_hline(yintercept = .01)
ggplot(filter(all_di, singletons == F), aes(x = skew_percentile)) +
  theme_bw() +
  geom_histogram(bins = 100) +
  facet_wrap(vars(found_all), scales = "free_y")
```

When we found all the samples, the percentiles are more evenly distributed. I do not read much into the spike at 0 for those communities, because skewness is bizarre for very small communities.

Here is how skewness maps with S and N:
```{r skew v s and n, fig.dim = c(10,10)}

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = skew_percentile)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "plasma", end = .9)

# ggplot(filter(all_di, singletons == F), aes(x = log(n0/s0), y = skew_percentile, color = prop_found)) +
#   geom_point() +
#   theme_bw() +
#   scale_color_viridis_c(option = "plasma", end = .9, direction = -1) +
#   theme(legend.position = "top")
# 
# ggplot(filter(all_di, singletons == F), aes(x = log(n0 / s0), y = skew_percentile, color = log(s0))) +
#   geom_point(alpha = .2) +
#   theme_bw() +
#   theme(legend.position = "top") + scale_color_viridis_c()
```

The very low skewness values are down in the very small and very weird communities. There may be variation along S and N elsewhere, but it is hard to parse.

## Simpson

Here is the overall evenness distribution, and split by whether we found 'em all:

```{r simpson percentiles}
ggplot(filter(all_di, singletons == F), aes(x = simpson_percentile, y =..count.. / sum(..count..))) +
  geom_histogram(bins = 100) +
  theme_bw() +
  geom_hline(yintercept = .01)

ggplot(filter(all_di, singletons == F), aes(x = simpson_percentile)) +
  theme_bw() +
  geom_histogram(bins = 100) +
  facet_wrap(vars(found_all))

```

Simpson is less evenly distributed than skewness. Again, where we found them all, we don't see the disproportionately common low percentile values.

Here is how Simpson behaves in S and N space:

```{r even v s and n, fig.dim = c(10,10)}
# ggplot(filter(all_di, singletons == F), aes(x = log(n0 / s0), y = simpson_percentile, color = log(s0))) +
#   geom_point(alpha = .2) +
#   theme_bw() +
#   theme(legend.position = "top") + scale_color_viridis_c()

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = simpson_percentile)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "plasma", end = .9)
# ggplot(filter(all_di, singletons == F), aes(x = log(s0/n0), y = simpson_percentile, color = prop_found)) +
#   geom_point() +
#   theme_bw() +
#   scale_color_viridis_c(option = "plasma", end = .9, direction = -1) +
#   theme(legend.position = "top")
```

There is unusual behavior where S is large and N/S is relatively small (log N/S <= 1.5), where evenness is unusually *high*.

For both skew and evenness, we do not see non-extreme percentile values in large communities:

```{r non extremes}

ggplot(filter(all_di, singletons == F), aes(x = log(n0 / s0), y = skew_percentile, color = log(n0))) +
  geom_point(alpha = .1) +
  theme_bw() +
  theme(legend.position = "top") + 
  scale_color_viridis_c(end = .9)

ggplot(filter(all_di, singletons == F), aes(x = log(n0 / s0), y = simpson_percentile, color = log(n0))) +
  geom_point(alpha = .1) +
  theme_bw() +
  theme(legend.position = "top") + 
  scale_color_viridis_c(end = .9)

```

## Singletons

Here is how singletons change percentiles, broken out by whether or not we found all the samples:

```{r singletons overall}

singletons_effect <- all_di %>%
  select(dat, nsamples, site, singletons, skew_percentile, simpson_percentile, found_all) %>%
  tidyr::pivot_wider(names_from = singletons, values_from = c(skew_percentile, simpson_percentile, nsamples, found_all)) %>%
  mutate(skewdiff = skew_percentile_TRUE - skew_percentile_FALSE,
         simpdiff= simpson_percentile_TRUE - simpson_percentile_FALSE,
         found_all = (found_all_TRUE + found_all_FALSE) == 2)

ggplot(data = singletons_effect, aes(x = skew_percentile_FALSE, y = skew_percentile_TRUE)) +
  geom_point(alpha = .1) +
  theme_bw() +
  geom_abline(intercept = 0, slope = 1, color = "green") +
  facet_wrap(vars(found_all))

ggplot(data = singletons_effect, aes(x = simpson_percentile_FALSE, y = simpson_percentile_TRUE)) +
  geom_point(alpha = .1) +
  theme_bw() +
  geom_abline(intercept = 0, slope = 1, color = "green") +
  facet_wrap(vars(found_all))

```

The rarefaction-inflated datasets are strongly // the raw vectors. They have more extreme skewness and evenness values, relative to their feasible sets, than the raw vectors. This is almost always true for evenness, with a little more noise in the skewness signal. But either way, very strong.

<!-- # Manipulations -->
<!-- # MACD -->

<!-- ```{r load macd, include =F} -->
<!-- cache_loc = "macdb" -->

<!-- ## Set up the cache and config -->
<!-- db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", paste0("drake-cache-", cache_loc, ".sqlite"))) -->
<!-- cache <- storr::storr_dbi("datatable", "keystable", db) -->

<!-- all_di_macd <- readd(all_di_obs, cache = cache)  -->
<!-- all_di_macd_manip <- readd(all_di_manip, cache = cache) -->
<!-- macd_dat <- readd(dat_s_dat_macdb, cache = cache) -->

<!-- DBI::dbDisconnect(db) -->
<!-- rm(cache) -->

<!-- all_di_macd <- all_di_macd %>% -->
<!--   mutate(log_nparts = log(gmp:::as.double.bigz(nparts)), -->
<!--          log_nsamples = log(nsamples)) %>% -->
<!--   mutate(prop_found = exp(log_nsamples - log_nparts)) %>% -->
<!--   mutate(found_all = prop_found==1) -->
<!-- #  -->
<!-- # all_di_macd_manip <- all_di_macd_manip %>% -->
<!-- #   mutate(log_nparts = log(gmp:::as.double.bigz(nparts)), -->
<!-- #          log_nsamples = log(nsamples)) %>% -->
<!-- #   mutate(prop_found = exp(log_nsamples - log_nparts)) %>% -->
<!-- #   mutate(found_all = prop_found==1) -->

<!-- ggplot(data = all_di, aes(x = log(s0), y = log(n0))) + geom_point(alpha = .1) + geom_point(data = all_di_macd, aes(x = log(s0), y = log(n0)), color = "blue", alpha = .9) + theme_bw() -->

<!-- ggplot(data = filter(all_di, !singletons), aes(x = log(s0), y = log(n0))) + geom_point(alpha = .1) + geom_point(data = filter(all_di_macd, !singletons), aes(x = log(s0), y = log(n0)), color = "blue", alpha = .9) + theme_bw() -->

<!-- ``` -->

<!-- Here are the distributions of skew and evenness, overall. -->

<!-- ```{r macd overall, include =F} -->

<!-- ggplot(data = all_di_macd, aes(x = skew_percentile)) + -->
<!--   geom_histogram(bins = 100) + -->
<!--   theme_bw() + -->
<!--   facet_wrap(vars(found_all), scales = "free_y") -->

<!-- ggplot(data = all_di_macd, aes(x = simpson_percentile)) + -->
<!--   geom_histogram(bins = 100) + -->
<!--   theme_bw() + -->
<!--   facet_wrap(vars(found_all), scales = "free_y") -->
<!-- ``` -->

<!-- Here is how manipulation affects things: -->

<!-- ```{r macd ctrlcomp, fig.dim = c(4,4), include =F} -->

<!-- ggplot(data = all_di_macd_manip, aes(x = skew_percentile, y = ctrl_skew_percentile)) + -->
<!--   geom_point(alpha = .5) + -->
<!--   #  xlim(0, 100) + -->
<!--   # ylim(0, 100) + -->
<!--   theme_bw() + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") -->

<!-- ggplot(data = all_di_macd_manip, aes(x = simpson_percentile, y = ctrl_skew_percentile)) + -->
<!--   geom_point(alpha = .5) + -->
<!--   xlim(0, 100) + -->
<!--   ylim(0, 100) + -->
<!--   theme_bw() + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") -->

<!-- ggplot(data = all_di_macd_manip, aes(x = simpson_change)) + -->
<!--   geom_histogram() + -->
<!--   theme_bw() -->

<!-- ggplot(data = all_di_macd_manip, aes(x = skew_change)) + -->
<!--   geom_histogram() + -->
<!--   theme_bw() -->

<!-- ggplot(data = all_di_macd_manip, aes(x = ctrl_simpson_percentile, y = simpson_change)) + -->
<!--   geom_point(alpha = .3) + -->
<!--   geom_hline(yintercept = 0) + -->
<!--   theme_bw() -->

<!-- ggplot(data = all_di_macd_manip, aes(x = ctrl_skew_percentile, y = skew_change)) + -->
<!--   geom_point(alpha = .3) + -->
<!--   geom_hline(yintercept = 0) + -->
<!--   theme_bw() -->

<!-- print(t.test(all_di_macd_manip$skew_percentile, all_di_macd_manip$ctrl_skew_percentile, paired = T) -->
<!-- ) -->
<!-- print(t.test(all_di_macd_manip$simpson_percentile, all_di_macd_manip$ctrl_simpson_percentile, paired = T) -->
<!-- ) -->

<!-- print(wilcox.test(all_di_macd_manip$skew_percentile, all_di_macd_manip$ctrl_skew_percentile, paired = T)) -->
<!-- print(wilcox.test(all_di_macd_manip$simpson_percentile, all_di_macd_manip$ctrl_simpson_percentile, paired = T)) -->

<!-- ``` -->

<!-- Change is going to be bounded at 100 and 0: you can't go up or down from there. (Another argument for increasing the number of samples?) -->

<!-- # Portal plant manips -->

<!-- ```{r load pp, include = F} -->
<!-- cache_loc = "portal_plants_manip" -->

<!-- ## Set up the cache and config -->
<!-- db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", paste0("drake-cache-", cache_loc, ".sqlite"))) -->
<!-- cache <- storr::storr_dbi("datatable", "keystable", db) -->

<!-- all_di_p<- readd(all_di_manip, cache = cache)  -->

<!-- all_di_p <- all_di_p %>% -->
<!--   mutate(log_nparts = log(gmp:::as.double.bigz(nparts)), -->
<!--          log_nsamples = log(nsamples)) %>% -->
<!--   mutate(prop_found = exp(log_nsamples - log_nparts)) %>% -->
<!--   mutate(found_all = prop_found==1) -->

<!-- DBI::dbDisconnect(db) -->
<!-- rm(cache) -->
<!-- ``` -->

<!-- Nsamples, singletons -->

<!-- ```{r pp overall, include = F} -->
<!-- #  -->
<!-- # ggplot(data = all_di_p, aes(x = nsamples)) + -->
<!-- #   geom_histogram(bins = 100) + -->
<!-- #   theme_bw() + -->
<!-- #   geom_vline(xintercept = 2000) -->
<!-- #  -->
<!-- # all_di_p <- filter(all_di_p, nsamples >= 2000) -->

<!-- ggplot(data = all_di_p, aes(x = skew_percentile)) + -->
<!--   geom_histogram(bins = 100) + -->
<!--   theme_bw() + -->
<!--   facet_wrap(vars(found_all), scales = "free_y") -->

<!-- ggplot(data = all_di_p, aes(x = simpson_percentile)) + -->
<!--   geom_histogram(bins = 100) + -->
<!--   theme_bw() + -->
<!--   facet_wrap(vars(found_all), scales = "free_y") -->
<!-- #  -->
<!-- # pp_singletons_effect <- all_di_p %>% -->
<!-- #   select(singletons, skew_percentile, simpson_percentile, year, plot, treatment, season) %>% -->
<!-- #   tidyr::pivot_wider(names_from = singletons, values_from = c("skew_percentile", "simpson_percentile")) -->
<!-- #  -->
<!-- # ggplot(data = pp_singletons_effect, aes(x = skew_percentile_FALSE, y  = skew_percentile_TRUE)) + -->
<!-- #   geom_point(alpha = .1) + -->
<!-- #   geom_abline(intercept = 0, slope = 1, color = "green") + -->
<!-- #   theme_bw() -->
<!-- #  -->
<!-- #  -->
<!-- # ggplot(data = pp_singletons_effect, aes(x = simpson_percentile_FALSE, y  = simpson_percentile_TRUE)) + -->
<!-- #   geom_point(alpha = .1) + -->
<!-- #   geom_abline(intercept = 0, slope = 1, color = "green") + -->
<!-- #   theme_bw() -->
<!-- ``` -->

<!-- By treatment, season -->

<!-- ```{r portal trtmt, include =F} -->
<!-- ggplot(data = filter(all_di_p, singletons == F), aes(x = skew_percentile)) + -->
<!--   geom_histogram(bins = 100) + -->
<!--   theme_bw() + -->
<!--   facet_wrap(vars(season, treatment), scales = "free_y") -->

<!-- ggplot(data = filter(all_di_p, singletons == F), aes(x = treatment, y = skew_percentile)) + -->
<!--   geom_boxplot() + -->
<!--   theme_bw() + -->
<!--   facet_wrap(vars(season)) -->

<!-- ggplot(data = filter(all_di_p, singletons == F), aes(x = simpson_percentile)) + -->
<!--   geom_histogram(bins = 100) + -->
<!--   theme_bw() + -->
<!--   facet_wrap(vars(season, treatment), scales = "free_y") -->

<!-- ggplot(data = filter(all_di_p, singletons == F), aes(x = treatment, y = simpson_percentile)) + -->
<!--   geom_boxplot() + -->
<!--   theme_bw() + -->
<!--   facet_wrap(vars(season)) -->
<!-- ``` -->

<!-- <!-- By year --> -->

<!-- <!-- ```{r plants year, fig.height = 25, fig.width = 5} --> -->

<!-- <!-- ggplot(data = filter(all_di_p, singletons == F), aes(x = treatment, y = skew_percentile)) + --> -->
<!-- <!--   geom_boxplot() + --> -->
<!-- <!--   theme_bw() + --> -->
<!-- <!--   facet_grid(rows = vars(year), cols = vars(season), scales = "free_y") --> -->
<!-- <!-- ggplot(data = filter(all_di_p, singletons == F), aes(x = treatment, y = simpson_percentile)) + --> -->
<!-- <!--   geom_boxplot() + --> -->
<!-- <!--   theme_bw() + --> -->
<!-- <!--   facet_grid(rows = vars(year), cols = vars(season), scales = "free_y") --> -->
<!-- <!-- ``` --> -->

<!-- Median -->
<!-- ```{r pp median, fig.width = 12, fig.height = 12, include =F} -->

<!-- pp_median <- all_di_p %>% -->
<!--   filter(singletons == F) %>% -->
<!--   group_by(season, year, treatment) %>% -->
<!--   summarize(skew_percentile = median(skew_percentile), -->
<!--             simpson_percentile = median(simpson_percentile)) %>% -->
<!--   ungroup() %>% -->
<!--   tidyr::pivot_wider(names_from = treatment, values_from = c(skew_percentile, simpson_percentile)) -->

<!-- sk_x <- ggplot(data = pp_median, aes(x = skew_percentile_control, y = skew_percentile_exclosure)) + -->
<!--   geom_point(alpha = .5) + -->
<!--   facet_wrap(vars(season)) + -->
<!--   theme_bw() + -->
<!--   ggtitle("Skew Ctrl v Exclosure") + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") + -->
<!--   xlim(40, 100) + -->
<!--   ylim(40, 100) -->

<!-- sk_r <- ggplot(data = pp_median, aes(x = skew_percentile_control, y = skew_percentile_removal)) + -->
<!--   geom_point(alpha = .5) + -->
<!--   facet_wrap(vars(season)) + -->
<!--   theme_bw() + -->
<!--   ggtitle("Skew Ctrl v Removal") + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") + -->
<!--   xlim(40, 100) + -->
<!--   ylim(40, 100) -->

<!-- gridExtra::grid.arrange(grobs = list(sk_x, sk_r), nrow = 2) -->

<!-- sp_x <- ggplot(data = pp_median, aes(x = simpson_percentile_control, y = simpson_percentile_exclosure)) + -->
<!--   geom_point(alpha = .5) + -->
<!--   facet_wrap(vars(season)) + -->
<!--   theme_bw() + -->
<!--   ggtitle("Simpson Ctrl v Exclosure") + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") + -->
<!--   xlim(0, 40) + -->
<!--   ylim(0, 40) -->

<!-- sp_r <- ggplot(data = pp_median, aes(x = simpson_percentile_control, y = simpson_percentile_removal)) + -->
<!--   geom_point(alpha = .5) + -->
<!--   facet_wrap(vars(season)) + -->
<!--   theme_bw() + -->
<!--   ggtitle("Simpson Ctrl v Removal") + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") + -->
<!--   xlim(0, 40) + -->
<!--   ylim(0, 40) -->

<!-- gridExtra::grid.arrange(grobs = list(sp_x, sp_r), nrow = 2) -->
<!-- ``` -->

```{r nparts v rov}

all_di_plot <- filter(all_di, !singletons)

ggplot(all_di_plot, aes(x = log_nparts, y = simpson_mean)) +
  geom_point(alpha = .1) +
  theme_bw() +
  geom_vline(xintercept = 10)

ggplot(all_di_plot, aes(x = log_nparts, y = simpson_sd)) +
  geom_point(alpha = .1) +
  theme_bw() +
  geom_vline(xintercept = 10)

ggplot(all_di_plot, aes(x = log_nparts, y = abs(simpson_sd/simpson_mean))) +
  geom_point(alpha = .1) +
  theme_bw() +
  geom_vline(xintercept = 10)

ggplot(all_di_plot, aes(x = abs(simpson_sd/simpson_mean), y = simpson_percentile, color = log_nparts > 10)) +
  geom_point(alpha = .05) +
  theme_bw() +
  scale_color_viridis_d(end = .8)

ggplot(filter(all_di_plot, s0 > 3, !is.infinite(skew_mean), !is.infinite(skew_sd)), aes(x = log_nparts, y = skew_mean)) +
  geom_point(alpha = .1) +
  theme_bw() +
  geom_vline(xintercept = 10)

ggplot(filter(all_di_plot, s0 > 3, !is.infinite(skew_mean), !is.infinite(skew_sd)), aes(x = log_nparts, y = skew_sd)) +
  geom_point(alpha = .1) +
  theme_bw() +
  geom_vline(xintercept = 10)

ggplot(filter(all_di_plot, s0 > 3, !is.infinite(skew_mean), !is.infinite(skew_sd)), aes(x = log_nparts, y = abs(skew_sd/skew_mean))) +
  geom_point(alpha = .1) +
  theme_bw() +
  geom_vline(xintercept = 10)

ggplot(filter(all_di_plot, s0 > 3, !is.infinite(skew_mean), !is.infinite(skew_sd)), aes(x = abs(skew_sd/skew_mean), y = skew_percentile, color = log_nparts > 10)) +
  geom_point(alpha = .05) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  xlim(0,4)
```