### 2020-05-01 analysis/reports/95_interval_portal.Rmd
try this

---
title: "Synthesis report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log150_nparts = log(gmp:::as.double.bigz(nparts), base = 150)) %>%
  mutate(prop_found = exp(log_nsamples - log_nparts)) %>%
  mutate(found_all = prop_found==1) %>%
  filter(s0 >= 2,
         n0 > s0,
         dat != "fia_small")
```

# Datasets in S and N space

Here is where our communities fall in S and N space:

```{r datasets in s and n space}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = dat)) +
  geom_point(alpha = .8) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_d()

```

Here is how that translates into the size of the feasible set:

```{r size of fs}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = log150_nparts)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

ggplot(filter(all_di, singletons == F), aes(x = log_nparts, y = ..count..)) +
  geom_density() +
  theme_bw() +
  geom_vline(xintercept = c(10, 15))

ggplot(filter(all_di, singletons == F), aes(x = log_nparts, y = ..count..)) +
  geom_density() +
  theme_bw() +
  geom_vline(xintercept = c(10, 15)) +
  facet_wrap(vars(dat), scales = "free")

```

The small datasets are basically all FIA.

Note that the color scale is log transformed, so the largest communities have e^`r max(all_di$log_nparts)`, or `r exp(max(all_di$log_nparts))`, elements in the feasible set!

# 95 interval vs nparts

```{r 95 interval}

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log(s0), y = log(n0), color = skew_95_ratio)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)+
  facet_wrap(vars(dat), scales = "free")

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = simpson_95_ratio)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)+
  facet_wrap(vars(dat), scales = "free")

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log_nparts, y = skew_95_ratio, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c() +
  facet_wrap(vars(dat), scales = "free")

ggplot(filter(all_di, singletons == F), aes(x = log_nparts, y = simpson_95_ratio, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c()+
  facet_wrap(vars(dat), scales = "free")

```

# Binned by ranked nparts and ranked 95 intervals

```{r binned skew}

all_di_skew <- all_di %>%
  filter(!singletons, s0 > 2) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(skew_95_ratio)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw()
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw()

ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw()

```
```{r binned simpson}

all_di_simpson <- all_di %>%
  filter(!singletons) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(simpson_95_ratio)%>%
  mutate(ratio_rank = row_number()) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))))

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw()
ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw()

```

# Small mcdb

```{r small mcdb}
small_di <- filter(all_di, dat == "mcdb", log_nparts < 20)

ggplot(small_di, aes(x = log_nparts, y = skew_95_ratio)) +
  geom_point() +
  theme_bw()

filter(small_di, skew_95_ratio == min(small_di$skew_95_ratio, na.rm = T))

```

```{r load one site}

site = 1879

library(drake)

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache-mcdb.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)
cache$del(key = "lock", namespace = "session")

loadd(di_fs_1879_TRUE, cache = cache)

twosided_interval <- quantile(filter(di_fs_1879_TRUE, source!= "observed")$skew, probs = c(0.025, 0.975))

onesided_interval <- quantile(filter(di_fs_1879_TRUE, source!= "observed")$skew, probs = c(.95))

ggplot(filter(di_fs_1879_TRUE, source != "observed"), aes(x =skew)) +
  geom_density() +
  geom_vline(xintercept = twosided_interval) +
  geom_vline(xintercept = onesided_interval, color = "red")

# 1469 - more middle of the pack

loadd(di_fs_1469_TRUE, cache = cache)

twosided_interval <- quantile(filter(di_fs_1469_TRUE, source!= "observed")$skew, probs = c(0.025, 0.975))

onesided_interval <- quantile(filter(di_fs_1469_TRUE, source!= "observed")$skew, probs = c(.95))

ggplot(filter(di_fs_1469_TRUE, source != "observed"), aes(x =skew)) +
  geom_density() +
  geom_vline(xintercept = twosided_interval) +
  geom_vline(xintercept = onesided_interval, color = "red")

#1218 - very high

loadd(di_fs_1218_TRUE, cache = cache)

twosided_interval <- quantile(filter(di_fs_1218_TRUE, source!= "observed")$skew, probs = c(0.025, 0.975), na.rm=T)

onesided_interval <- quantile(filter(di_fs_1218_TRUE, source!= "observed")$skew, probs = c(.95), na.rm=T)

ggplot(filter(di_fs_1218_TRUE, source != "observed"), aes(x =skew)) +
  geom_density() +
  geom_vline(xintercept = twosided_interval) +
  geom_vline(xintercept = onesided_interval, color = "red")

```