### 2020-05-17 analysis/reports/figs.Rmd
rewrites

---
title: "Synthesis report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log150_nparts = log(gmp:::as.double.bigz(nparts), base = 150)) %>%
  mutate(prop_found = exp(log_nsamples - log_nparts)) %>%
  mutate(found_all = prop_found==1) %>%
  filter(s0 >= 2,
         n0 > s0,
         dat %in% c("bbs", "gentry", "fia_short", "fia_small", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(substr(dat, 0, 3) == "fia", yes = "fia", no = dat))
```

# Datasets in S and N space

Here is where our communities fall in S and N space:

```{r datasets in s and n space}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = dat)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_d(end = .9)

```

Here is how that translates into the size of the feasible set:

```{r size of fs}
ggplot(filter(all_di, log_nparts <= 100, singletons == F), aes(x = log(s0), y = log(n0), color = log_nparts)) +
  geom_point(alpha = .5) +
  geom_point(data = filter(all_di, log_nparts > 100), color = "yellow", alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

ggplot(filter(all_di, singletons == F), aes(x = log_nparts, y = ..count..)) +
  geom_density() +
  theme_bw() +
  geom_vline(xintercept = c(10, 15))

ggplot(filter(all_di, singletons == F), aes(x = log_nparts, y = ..count..)) +
  geom_density() +
  theme_bw() +
  geom_vline(xintercept = c(10, 15)) +
  facet_wrap(vars(dat), scales = "free")

```

The small datasets are basically all FIA.

Note that the color scale is log transformed, so the largest communities have e^`r max(all_di$log_nparts)`, or `r exp(max(all_di$log_nparts))`, elements in the feasible set!

# 95 interval (one tailed) vs nparts

```{r 95 interval}

skew_1t = ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log(s0), y = log(n0), color = skew_95_ratio_1t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

simpson_1t = ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = simpson_95_ratio_1t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

gridExtra::grid.arrange(grobs = list(skew_1t, simpson_1t), nrow = 1)

```

# Overall percentile values

```{r overall}

ggplot(filter(all_di, !singletons, s0 > 2), aes(x = skew_percentile)) +
  geom_histogram() +
  theme_bw() +
  ggtitle("Skew percentile") +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(filter(all_di, !singletons), aes(x = simpson_percentile)) +
  geom_histogram() +
  theme_bw() +
  ggtitle("Simpson percentile")+
  facet_wrap(vars(dat), scales = "free_y")

```

# Binned by ranked nparts and ranked 95 intervals

```{r binned skew}

all_di_skew <- all_di %>%
  filter(!singletons, s0 > 2) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(skew_95_ratio_2t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: Nparts")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: 95% interval ratio")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```

```{r binned simpson}

all_di_simpson <- all_di %>%
  filter(!singletons) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(simpson_95_ratio_2t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
    ggtitle("1/10th bins: Nparts")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw()+
  ggtitle("1/10th bins: 95% interval ratio")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```

```{r relationships in nparts bins}

ggplot(all_di_skew, aes(x = skew_95_ratio_2t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

ggplot(all_di_simpson, aes(x = simpson_95_ratio_2t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

```

# 95 interval vs nparts

```{r 1t 95 interval}

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log(s0), y = log(n0), color = skew_95_ratio_1t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = simpson_95_ratio_1t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

```

```{r 1t 95 ratio v nparts scatter}

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log_nparts, y = skew_95_ratio_1t, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c()

ggplot(filter(all_di, singletons == F), aes(x = log_nparts, y = simpson_95_ratio_1t, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c()

```

```{r 1t percentiles v 95 interval}

ggplot(filter(all_di, !singletons), aes(x = skew_95_ratio_1t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw()

ggplot(filter(all_di, !singletons), aes(x = simpson_95_ratio_1t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw()

```

# Binned by ranked nparts and ranked 95 intervals

```{r 1t binned skew}

all_di_skew <- all_di %>%
  filter(!singletons, s0 > 2) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(skew_95_ratio_1t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: Nparts")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: 95% interval ratio")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```

```{r 1t binned simpson}

all_di_simpson <- all_di %>%
  filter(!singletons) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(simpson_95_ratio_1t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
    ggtitle("1/10th bins: Nparts")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw()+
  ggtitle("1/10th bins: 95% interval ratio")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```

```{r 1t relationships in nparts bins}

ggplot(all_di_skew, aes(x = skew_95_ratio_1t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

ggplot(all_di_simpson, aes(x = simpson_95_ratio_1t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

```

```{r overall relationships}

ggplot(filter(all_di, !singletons), aes(x = skew_percentile)) +
  geom_histogram() +
  theme_bw()

ggplot(filter(all_di, !singletons), aes(x = simpson_percentile)) +
  geom_histogram() +
  theme_bw()

```

### Remove FIA

```{r overall relationships no fia}

ggplot(filter(all_di, !singletons, dat != "fia_short", dat != "fia_small"), aes(x = skew_percentile)) +
  geom_histogram() +
  theme_bw()

ggplot(filter(all_di, !singletons, dat != "fia_short", dat != "fia_small"), aes(x = simpson_percentile)) +
  geom_histogram() +
  theme_bw()

```

```{r overall relationships by da t}

ggplot(filter(all_di, !singletons, s0 > 2), aes(x = skew_percentile)) +
  geom_histogram() +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(filter(all_di, !singletons), aes(x = simpson_percentile)) +
  geom_histogram() +
  theme_bw()+
  facet_wrap(vars(dat), scales = "free_y")

```

```{r gentry}

ggplot(filter(all_di, dat == "gentry", !singletons), aes(x = log(n0/s0), y = simpson_percentile)) +
  geom_point()

gentry_cat <- filter(all_di, dat == "gentry", !singletons) %>%
  mutate(low_avgn = log(n0/s0) < 1.25)

ggplot(gentry_cat, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(low_avgn), scales = "free_y")

```

```{r 1t hists}

ggplot(all_di_skew, aes(x = skew_95_ratio_1t))  +
geom_histogram() +   theme_bw() +
  facet_wrap(vars(nparts_rank_binned))

ggplot(all_di_skew, aes(x = skew_95_ratio_1t)) +
  geom_histogram(aes(x = skew_percentile / 100, y = ..count.. * 5), fill = "green", alpha = .5) + geom_histogram(alpha = .5) +  
theme_bw() +
  facet_wrap(vars(nparts_rank_binned))

ggplot(all_di_simpson, aes(x = simpson_95_ratio_1t))  +
geom_histogram() +   theme_bw() +
  facet_wrap(vars(nparts_rank_binned))

ggplot(all_di_simpson, aes(x = simpson_95_ratio_1t)) +
  geom_histogram(aes(x = simpson_percentile / 100, y = ..count.. * 5), fill = "green", alpha = .5) + geom_histogram(alpha = .5) +  
theme_bw() +
  facet_wrap(vars(nparts_rank_binned))

```