### 2020-07-31 analysis/howto.Rmd
working

---
title: "How to run"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installations

You will need to download and install hte package for this research compendium, as well as packages to run the analytical workflow and optionally HPC.

```{r install packages, echo = T, eval  = F}
devtools::install_github("diazrenata/scadsanalysis@clean-and-tests")
install.packages("drake")
install.packages("DBI")
install.packages("storr")
install.packages("clustermq")

```

```{r load}
library(scadsanalysis)

```

# Download and prep data

```{r download data, echo = T, eval = F}
download_data()
```

# Get the p tables

Either copy from directory, or run `analysis/helper_scripts/ptables_plan.R`. RMD runs this on a HPC cluster using `submit_p_pipeline.sbatch`. These are very large files.

# Run plans

Either source one of the pipelines directory or run via sbatch.