### 2020-08-05 analysis/reports/less_than_equal.Rmd
move old reporst to old_reports directory

---
title: "Figures and results for main manuscript"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 > s0,
         !singletons,
         dat %in% c("bbs", "fia_short", "fia_small", "gentry", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))
```

# Final dataset in S and N space

```{r final dataset s and n space}

ggplot(all_di, aes(x = log_s0, y = log_n0, color = dat)) +
  geom_point(alpha = .01) +
  geom_point(data = filter(all_di, log_nparts > log(50)), alpha = .4) +
  theme_bw() +
  scale_color_viridis_d() +
  ggtitle("Communities by S and N")

```

```{r skew change}

ggplot(filter(all_di, s0 > 2), aes(x = skew_percentile_excl, y = skew_percentile, color = nsamples)) +
  geom_point() +
  facet_wrap(vars(dat))

all_di <- all_di  %>%
  mutate(skew_change = skew_percentile - skew_percentile_excl,
         even_change = simpson_percentile -
           simpson_percentile_excl)

ggplot(filter(all_di, s0 > 2), aes(x = log_nparts, y = skew_change)) +
  geom_point() +
  xlim(0,15) +
  facet_wrap(vars(dat))

```

```{r hists}

```