### 2020-08-05 analysis/reports/smallness.Rmd
shiny app for size comparisons

---
title: "Size class small comparisons"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

```

```{r real data, include=FALSE}

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 > s0,
         !singletons,
         dat %in% c("bbs", "fia_short", "fia_small", "gentry", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))
```

##### Results for small (FIA sized) communities

Here are the results when we try to tell whether FIA communities differ qualitatively from similarly-sized communities from other datasets...

###### Small sites

Note that, even for small sites, "other datasets" are generally larger than FIA....

```{r small, fig.dim = c(4,3)}

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia = filter(all_di, dat == "fia")

small_di <- all_di %>%
  filter(s0 <= fia_max_s,
         n0 <= fia_max_n) %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"),
         right_corner = FALSE)

for(i in 1:nrow(small_di)) {
  if(small_di$dat[i] != "fia") {
    fia_match_s0 = filter(fia, s0 >= small_di$s0[i])

    max_n0_for_this_s0 = max(fia_match_s0$n0)

    small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0

  }
}

small_di <- small_di %>%
  filter(!right_corner)

ggplot(small_di, aes(x = s0, y = n0, color = log_nparts)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .7) +
  facet_wrap(vars(fia_yn)) +
  theme(legend.position = "top")

ggplot(small_di, aes(log_nparts)) +
  geom_histogram(bins = 30) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y")

```

###### Skew for small sites
```{r skewness small, fig.dim = c(4, 3)}

ggplot(filter(small_di, s0 >2), aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Skew, <=")

ggplot(filter(small_di, s0 >2), aes(x = skew_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Skew, <")

small_di %>%
  filter(s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high_ltet = mean(skew_percentile > 95),
            prop_skew_high_lt = mean(skew_percentile_excl > 95),
            n_skew_sites = dplyr::n())

```

###### Evenness
```{r evenness result small, fig.dim = c(4, 3)}

ggplot(filter(small_di), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Even, <=")

ggplot(filter(small_di), aes(x = simpson_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Even, <")

small_di %>%
  group_by(fia_yn) %>%
  summarize(prop_even_low_ltet = mean(simpson_percentile < 5),
            prop_even_low_lt = mean(simpson_percentile_excl < 5),
            n_even_sites = dplyr::n())

```

For these comparisons, 
- For skewness, the effect is weak for both FIA and not-FIA (5% and 8% in the extremes, respectively)
- For evenness, the effect is much stronger for not-FIA (8 vs 15%)
- < vs <= matters quite a bit
    - Many of these sites can be quite small. Since we flagged a concern about having fewer than, arbitrarily, 100 elements, let's remove those:

###### Skew for small sites, removing very small
```{r skewness small no tiny, fig.dim = c(4, 3)}

small_di <- small_di %>%
  filter(nparts > 100)

ggplot(filter(small_di, s0 >2), aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Skew, <=")

ggplot(filter(small_di, s0 >2), aes(x = skew_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Skew, <")

small_di %>%
  filter(s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high_ltet = mean(skew_percentile > 95),
            prop_skew_high_lt = mean(skew_percentile_excl > 95),
            n_skew_sites = dplyr::n())

```

###### Evenness
```{r evenness result small no tiny, fig.dim = c(4, 3)}

ggplot(filter(small_di), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Even, <=")

ggplot(filter(small_di), aes(x = simpson_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Even, <")

small_di %>%
  group_by(fia_yn) %>%
  summarize(prop_even_low_ltet = mean(simpson_percentile < 5),
            prop_even_low_lt = mean(simpson_percentile_excl < 5),
            n_even_sites = dplyr::n())

```

Removing sites with fewer than 100 elements in the feasible set

- *eliminates the difference between <= and <*
- We see a difference between FIA and other datasets 
    - For skew, 6 vs 12% in the extremes
    - For even, 10 vs 25% in the extremes

I am skeptical of these comparisons. It may be that FIA is behaving qualitatively differently, but it may also be that, because the other datasets are less concentrated at *very small* FS, the effect of FS size still drives the difference even when we filter like this. On the strength of this filtering I'm not confident arguing *either* that FIA is qualitatively different, independent of size, or that these subsets are behaving the same.

I think a stronger comparison would be to break into size classes based on the number of parts.