### 2020-08-06 analysis/reports/smallness.Rmd
DIRECT COMPARISONS, update di_obs test

---
title: "Size class small comparisons"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, message = FALSE) 
```

```{r real data, include=FALSE}

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 > s0,
         !singletons,
         dat %in% c("bbs", "fia_short", "fia_small", "gentry", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))
```

The FIA sites stand out from the other datasets in that, overall, they do not show pronouncedly extreme values. This is especially the case for skewness:

```{r overall hists}

ggplot(filter(all_di, s0 > 2), aes(skew_percentile_excl)) +
  geom_histogram(bins = 20) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(filter(all_di), aes(simpson_percentile)) +
  geom_histogram(bins = 20) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")
```

One possible explanation for this is that the FIA sites tend to have quite small S and N, leading to quite a small feasible set and potentially quite a broad distribution of expected values for skewness and evenness.

It is also possible that the FIA datasets differ qualitatively from other datasets, and that this drives the difference.

We may be able to disentangle these possibilities by

- Comparing the FIA sites to sites from other datasets that are of a similar size
- Looking at the breadth of the distributions for these sites relative to larger sites

If we compare the FIA sites to sites from other datasets that are of a similar size, the comparison is limited by the fact that our sites are nonuniformly distributed in SxN space, or in terms of the the number of elements in their feasible sets. Specifically, the FIA datasets are concentrated towards **especially small** feasible sets compared "other datasets" - even the ones that fall within the general range found in FIA.

```{r small, fig.dim = c(4,3)}

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia = filter(all_di, dat == "fia")

small_di <- all_di %>%
  filter(s0 <= fia_max_s,
         n0 <= fia_max_n) %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"),
         right_corner = FALSE)

for(i in 1:nrow(small_di)) {
  if(small_di$dat[i] != "fia") {
    fia_match_s0 = filter(fia, s0 >= small_di$s0[i])

    max_n0_for_this_s0 = max(fia_match_s0$n0)

    small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0

  }
}

small_di <- small_di %>%
  filter(!right_corner)

ggplot(small_di, aes(x = s0, y = n0, color = log_nparts)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .7) +
  facet_wrap(vars(fia_yn)) +
  theme(legend.position = "top")

ggplot(small_di, aes(log_nparts)) +
  geom_histogram(bins = 30) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y")

```

This means that an overall histogram for FIA would be representative more of sites in the 4-7 parts range, while one for "other datasets" would be representative of more of a spread from 3-16 parts. Given that we suspect the number of parts is quite important, this is a sticky problem.

One option, which I haven't considered before, is to subsample __FIA__ to match "other datasets". 
If we subsample so there is equal representation of all S and N combinations between FIA and "other datasets", we end up with 309 combinations represented and a total of 488 comparisons.

```{r subsampling}

not_fia <- filter(small_di, fia_yn != "fia")
fia = filter(small_di, dat == "fia")

small_di_s_n <- small_di %>%
  select(s0, n0) %>%
  distinct() %>%
  mutate(in_fia = FALSE,
         in_not_fia = FALSE)

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)
  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  small_di_s_n$in_fia[i] = nrow(fia_matches) > 0 
  small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0

}

small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia)

set.seed(1977)

subsamples <- list()

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)

  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches))

  fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F)
  not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F)

  subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ]))

}

sub_di <- bind_rows(subsamples)
# 
# ggplot(sub_di, aes(log(s0), log(n0), color = fia_yn)) +
#   geom_point() +
#   facet_wrap(vars(fia_yn))
# 
# 
# ggplot(sub_di, aes(log_nparts)) +
#   geom_histogram() +
#   facet_wrap(vars(fia_yn))

ggplot(small_di, aes(log(s0), log(n0))) +
  geom_point(color = "gray") +
  theme_bw() +
  geom_point(data = sub_di, alpha = 1)

ggplot(sub_di, aes(log_nparts)) +
  geom_histogram() +
  theme_bw()
```

The dark dots are the sites for which we can find at least one exact match in s0 and n0 between a FIA site and a site from another dataset. The histogram is the distribution of FS sizes represented in the subsample.

Here are the results for that subsample:

```{r subsample results}

ggplot(sub_di, aes(skew_percentile_excl)) +
  geom_histogram()+
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y")

ggplot(sub_di, aes(simpson_percentile)) +
  geom_histogram()+
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y")

sub_di %>%
  filter(s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95),
            prop_even_low = mean(simpson_percentile < 5),
            nsites = dplyr::n())

```

For the subsample that is directly comparable, there is **not a difference** between FIA and other datasets in the distribution of percentile values. This is a very small subsample relative to all the sites in both datasets - FIA has 20,000 sites overall, and we are only looking at 500 of them. However, given how sensitive results aggregated in this way are to the relative frequency of SxN combinations/FS sizes in the sample, I believe this is the most robust way we can ask whether *being from FIA* causes detectably different outcomes.