### 2020-08-07 analysis/reports/rov_metric.Rmd
run this

---
title: "Explaining and exploring the 95% interval metric"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, message = FALSE) 
```

```{r real data, include=FALSE}

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 > s0,
         !singletons,
         s0 > 1,
         dat %in% c("bbs", "fia_short", "fia_small", "gentry", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))
```

One of the crucial threads of reasoning for comparing observations to an emergent statistical expectation is that the expectation be sufficiently specific to allow us to make robust comparisons. *Often*, with sufficiently large systems (also described as sufficiently large scale separation between the subcomponents and the aggregate characteristic of interest; H&L), we can rely on the expectation being extremely narrowly peaked. It is, however, not clear what "sufficiently large" means. We can expect that *some* ecological systems are definitely not large in this sense - if there are only 6 species and 3 individuals, the feasible set is trivially small. Nor is it a given that the expectations do actually get appreciably more peaked with increasing S and N.

We would like to know:

* How the shape, and specifically the narrowness, of the expectation changes over an ecologically-relevant gradient in S and N
* If, as we might expect given the above intuition about larger systems, small systems have less steep peaks: whether the change in the shape of the expectation corresponds with the change in our results. 
    * This is kind of slippery. Ideally we would find a way to measure our statistical power.

To do this we need to establish

* A way of measuring the shape, and specifically the narrowness, of the expectation
* How the shape changes over gradients
* How the shape constrains the kind of deviation we *could* detect (?)
* How the shape maps on to the presence/absence of detected deviations

## Illustrating the expectation

We obtain the expectation by drawing samples from the feasible set and constructing the distribution of values for *summary statistics* for all of those elements.

## Measuring the shape (narrowness)

## Mapping the shape over a gradient in S and N

It can be difficult to interpret the distribution of interval values from our actual datasets, because they are irregularly distributed over S by N space. So we can drop a net over the relevant S by N gradient and sample *regularly* from it.

```{r define net}

all_di_net <- all_di %>%
  select(s0, n0) %>%
  distinct()

ggplot(all_di_net, aes(log(s0), log(n0))) +
  geom_point() +
  geom_vline(xintercept = log(200)) +
  geom_hline(yintercept = c(log(40720)))

net_s0 <- floor(exp(seq(.5, log(200), by = .3))) %>%
  unique()

net_n0 <- 1+ floor(exp(seq(.5, 10, by = .75))) %>%
  unique()

net <- expand.grid(net_s0, net_n0) %>%
  rename(s0 = Var1, n0 = Var2) %>%
  filter(n0 > s0,
         s0 > 1)

ggplot(all_di_net, aes(log(s0), log(n0))) +
  geom_point() +
  geom_vline(xintercept = log(200)) +
  geom_hline(yintercept = c(log(40720))) +
  geom_point(data = net, color = "purple")

```