### 2020-11-28 analysis/reports/filtering.Rmd
final hao edits and start filtering vignette

---
title: "Filtering datasets"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

theme_set(theme_bw())
options(ggplot2.discrete.color = scale_color_viridis_d(end = .8))

knitr::opts_chunk$set(fig.width=3, fig.height=2.5, warning = FALSE, message = FALSE) 
```

The filtering of datasets to exclude very small or very large communities happens at the very beginning of the analysis, when we download and *process* the data.

```{r}

bbs <- read.csv(here::here("working-data", "abund_data", "bbs_spab.csv"),  stringsAsFactors = F, header = F, skip = 2)

colnames(bbs) <- c("site", "year", "species", "abund")

bbs_statevars <- bbs %>%
  group_by(site) %>%
  summarize(
    nyears = length(unique(year)),
    nspecies = dplyr::n(),
    nind = sum(abund)
  )

ggplot(bbs_statevars, aes(nspecies, nind)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("BBS")

unique(bbs_statevars$nyears)

```

No communities in BBS are outside the maxima. All communities (in this dataset) are only represnted with 1 year.

```{r}

gentry <- read.csv(here::here("working-data", "abund_data", "gentry_spab.csv"),  stringsAsFactors = F, header = F, skip = 2)

colnames(gentry) <- c("site", "year", "species", "abund")

gentry_statevars <- gentry %>%
  group_by(site) %>%
  summarize(
    nyears = length(unique(year)),
    nspecies = dplyr::n(),
    nind = sum(abund)
  ) %>%
  mutate(outside_maxima = (nspecies > 200) | (nind > 40720))

ggplot(gentry_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("gentry")

```

Gentry has `r sum(gentry$outside_maxima)` communities, out of `r nrow(gentry)` communities, outside the maxima. RMD just realized, I left these in because they are still sample-able - they have sufficiently low N.

```{r}

miscabund <- read.csv(here::here("working-data", "abund_data", "misc_abund_spab.csv"))

    miscabund <- miscabund %>%
      dplyr::rename(site = Site_ID,
                    abund = Abundance)

miscabund_statevars <- miscabund %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  ) %>%
  mutate(outside_maxima = (nspecies > 200) | (nind > 40720))

ggplot(miscabund_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("miscabund")

sum(is.na(miscabund_statevars$nind))

# how many of those have 0

nrow(filter(miscabund_statevars, is.na(nind), nind_na_rm == 0))

filter(miscabund_statevars, is.na(nind), nind_na_rm >0 )

```

Misc abund has `{r} sum(miscabund_statevars$outside_maxima, na.rm =T)` communities that get removed. They are all removed because they have high numbers of individiuals.

There are 130 datasets where there are a mixture of relative and count abundances included. of those, 124 have 0 individuals in the count. THose were not included. There were **6** that have a mixutre and a nonzero number counted. These were INCLUDED in the initial analysis, with relative abundance species completely removed.

```{r}

miscabund_within <- miscabund_statevars %>%
  filter(nspecies <= 200, nind <= 40720)

miscabund_filtered <- load_dataset("misc_abund_short")

miscabund_sites <- unique(miscabund_within$site)
miscabund_filtered_sites <- unique(miscabund_filtered$site)

included_sites <- as.numeric(miscabund_filtered_sites)

setdiff(included_sites, miscabund_sites)

```

```{r}

fia <- read.csv(here::here("working-data", "abund_data", "fia_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

      colnames(fia) <- c("site", "year", "species", "abund")

fia_statevars <- fia %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  ) %>%
  mutate(outside_maxima = (nspecies > 200) | (nind > 40720),
         n_equals_s = nind == nspecies)
sum(fia_statevars$n_equals_s)

ggplot(fia_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("fia")

```

FIA has no extremely large datasets.

```{r}

ggplot(fia_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_vline(xintercept = 10) +
  ggtitle("fia")

ggplot(fia_statevars, aes(nspecies)) +
  geom_histogram() +
  geom_vline(xintercept = 10)

```

```{r}

mcdb <- read.csv(here::here("working-data", "abund_data", "mcdb_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

      colnames(mcdb) <- c("site", "year", "species", "abund")

mcdb_statevars <- mcdb %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  ) %>%
  mutate(outside_maxima = (nspecies > 200) | (nind > 40720),
         n_equals_s = nind == nspecies)

ggplot(mcdb_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("mcdb")

sum(mcdb_statevars$n_equals_s)
```

```{r}

all_statevars <- miscabund_statevars %>%
  select(-nind) %>%
  rename(nind = nind_na_rm) %>%
  bind_rows(bbs_statevars, fia_statevars, gentry_statevars, mcdb_statevars) %>%
  filter(!outside_maxima) %>%
  mutate(n_equals_s = nind == nspecies)

sum(all_statevars$n_equals_s)

```