### 2020-12-19 analysis/howto.Rmd

---
title: "How to run"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installations

You will need to download and install hte package for this research compendium, as well as packages to run the analytical workflow and optionally HPC.

```{r install packages, echo = T, eval  = F}
devtools::install_github("diazrenata/scadsanalysis@clean-and-tests")
install.packages("drake")
install.packages("DBI")
install.packages("storr")
install.packages("clustermq")

```

```{r load}
library(scadsanalysis)

```

# Download and prep data

```{r download data, echo = T, eval = F}
download_data()
```

# Get the p tables

Either copy from directory, or run `analysis/helper_scripts/ptables_plan.R`. RMD runs this on a HPC cluster using `submit_p_pipeline.sbatch`. These are very large files.

# Run plans

Either source one of the pipelines directory or run via sbatch. Running directly is going to take a lot of CPU.

The plans render dataset-level reports, but these have not been updated in a while.

# Collect all_di

Run `analysis/helper_scripts/make_all_di.R` to collect the results from each of the pipelines into one dataframe.

# Render manuscript-level reports

Running or knitting (an as-yet-uncompiled) doc will reproduce the figures and results summaries presented in the manuscript.

(rarefaction doc) will reproduce the rarefaction analysis in the supplement.

(range of variation doc) will reproduce the range of variation analysis in the supplement.