### 2020-12-19 analysis/reports/accessory/less_than_equal.Rmd

---
title: "Effect of < v <= in percentile definition"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

```

You can define the %ile score as either

- the % of values in a comparison vector that are **less than** a focal value
- the % of values that are **less than or equal to** that value

You see both definitions out in the world.

When I initially coded up the analysis, I used **<**. I didn't think about it much and as I was writing the manuscript I wrote **<=**. So that got me curious if it matters.

It *should not* matter too much as long as there are many unique values in the comparison vector. However, if there are only a handful, and the focal value is equal to an element of the comparison vector, using <= will cause the score to jump up. The jumps will be in increments of 100/n, with n the number of elements in the comparison vector. Meaning if there is only one value, < will give a score of 0 and <= will give a score of 100! And if there are ties, and the focal value is equal to the tie value, the jump will be (100 / n) * m, with m the number of elements in the comparison vector equal to the tie value.

#### Demo with toy data

```{r a handful of values, echo = T}

percentile_lt <- function(focal, comparison) {
  100 * (sum(comparison < focal) / length(comparison))
}

percentile_ltet <- function(focal, comparison) {
  100 * (sum(comparison <= focal) / length(comparison))
}

# Many values

many <- seq(1, 100, by = .1)

focal <- runif(1, 1, 100)
focal <- round(focal, 1) # rounding so focal == a value in many

percentile_lt(focal, many)
percentile_ltet(focal, many)

# A few values
few <- seq(1, 101, by = 20)

focal <- ceiling(focal / 20) * 20 + 1 # again, rounding so focal == a value in few

percentile_lt(focal, few)

percentile_ltet(focal, few)

# A few values with ties
few_with_ties <- c(seq(1, 81, by = 20), 41)

focal <- 41 # if focal == tie value, you get the largest possible jump

percentile_lt(focal, few_with_ties)

percentile_ltet(focal, few_with_ties)

# Just one value

one <- runif(1, 1, 100)

focal <- one # if there is only 1 element in one, focal MUST be that element

percentile_lt(focal, one)
percentile_ltet(focal, one)

```

### For our purposes

We can expect using <= to give us, if anything, higher percentile scores. We can expect this effect to be more pronounced for distributions with relatively few unique values. This is probably strongly correlated with the number of unique elements in the feasible set and the number of unique samples we found from the feasible set, but is not necessarily identical to it because two unique samples can have the same values for skewness or evenness. This will create **ties**, which can cause quite large jumps.

If the comparison vector is short, and/or has repeated values, we may get noticeably higher scores using <= than using <. Usually I think being short has a larger effect than having ties - unless you have a very long vector that is all ties, which doesn't happen in this context. Being short also makes ties more impactful.

Especially if the feasible set is small, the focal value is guaranteed to be an element of the comparison vector. That is because the comparison vector, for small FS, is an exhaustive account of all possible values for the focal value.

#### Change in %ile score

```{r real data, include=FALSE}

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 > s0,
         !singletons,
         dat %in% c("bbs", "fia_short", "fia_small", "gentry", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))
```

##### Skewness
```{r skew change, fig.dim = c(4,3)}

ggplot(filter(all_di, s0 > 2), aes(x = skew_percentile_excl, y = skew_percentile, color = nsamples)) +
  geom_point(alpha = .1) +
  facet_wrap(vars(dat)) +
  scale_color_viridis_c(option = "plasma") +
  theme_bw()

all_di <- all_di  %>%
  mutate(skew_change = skew_percentile - skew_percentile_excl,
         even_change = simpson_percentile -
           simpson_percentile_excl)

ggplot(filter(all_di, s0 > 2), aes(x = nsamples, y = skew_change)) +
  geom_point(alpha = .1) +
  facet_wrap(vars(dat)) +
  theme_bw() +
  geom_vline(xintercept = 150)

```

For skewness,

- Jumps occur and can be quite large, including jumps from 0 to 100!
    - Using < or <= could qualitatively change our interpretation of whether we see overrepresentation in the extremes
    - Because the one-tail for skewness is asking whether observed values are usually **higher** than the comparison vectors, using **<=** will give us a **stronger** apparent effect
- Jumps occur only when the number of samples is low - for the most part, well below 500 and usually below 150 (vertical line on right graph). 
    - So < vs <= is mostly likely to be important for small FS
        - (The number of samples is strongly correlated with the size of the FS, up to a threshold. Currently running an update to `pull_di` to see how the size of FS and # samples is correlated to the # unique values in the comparison vectors.)

##### Evenness
```{r even change, fig.dim = c(4,3)}

ggplot(filter(all_di), aes(x = simpson_percentile_excl, y = simpson_percentile, color = nsamples)) +
  geom_point(alpha = .1) +
  facet_wrap(vars(dat)) +
  scale_color_viridis_c(option = "plasma") +
  theme_bw()

ggplot(filter(all_di), aes(x = nsamples, y = even_change)) +
  geom_point(alpha = .1) +
  facet_wrap(vars(dat)) +
  theme_bw()

```

For evenness, we see the same general outcome as for skewness. However,

- The one-tail for evenness is asking whether the observed values are **lower** than the comparison. So using <= will tend to weaken the effect we perceive.
- The jumps for evenness are confined to even smaller numbers of samples than skewness. 
    - My intuition is that this is because you tend to get ties more often with skewness than with evenness (skewness is a little bit of a strange creature). But I haven't confirmed this.

#### The conservative option

While both <= and < are valid more generally, I think there may be reasoning to prefer one over the other in this context.

Specifically, we want to know how *weird* the observation is. If the observation is equal to a very common value in the comparison vector, it's *not that weird*. So we don't want all those ties to count as values that the observation is more extreme than.

Because we are using less than, and the question for skewness is whether the observation is unusually *high*, the conservative approach for skewness is to use **<**. This will not allow an observation that matches a common tie value to count as more extreme than the ties.

The question for evenness is whether the observation is unusually low. If the observation is equal to a common tie value, using < will put the observation just to the left of those ties, effectively. It will be driven artificially low. Using <=, however, will count those tie elements as values to the left of the observation, and make it look less extreme. Which it is, if it matches all those ties.

So using < for skewness and <= for evenness gives us an appropriately conservative estimate of how weird the focal values are in the case where the focal values match common ties in the comparison vectors.

I am not sure if this logic really addresses the scenario where the comparison vectors do not have ties, but are quite short. My intuition there is that extremely short vectors - roughly on a scale of fewer than 20-100 values - are not very informative in this framework. Using <= and < for skew and even, respectively, is still the conservative tack in terms of detecting deviations.

Also, the impact on "statistical significance" is strongest for very short vectors. That is because, in the absence of ties, the jump increment is 100/n, with n the length of the comparison vector. If we define significance using an arbitrary cutoff of being more extreme than 95% of comparison values, then for the jump to matter it must get you from below 95 to 95 (for skew) or from above 5 to 5 (for even). If there are even 100 values in the comparison vector, the jump should only affect observations landing at the 94th or 6th percentiles. If there are fewer, it becomes more problematic: if there are only 5 values, using the more generous condition will move you from a score of 80 to a score of 100. But if there are only 5 values, this approach is almost certainly not appropriate anyway....(I can't prove it, but I don't think 5 is enough to define an attractor...!)

#### Effects on results

##### Overall results

Here are the overall result histograms and tables using < and <=...

###### Skew
```{r skewness result, fig.dim = c(4, 3)}

ggplot(filter(all_di, s0 >2), aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") +
  ggtitle("Skew, <=")

ggplot(filter(all_di, s0 >2), aes(x = skew_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") +
  ggtitle("Skew, <")

all_di %>%
  filter(s0 > 2) %>%
  group_by(dat) %>%
  summarize(prop_skew_high_ltet = mean(skew_percentile > 95),
            prop_skew_high_lt = mean(skew_percentile_excl > 95),
            n_skew_sites = dplyr::n())

```

###### Evenness
```{r evenness result, fig.dim = c(4, 3)}

ggplot(filter(all_di, s0 >2), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") +
  ggtitle("Even, <=")

ggplot(filter(all_di, s0 >2), aes(x = simpson_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") +
  ggtitle("Even, <")

all_di %>%
  group_by(dat) %>%
  summarize(prop_even_low_ltet = mean(simpson_percentile < 5),
            prop_even_low_lt = mean(simpson_percentile_excl < 5),
            n_even_sites = dplyr::n())

```

**Generally**, we do not see a big difference between < and <= for the overall results. We see the biggest changes for FIA and MCDB:
- For FIA, using the more conservative option (< for skew and <= for even) reduces the % of extreme observations from 7 to 5 for skew and from 11 to 8 for even.
- For MCDB, the reductions are from 16 to 14 for skew and from 39 to 27 for even.

These datasets have the greatest representation of smallish sites. Which leads us to...

##### Results for small (FIA sized) communities

This is most likely to matter for small communities with small FS.

Here are the results when we try to tell whether FIA communities differ qualitatively from similarly-sized communities from other datasets...

###### Small sites

Note that, even for small sites, "other datasets" are generally larger than FIA....

```{r small, fig.dim = c(4,3)}

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia = filter(all_di, dat == "fia")

small_di <- all_di %>%
  filter(s0 <= fia_max_s,
         n0 <= fia_max_n) %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"),
         right_corner = FALSE)

for(i in 1:nrow(small_di)) {
  if(small_di$dat[i] != "fia") {
    fia_match_s0 = filter(fia, s0 >= small_di$s0[i])

    max_n0_for_this_s0 = max(fia_match_s0$n0)

    small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0

  }
}

small_di <- small_di %>%
  filter(!right_corner)

ggplot(small_di, aes(x = s0, y = n0, color = log_nparts)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .7) +
  facet_wrap(vars(fia_yn)) +
  theme(legend.position = "top")

ggplot(small_di, aes(log_nparts)) +
  geom_histogram(bins = 30) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y")

```

###### Skew for small sites
```{r skewness small, fig.dim = c(4, 3)}

ggplot(filter(small_di, s0 >2), aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Skew, <=")

ggplot(filter(small_di, s0 >2), aes(x = skew_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Skew, <")

small_di %>%
  filter(s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high_ltet = mean(skew_percentile > 95),
            prop_skew_high_lt = mean(skew_percentile_excl > 95),
            n_skew_sites = dplyr::n())

```

###### Evenness
```{r evenness result small, fig.dim = c(4, 3)}

ggplot(filter(small_di), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Even, <=")

ggplot(filter(small_di), aes(x = simpson_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Even, <")

small_di %>%
  group_by(fia_yn) %>%
  summarize(prop_even_low_ltet = mean(simpson_percentile < 5),
            prop_even_low_lt = mean(simpson_percentile_excl < 5),
            n_even_sites = dplyr::n())

```

For these comparisons, 
- For skewness, the effect is weak for both FIA and not-FIA (5% and 8% in the extremes, respectively)
- For evenness, the effect is much stronger for not-FIA (8 vs 15%)
- < vs <= matters quite a bit
    - Many of these sites can be quite small. Since we flagged a concern about having fewer than, arbitrarily, 100 elements, let's remove those:

###### Skew for small sites, removing very small
```{r skewness small no tiny, fig.dim = c(4, 3)}

small_di <- small_di %>%
  filter(nparts > 100)

ggplot(filter(small_di, s0 >2), aes(x = skew_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Skew, <=")

ggplot(filter(small_di, s0 >2), aes(x = skew_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Skew, <")

small_di %>%
  filter(s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high_ltet = mean(skew_percentile > 95),
            prop_skew_high_lt = mean(skew_percentile_excl > 95),
            n_skew_sites = dplyr::n())

```

###### Evenness
```{r evenness result small no tiny, fig.dim = c(4, 3)}

ggplot(filter(small_di), aes(x = simpson_percentile)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Even, <=")

ggplot(filter(small_di), aes(x = simpson_percentile_excl)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("Even, <")

small_di %>%
  group_by(fia_yn) %>%
  summarize(prop_even_low_ltet = mean(simpson_percentile < 5),
            prop_even_low_lt = mean(simpson_percentile_excl < 5),
            n_even_sites = dplyr::n())

```

Removing sites with fewer than 100 elements in the feasible set

- *eliminates the difference between <= and <*
- We see a difference between FIA and other datasets 
    - For skew, 6 vs 12% in the extremes
    - For even, 10 vs 25% in the extremes

I am skeptical of these comparisons. It may be that FIA is behaving qualitatively differently, but it may also be that, because the other datasets are less concentrated at *very small* FS, the effect of FS size still drives the difference even when we filter like this. On the strength of this filtering I'm not confident arguing *either* that FIA is qualitatively different, independent of size, or that these subsets are behaving the same.

I think a stronger comparison would be to break into size classes based on the number of parts.