### 2021-02-13 analysis/rev_prototyping/prop_off.Rmd
modifications to metrics to add prop off actual and nsingletons and shannon

---
title: "Shannon diversity"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())
## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache-mcdb.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)
cache$del(key = "lock", namespace = "session")

loadd(fs_1002_FALSE, cache = cache)

loadd(all_di_obs, cache = cache)

```
```{r}

ggplot(filter(fs_1002_FALSE, source == "observed"), aes(rank, abund)) +
  geom_line()

```

```