### 2021-02-14 analysis/rev_prototyping/jackknife.Rmd
fixese

---
title: "Jackknife"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())
library(scadsanalysis)

```
```{r}
mcdb <- load_dataset("mcdb")

mcdb_sv <- get_statevars(mcdb)

a_site <- filter(mcdb, site == 1270)
```

Using community 1270 from MCDB because with S = 8, N = 542 it's not miniscule but it's small enough that things will run easily locally. Results might be on the small end.

```{r}

```
```{r, echo = T}

make_ind_pool_species <- function(a_vect) {

  rep(a_vect[1], times = a_vect[2])

}

sample_ind_pool <- function(ind_pool, nind_to_sample) {
 species = sample(ind_pool, size = nind_to_sample, replace = FALSE)

 return(data.frame(species = species))
}

jacknife <- function(site_df, prop_of_ind = .6, nsims = 10) {

  nind <- sum(site_df$abund)

  nind_to_sample <- ceiling(prop_of_ind * nind)

  nspp <- nrow(site_df)

  ind_pool_mat <- as.matrix(dplyr::select(site_df, rank, abund))

  ind_pool <- unlist(apply(ind_pool_mat, MARGIN = 1, FUN = make_ind_pool_species))

  sampled_ind <- replicate(n = nsims, expr = sample_ind_pool(ind_pool, nind_to_sample), simplify = F)

  sampled_ind = unique(sampled_ind)

# 
#   sampled_ind <- rmultinom(n = nsims, size = nind_to_sample, prob = site_df$abund)

  sampled_ind_df <- dplyr::bind_rows(sampled_ind, .id = "sim") %>%
    dplyr::group_by(sim, species) %>%
    dplyr::summarize(abund = dplyr::n()) %>%
    dplyr::ungroup() %>%
    dplyr::filter(abund > 0) %>%
    dplyr::select(-species) %>%
    dplyr::group_by(sim) %>%
    dplyr::arrange(abund) %>%
    dplyr::mutate(rank = dplyr::row_number()) %>%
    dplyr::ungroup() %>%
    dplyr::arrange(sim, rank)

  site_df_identifiers <- site_df %>%
    select(dat, site, singletons, source) %>%
    distinct()

  sampled_ind_df <- sampled_ind_df %>%
    cbind(site_df_identifiers) %>%
    dplyr::mutate(site = paste0(site, "_jacknife_", sim)) %>%
    dplyr::mutate(sim = -99)

  return(sampled_ind_df)

  }

set.seed(1977)

a_site_jk <- jacknife(a_site)

```

```{r, fig.dim= c(4,3)}

ggplot(a_site, aes(rank, abund)) +
  geom_point() +
  geom_line() +
  ggtitle("Actual")

ggplot(a_site_jk, aes(rank, abund, group = site)) +
  geom_line(alpha = .3) +
  geom_point(alpha = .3) +
  ggtitle("Resamples")

```
```{r}

jacknife_whole_dataset <- function(dataset, prop_of_ind = .6, nsims = 10) {

  sites_list <- unique(dataset$site)

  all_sites <- lapply(sites_list, FUN = function(which_site, whole_dataset) return(dplyr::filter(whole_dataset, site == which_site)), whole_dataset = dataset)

  all_jacknifes <- lapply(all_sites, FUN = jacknife, prop_of_ind = prop_of_ind, nsims = nsims)

  all_jacknifes <- dplyr::bind_rows(all_jacknifes)

  return(all_jacknifes)
}

mcdb_jacknifes <- jacknife_whole_dataset(mcdb)

write.csv(mcdb_jacknifes, here::here("analysis", "rev_prototyping", "jacknifed_datasets", "mcdb_jk.csv"), row.names = F)

```