### 2021-02-22 analysis/rev_prototyping/additional_metrics_results.Rmd
prelim stuff

---
title: "Other metrics"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())

all_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"))

all_di <- all_di %>%
  filter(s0 > 1,
         s0 != n0,
         n0 != s0 + 1)

```

## Proportion off

```{r}

ggplot(filter(all_di), aes(mean_po_comparison_percentile)) +
  geom_histogram() +
  facet_wrap(vars(dat, singletons), scales = "free_y")

all_di %>% group_by(singletons, dat) %>% summarize(po_sig = mean(mean_po_comparison_percentile > 95))

ggplot(all_di, aes(mean_po_comparison_95, mean_po_comparison, color = log(nparts))) +
  geom_point() +
  geom_line(aes(mean_po_comparison_95, mean_po_comparison_95))

all_di %>%
  filter(mean_po_comparison_percentile > 95) %>%
  group_by(singletons) %>%
  summarize(diff = mean(mean_po_comparison - mean_po_comparison_95))

ggplot(all_di, aes(nparts, mean_po_comparison_95_ratio_1t)) +
  geom_point() +
  scale_x_log10()

ggplot(all_di, aes(nparts, mean_po_comparison_95_ratio_1t)) +
  geom_point() +
  scale_x_log10() +
  xlim(0, 10000)

```

```{r}

ggplot(all_di, aes(mean_po_comparison_sims, mean_po_comparison, color = mean_po_comparison_percentile > 95)) +
  geom_point(alpha = .3) +
  facet_wrap(vars(dat, singletons))

ggplot(all_di, aes(mean_po_comparison - mean_po_comparison_sims, color = mean_po_comparison_percentile > 95)) +
  geom_histogram(alpha = .3) +
  facet_wrap(vars(dat, singletons))

```

## Number of singletons

## Shannon diversity