### 2021-02-22 analysis/rev_prototyping/jacknife_results.Rmd

---
title: "Jackknife"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 3, fig.height = 6)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())
library(scadsanalysis)

all_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"), stringsAsFactors = F)

jk_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di_jk.csv"))

jk_di <- jk_di %>%
  group_by_all() %>%
  mutate(site_source = unlist(strsplit(site, "_")[[1]][[1]])) %>%
  ungroup() %>%
  select(dat, site, site_source, s0, n0, skew_percentile_excl, simpson_percentile, shannon_percentile, nsingletons_percentile, nsingletons_percentile_excl, nsingletons, nsingletons_95, mean_po_comparison_percentile, mean_po_comparison_sims, mean_po_comparison_95, skew_95_ratio_1t, simpson_95_ratio_1t, shannon_95_ratio_1t, mean_po_comparison_95_ratio_1t, nparts)

jk_di_means <- jk_di %>%
  group_by(dat, site_source) %>%
  summarise_at(vars(-group_cols(), -site), mean)

all_di_actual <- all_di %>%
    filter(!singletons) %>%
  mutate(site_source = as.character(site)) %>%
  select(dat, site_source, s0, n0, skew_percentile_excl, simpson_percentile, shannon_percentile, nsingletons_percentile, nsingletons_percentile_excl, nsingletons, nsingletons_95, mean_po_comparison_percentile, mean_po_comparison_sims, mean_po_comparison_95, skew_95_ratio_1t, simpson_95_ratio_1t, shannon_95_ratio_1t, mean_po_comparison_95_ratio_1t, nparts)

colnames(all_di_actual)[ 3:19] <- paste0(colnames(all_di_actual)[ 3:19], "_actual")

jk_di_results <- left_join(jk_di, all_di_actual)
jk_di_mean_results <- left_join(jk_di_means, all_di_actual)

jk_di_mean_results <- jk_di_mean_results %>%
  mutate(skew_actual_sig = skew_percentile_excl_actual > 95,
         skew_sig = skew_percentile_excl > 95) %>%
  mutate(skew_change = paste0("actual", skew_actual_sig, "_jk", skew_sig))

```

<!-- ## S and N -->

<!-- ```{r} -->

<!-- ggplot(jk_di_mean_results, aes(s0_actual, s0)) + -->
<!--   geom_point() + -->
<!--   geom_line(aes(s0_actual, s0_actual)) + -->
<!--   facet_wrap(vars(dat), scales = "free") -->

<!-- ggplot(jk_di_mean_results, aes(n0_actual, n0)) + -->
<!--   geom_point() + -->
<!--   geom_line(aes(n0_actual, n0_actual)) + -->
<!--   facet_wrap(vars(dat), scales = "free") -->

<!-- ``` -->

## Skewness

```{r}
ggplot(filter(jk_di_mean_results, s0 > 2), aes(skew_percentile_excl)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

ggplot(filter(jk_di_mean_results, s0 > 2), aes(skew_percentile_excl_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

jk_di_mean_results  %>% 
 filter(s0 > 2) %>%
  group_by(dat) %>%
  summarize(prop_skew_high_jk = mean(skew_percentile_excl > 95),
            prop_skew_high_raw = mean(skew_percentile_excl_actual > 95),
            nsites_included = dplyr::n())
```

## Evenness

```{r}
ggplot(filter(jk_di_mean_results), aes(simpson_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 5)

ggplot(filter(jk_di_mean_results), aes(simpson_percentile_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 5)

jk_di_mean_results  %>% 
  group_by(dat) %>%
  summarize(prop_simpson_high_jk = mean(simpson_percentile < 5),
            prop_simpson_high_raw = mean(simpson_percentile_actual < 5),
            nsites_included = dplyr::n())

```

## Shannon

```{r}
ggplot(filter(jk_di_mean_results), aes(shannon_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 5)

ggplot(filter(jk_di_mean_results), aes(shannon_percentile_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 5)

jk_di_mean_results  %>% 
  group_by(dat) %>%
  summarize(prop_shannon_high_jk = mean(shannon_percentile < 5),
            prop_shannon_high_raw = mean(shannon_percentile_actual < 5),
            nsites_included = dplyr::n())

```

## Percent off

```{r}
ggplot(filter(jk_di_mean_results), aes(mean_po_comparison_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

ggplot(filter(jk_di_mean_results), aes(mean_po_comparison_percentile_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

jk_di_mean_results  %>% 
  group_by(dat) %>%
  summarize(prop_mean_po_comparison_high_jk = mean(mean_po_comparison_percentile > 95),
            prop_mean_po_comparison_high_raw = mean(mean_po_comparison_percentile_actual > 95),
            nsites_included = dplyr::n())

```

## Nsingletons

```{r}
ggplot(filter(jk_di_mean_results), aes(nsingletons_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

ggplot(filter(jk_di_mean_results), aes(nsingletons_percentile_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

jk_di_mean_results  %>% 
  group_by(dat) %>%
  summarize(prop_nsingletons_high_jk = mean(nsingletons_percentile > 95),
            prop_nsingletons_high_raw = mean(nsingletons_percentile_actual > 95),
            nsites_included = dplyr::n())

```

## Nsingletons excl

```{r}
ggplot(filter(jk_di_mean_results), aes(nsingletons_percentile_excl)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

ggplot(filter(jk_di_mean_results), aes(nsingletons_percentile_excl_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

jk_di_mean_results  %>% 
  group_by(dat) %>%
  summarize(prop_nsingletons_high_jk = mean(nsingletons_percentile_excl > 95),
            prop_nsingletons_high_raw = mean(nsingletons_percentile_excl_actual > 95),
            nsites_included = dplyr::n())

```