### 2021-02-22 analysis/rev_prototyping/old/nsingletons.Rmd

---
title: "Shannon diversity"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         #!singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset))

```
```{r}

ggplot(all_di, aes(nsingletons_percentile)) +
  geom_histogram() +
  facet_wrap(vars(dat, singletons), scales = "free_y") +
  xlim(0, 100)

ggplot(all_di, aes(nsingletons_95, nsingletons)) +
  geom_point(alpha = .2) +
  geom_line(aes(nsingletons_95, nsingletons_95)) +
  facet_wrap(vars(dat, singletons), scales = "free")

all_di <- all_di %>%
  mutate(prop_singles_actual = nsingletons / s0,
         sim_prop_singles_95 = nsingletons_95 / s0,
         sim_prop_singles_mean = nsingletons_mean / s0) %>%
  mutate(
            actual_minus_95_prop = prop_singles_actual - sim_prop_singles_95)

all_di %>%
  mutate(high_singles=nsingletons >= nsingletons_95,
         high_singles_ex = nsingletons > nsingletons_95,
         high_singles_mean = nsingletons > nsingletons_mean) %>%
  group_by(dat, singletons) %>%
  summarize(prop_high_singles = mean(high_singles),
            prop_high_singles_ex = mean(high_singles_ex),
            prop_high_singles_mean = mean(high_singles_mean),
            nsites= dplyr::n()) %>%
  ungroup()

```

<!-- ```{r} -->

<!-- ggplot(filter(all_di, nsingletons >= nsingletons_95), aes(nsingletons_mean, nsingletons)) + -->
<!--   geom_point(alpha = .2) + -->
<!-- #  geom_line(aes(nsingletons_mean, nsingletons_mean)) + -->
<!--     geom_line(aes(nsingletons_95, nsingletons_95)) + -->
<!--   facet_wrap(vars(dat), scales = "free") -->

<!-- ggplot(filter(all_di, nsingletons >= nsingletons_95), aes(nsingletons, nsingletons-nsingletons_95)) + -->
<!--   geom_point(alpha = .2) + -->
<!-- #  geom_line(aes(nsingletons_mean, nsingletons_mean)) + -->
<!--   facet_wrap(vars(dat), scales = "free") -->
<!-- ``` -->