### 2021-02-23 analysis/rev_prototyping/jacknife_results.Rmd

---
title: "Jackknife"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 3, fig.height = 6)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())
library(scadsanalysis)

all_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"), stringsAsFactors = F) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1))

jk_di_full <- read.csv(here::here("analysis", "rev_prototyping", "all_di_jk.csv"))

jk_di <- jk_di_full %>%
  group_by_all() %>%
  mutate(site_source = unlist(strsplit(site, "_")[[1]][[1]])) %>%
  ungroup() %>%
  select(dat, site, site_source, s0, n0, skew_percentile_excl,skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile, nsingletons_percentile, nsingletons_percentile_excl, nsingletons, nsingeltons_median, nsingletons_95, mean_po_comparison_percentile, mean_po_comparison_sims, mean_po_comparison_95, skew_95_ratio_1t, simpson_95_ratio_1t, shannon_95_ratio_1t, mean_po_comparison_95_ratio_1t, nparts)

jk_di_means <- jk_di %>%
  group_by(dat, site_source) %>%
  summarise_at(vars(-group_cols(), -site), mean)

all_di_actual <- all_di %>%
  filter(!singletons) %>%
  mutate(site_source = as.character(site)) %>%
  select(dat, site_source, s0, n0, skew_percentile_excl, skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile, nsingletons_percentile, nsingletons_percentile_excl, nsingletons,nsingeltons_median,  nsingletons_95, mean_po_comparison_percentile, mean_po_comparison_sims, mean_po_comparison_95, skew_95_ratio_1t, simpson_95_ratio_1t, shannon_95_ratio_1t, mean_po_comparison_95_ratio_1t, nparts)

colnames(all_di_actual)[ 3:ncol(all_di_actual)] <- paste0(colnames(all_di_actual)[3:ncol(all_di_actual)], "_actual")

jk_di_results <- left_join(jk_di, all_di_actual)
jk_di_mean_results <- left_join(jk_di_means, all_di_actual)

jk_di_mean_results <- jk_di_mean_results %>%
  mutate(skew_actual_sig = skew_percentile_excl_actual > 95,
         skew_sig = skew_percentile_excl > 95) %>%
  mutate(skew_change = paste0("actual", skew_actual_sig, "_jk", skew_sig)) %>%
  filter(nparts > 20)

```

<!-- ## S and N -->

<!-- ```{r} -->

<!-- ggplot(jk_di_mean_results, aes(s0_actual, s0)) + -->
<!--   geom_point() + -->
<!--   geom_line(aes(s0_actual, s0_actual)) + -->
<!--   facet_wrap(vars(dat), scales = "free") -->

<!-- ggplot(jk_di_mean_results, aes(n0_actual, n0)) + -->
<!--   geom_point() + -->
<!--   geom_line(aes(n0_actual, n0_actual)) + -->
<!--   facet_wrap(vars(dat), scales = "free") -->

<!-- ``` -->

## Skewness

```{r}
ggplot(filter(jk_di_mean_results, s0 > 2, nparts > 20), aes(skew_percentile_excl)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

ggplot(filter(jk_di_mean_results, s0 > 2, nparts > 20), aes(skew_percentile_excl_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

jk_di_mean_results  %>% 
  filter(s0 > 2, nparts > 20) %>%
  group_by(dat) %>%
  summarize(prop_skew_high_jk = mean(skew_percentile_excl > 95),
            prop_skew_high_raw = mean(skew_percentile_excl_actual > 95),
            prop_skew_low_jk = mean(skew_percentile < 5),
            prop_skew_low_raw = mean(skew_percentile_actual < 5),
            nsites_included = dplyr::n())
```

## Evenness

```{r}
ggplot(filter(jk_di_mean_results, nparts > 20), aes(simpson_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 5)

ggplot(filter(jk_di_mean_results, nparts > 20), aes(simpson_percentile_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 5)

jk_di_mean_results  %>% 
  group_by(dat) %>%
  filter(nparts > 20) %>%
  summarize(prop_simpson_high_jk = mean(simpson_percentile_excl > 95),
            prop_simpson_high_raw = mean(simpson_percentile_excl_actual > 95),
            prop_simpson_low_jk = mean(simpson_percentile< 5),
            prop_simpson_low_raw = mean(simpson_percentile_actual < 5),
            nsites_included = dplyr::n())

```

## Shannon

```{r}
ggplot(filter(jk_di_mean_results, nparts > 20), aes(shannon_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 5)

ggplot(filter(jk_di_mean_results, nparts > 20), aes(shannon_percentile_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 5)

jk_di_mean_results  %>% 
  group_by(dat) %>%
  filter(nparts > 20) %>%
  summarize(prop_shannon_high_jk = mean(shannon_percentile > 95),
            prop_shannon_high_raw = mean(shannon_percentile_actual > 95),
            prop_shannon_low_jk = mean(shannon_percentile< 5),
            prop_shannon_low_raw = mean(shannon_percentile_actual < 5),
            nsites_included = dplyr::n())

```

## Percent off

```{r}
ggplot(filter(jk_di_mean_results, nparts > 20), aes(mean_po_comparison_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

ggplot(filter(jk_di_mean_results, nparts > 20), aes(mean_po_comparison_percentile_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

jk_di_mean_results  %>% 
  group_by(dat) %>%
  filter(nparts > 20) %>%
  summarize(prop_mean_po_comparison_high_jk = mean(mean_po_comparison_percentile > 95),
            prop_mean_po_comparison_high_raw = mean(mean_po_comparison_percentile_actual > 95),
            nsites_included = dplyr::n())

```

<!-- ## Nsingletons -->

<!-- ```{r} -->
<!-- ggplot(filter(jk_di_mean_results, nparts > 20), aes(nsingletons_percentile_excl)) + -->
<!--   geom_histogram(bins = 100) + -->
<!--   facet_wrap(vars(dat), scales = "free_y", ncol = 1) + -->
<!--   geom_vline(xintercept = 95) -->

<!-- ggplot(filter(jk_di_mean_results, nparts > 20), aes(nsingletons_percentile_actual)) + -->
<!--   geom_histogram(bins = 100) + -->
<!--   facet_wrap(vars(dat), scales = "free_y", ncol = 1) + -->
<!--   geom_vline(xintercept = 95) -->

<!-- jk_di_mean_results  %>%  -->
<!--   group_by(dat) %>% -->
<!--   summarize(prop_nsingletons_high_jk = mean(nsingletons_percentile > 95), -->
<!--             prop_nsingletons_high_raw = mean(nsingletons_percentile_actual > 95), -->
<!--             nsites_included = dplyr::n()) -->

<!-- ``` -->

## Nsingletons excl

Visualization and analysis of nsingletons has a little more nuance than the others, because there are often relatively few values for nsingletons at all. 80% of sites have fewer than 20 singletons as the 95th percentile, which is just a rough way of saying that a **lot** of these are going to be sensitive to whether you define the percentiles as > or >=.

In general the strict > percentile will give you an (appropriately) conservative estimate of how many are extraordinarily **high** and the >= will give you an appropriate estimate of how many are unusually **low**. For most metrics it doesn't really matter, writ large, which you use, because ties are rare. For this one, you get large numbers of sites where a lot of values are = to the observed values, and the >= decision will therefore give you a jump of a lot of percentile scores.

For the "proportions high/low" calculations, we use *the > percentile for high* and *the >= percentile for low*. For visualization, because we are interested in both unusually high and unusually low scores, we can't just pick one or the other. The histogram using >= is reliable at the high end but has a misleading spike at 0, and vice versa. I am making these plots using the *mean*, which doesn't have the misleading spikes at the extremes but does smear things out a little bit.

```{r, fig.dim = c(5,5)}

jk_di_mean_results <- jk_di_mean_results %>%
  group_by_all() %>%
  mutate(singletons_percentile_jump = nsingletons_percentile - nsingletons_percentile_excl,
         singletons_percentile_jump_actual = nsingletons_percentile_actual - nsingletons_percentile_excl_actual,
         nsingletons_percentile_mean = sum(nsingletons_percentile, nsingletons_percentile_excl) / 2,
         nsingletons_percentile_mean_actual = sum(nsingletons_percentile_actual, nsingletons_percentile_excl_actual)/2) %>%
  ungroup()

```

```{r}
ggplot(filter(jk_di_mean_results, nparts > 20), aes(nsingletons_percentile_mean)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

ggplot(filter(jk_di_mean_results, nparts > 20), aes(nsingletons_percentile_mean_actual)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 95)

jk_di_mean_results  %>% 
  group_by(dat) %>%
  filter(nparts > 20) %>%
  summarize(prop_nsingletons_high_jk = mean(nsingletons_percentile_excl > 95),
            prop_nsingletons_high_raw = mean(nsingletons_percentile_excl_actual > 95),
         #   prop_nsingletons_mean_high_raw = mean(nsingletons_percentile_mean_actual > 95),
      #      prop_nsingletons_mean_high_jk = mean(nsingletons_percentile_mean > 95),
            prop_nsingletons_low_jk = mean(nsingletons_percentile < 5),
            prop_nsingletons_low_raw = mean(nsingletons_percentile_actual  < 5),
            nsites_included = dplyr::n())

```

<!-- ```{r, fig.dim = c(6,6)} -->

<!-- jk_di_mean_results <- jk_di_mean_results %>% -->
<!--  #select(-singletons_percentile_excl_high_insig, -singletons_percentile_low_insig) -->
<!--   group_by_all() %>% -->
<!--   mutate(prop_singletons = nsingletons / s0, -->
<!--          prop_singletons_actual = nsingletons_actual / s0, -->
<!--          singletons_percentile_high_change = paste0("actual", (nsingletons_percentile_excl_actual > 95), "_jk", (nsingletons_percentile_excl > 95)), -->
<!--          singletons_percentile_low_change = paste0("actual", (nsingletons_percentile_actual <5), "_jk", (nsingletons_percentile < 5))) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(prop_singletons_change = prop_singletons - prop_singletons_actual) -->

<!-- ggplot(jk_di_mean_results, aes(prop_singletons_actual, prop_singletons, color =singletons_percentile_high_change)) + geom_point() + geom_line(aes(prop_singletons_actual,prop_singletons_actual), inherit.aes = F) -->

<!-- ggplot(jk_di_mean_results, aes(nsingletons_actual, nsingletons,color = singletons_percentile_high_change)) + geom_point() + geom_line(aes(nsingletons_actual,nsingletons_actual), inherit.aes = F) +facet_wrap(vars(dat), scales = "free") + theme(legend.position = "top") -->

<!-- ggplot(jk_di_mean_results, aes(s0,n0, color =  -->
<!-- ggplot(jk_di_mean_results, aes(nsingletons_actual, nsingletons,color = singletons_percentile_high_change)) + geom_point() + geom_line(aes(nsingletons_actual,nsingletons_actual), inherit.aes = F) +facet_wrap(vars(dat), scales = "free") + theme(legend.position = "top") -->
<!-- )) + geom_point() +scale_x_log10() +scale_y_log10() + -->
<!--   facet_wrap(vars(dat), scales = "free") + -->
<!--   theme(legend.position = "top") -->

<!-- ggplot(jk_di_mean_results, aes(s0,n0, color = singletons_percentile_low_change)) + geom_point() +scale_x_log10() +scale_y_log10()+ -->
<!--   facet_wrap(vars(dat), scales = "free") + -->
<!--   theme(legend.position = "top") -->

<!-- ggplot(jk_di_mean_results, aes(n0/s0, nsingletons_percentile_excl, color = singletons_percentile_high_change)) + geom_point()+ -->
<!--   facet_wrap(vars(dat), scales = "free") + -->
<!--   theme(legend.position = "top") -->
<!-- ``` -->