### 2021-03-08 analysis/rev_prototyping/distinguish.Rmd

---
title: "Comparing to lognormal, Poisson, exponential"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, message = FALSE)

theme_set(theme_bw())

```

```{r}

tiny_sad_draws <- feasiblesads::sample_fs(3, 10, 1)
# 
# tiny_sad_tally <- feasiblesads::tally_sets(tiny_sad_draws)
# 
# count_if_species_distinguishable <- function(sad) {
#   
#   tallies <- data.frame(
#     species = unique(sad)
#   )
#   
#   for(i in 1:nrow(tallies)) {
#   tallies$abund[i] = sum(sad == tallies$species[i])
#   }
#   
#   uniques <- data.frame(
#     abund = unique(tallies$abund)
#   )
# 
#   for(i in 1:nrow(uniques)) {
#     uniques$nspp[i] = sum(tallies$abund == uniques$abund[i])
#   }
#   
#   ties <- filter(uniques, nspp > 1)
#     
#   
#   for(i in 1:nrow(ties)) {
#     ties$norders <- factorial(ties$nspp)
#   }
#   
#   nways <- 
# }

```

Say we care which species has which abundance, and to start, we can tell them apart even if they have different abundances. 
Then each element of the feasible set can be reshuffled however many ways there are to order S things, which is S!.

1,2,7
1,7,2
2,1,7
2,7,1
7,1,2
7,2,1

but also

2,4,4
2,4,4
4,2,4
4,2,4
4,4,2
4,4,2

If we can't tell them apart if they have different abundances...
you need to do some offset based on the number of ties.

2,4,4
4,2,4
4,4,2

Within a feasible set this would down-weight hte probability of landing on a FS with ties in species' abundances. Depending on whether those elements are more commonly the ones with long rare tails, or ones with lots of abundant species - which I think may depend on N/S - this would push things to be more or less even.