### 2021-03-11 analysis/rev_prototyping/revision_results.Rmd

---
title: "Updated results"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, message = FALSE)

theme_set(theme_bw())

all_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"), stringsAsFactors = F)

all_ct <-  read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"), stringsAsFactors = F)

fia_ct <- read.csv(here::here("fia_cts.csv"))

all_ct <- rbind(all_ct, fia_ct)

all_ct <- all_ct %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  left_join(all_ct)

all_di <- all_di %>%
  group_by_all() %>%
  mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl)) %>%
  ungroup()

all_di <- all_di %>%
  mutate(in_fia = ifelse(Dataset == "FIA", "FIA", "Other datasets"))

```

## Comparing observed to the baseline

### Central tendency

```{r}
# 
# ggplot(all_di, aes(sim_pos_from_best)) +
#   geom_histogram() +
#   facet_wrap(vars(dat), scales = "free_y") 
# 
# ggplot(all_di, aes(real_po_percentile_excl)) + geom_histogram() +
#   facet_wrap(vars(dat), scales = "free_y") + geom_vline(xintercept = 95)

ggplot(all_di, aes(real_po_percentile_mean)) + geom_histogram() +
  facet_wrap(vars(dat), scales = "free_y") + geom_vline(xintercept = 95)

all_di %>%
  group_by(dat) %>%
  summarize(prop_high_po = mean(real_po_percentile_excl > 95, na.rm = T),
            nsites = dplyr::n())

all_di %>%
  group_by(in_fia) %>%
  summarize(prop_high_po = mean(real_po_percentile_excl > 95, na.rm = T),
            nsites = dplyr::n())

ggplot(all_di, aes(sim_pos_from_best, real_po, color = real_po_percentile_excl > 95)) +
  geom_point(alpha = .3) +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "top") +
  scale_color_viridis_d(end = .8) +
  facet_wrap(vars(dat), scales = "fixed")

ggplot(all_di, aes(sim_pos_from_best, real_po, color = real_po_percentile_excl > 95)) +
  geom_point(alpha = .3) +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "top") +
  scale_color_viridis_d(end = .8)# +
#  facet_wrap(vars(dat), scales = "fixed")

filter(all_di, real_po_percentile_excl > 95) %>%
  mutate(real_to_mean_sim_po_ratio = real_po / sim_pos_from_best,
         real_to_mean_sim_po_difference = real_po - sim_pos_from_best,
         real_to_mean_sim_po_logratio = log(real_po / sim_pos_from_best)) %>%
  select(real_to_mean_sim_po_ratio, real_to_mean_sim_po_difference) %>%
  summarize_all(.funs = list(min = min,max = max))
```

### Shape metrics

```{r, fig.dim = c(12,12)}

all_di_shapes_long <- all_di  %>%
  filter(nparts > 40) %>%
  select(dat, skew_percentile, skew_percentile_excl, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl) %>%
  rename(skew = skew_percentile,
         skewX = skew_percentile_excl,
         simp = simpson_percentile,
         simpX = simpson_percentile_excl,
         shan = shannon_percentile,
         shanX = shannon_percentile_excl,
         nsng = nsingletons_percentile,
         nsngX = nsingletons_percentile_excl) %>%
  mutate(skewM = (skew + skewX) / 2,
         simpM = (simp+ simpX)/2,
         shanM = (shan+ shanX)/2,
         nsngM = (nsng+ nsngX)/2) %>%
  tidyr::pivot_longer(-dat, names_to = "metric", values_to = "value") %>%
  mutate(metric_name = substr(metric, 0, 4),
         metric_mod = ifelse(substr(metric, 5, 5) == "X", "excl", ifelse(substr(metric, 5,5) == "M", "mean", "incl")))

ggplot(filter(all_di_shapes_long, metric_mod == "mean"), aes(value)) +
  geom_histogram(bins = 40) +
  #facet_grid(rows = vars(dat), cols = vars(metric_name), scales = "free", switch = "y") +
  facet_wrap(vars(metric_name, dat), scales = "free", ncol = 5) +
  geom_vline(xintercept = c(2.5, 97.5))

all_di_shapes_high <- all_di_shapes_long %>%
  filter(metric_mod == "excl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_high = mean(value > 95))

all_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = dat, names_from = metric_name, values_from = proportion_high) %>% mutate(dir = "HIGH")

all_di_shapes_low <- all_di_shapes_long %>%
  filter(metric_mod == "incl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_low = mean(value <5))

all_di_shapes_low %>%
  tidyr::pivot_wider(id_cols = dat, names_from = metric_name, values_from = proportion_low) %>%
  mutate(dir = "LOW")

```
<!-- ``` -->
<!-- ```{r} -->

<!-- ggplot(all_di, aes((real_po / sim_pos_from_best), real_po_percentile_excl, color =  real_po_percentile_excl > 95)) + -->
<!--   geom_point(alpha = .3)  -->

<!-- all_di %>% -->
<!--   mutate(rat = real_po / sim_pos_from_best, -->
<!--          dev = real_po_percentile_excl > 95) %>% -->
<!--   group_by(dev) %>% -->
<!--   summarize(minrat = min(rat), -->
<!--             maxrat = max(rat), -->
<!--             meanrat = mean(rat), -->
<!--             mindis = min(real_po - sim_pos_from_best), -->
<!--             maxdis = max(real_po - sim_pos_from_best)) -->

<!-- ggplot(all_di, aes(nparts, sim_pos_from_best)) + -->
<!--   geom_point() + -->
<!--   scale_x_log10() -->

<!-- ggplot(all_di, aes(nparts, sim_pos_from_best, color = dat)) + -->
<!--   geom_point() + -->
<!--   scale_x_log10()  + -->
<!--   scale_color_viridis_d() + -->
<!--   theme(legend.position = "top") -->

<!-- ggplot(all_di, aes(s0, n0, color =  sim_pos_from_best)) + -->
<!--   geom_point() + -->
<!--   scale_x_log10()  + -->
<!--   scale_y_log10() + -->
<!--   scale_color_viridis_c() + -->
<!--   theme(legend.position = "top") -->
<!-- ``` -->