### 2021-03-16 analysis/rev_prototyping/manuscript_main_rev.Rmd
whoff

---
title: "Figures and results for main manuscript"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())

all_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"), stringsAsFactors = F)

all_ct <-  read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"), stringsAsFactors = F)

fia_ct <- read.csv(here::here("fia_cts.csv"))

all_ct <- rbind(all_ct, fia_ct)
all_ct <- all_ct %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  left_join(all_ct)

all_di <- all_di %>%
  group_by_all() %>%
  mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl),
         skew_percentile_mean = mean(skew_percentile, skew_percentile_excl),
        simpson_percentile_mean = mean(simpson_percentile,simpson_percentile_excl),
         shannon_percentile_mean = mean(shannon_percentile, shannon_percentile_excl),
         nsingletons_percentile_mean = mean(nsingletons_percentile, nsingletons_percentile_excl),) %>%
  ungroup()

all_di <- all_di %>%
  mutate(in_fia = ifelse(Dataset == "FIA", "FIA", "Other datasets"))

```

# Datasets by S and N (Figure 1)

```{r, fig.dim = c(7,7)}

s4_fig1 <- ggplot(all_di, aes(x = s0, y = n0, color = Dataset)) +
  geom_point(alpha = .1) +
  geom_point(data = filter(all_di), alpha = .4) +
  theme_bw() +
  scale_color_viridis_d(end = .9) +
  ggtitle("Communities by S and N") +
  xlab("Species richness (S); note log scale") +
  ylab("Total abundance (N); note log scale") +
  scale_x_log10() +
  scale_y_log10()

plot(s4_fig1)
```

# Illustrations of 95% interval (Figure 2)

To show the 95% interval, we need to load the distribution of shape metric values from the samples from the feasible set for a few communities. See rov_metric.md.

<!-- ```{r, fig.dim = c(7,7)} -->
<!-- library(drake) -->

<!-- db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache-net.sqlite")) -->
<!-- cache <- storr::storr_dbi("datatable", "keystable", db) -->
<!-- cache$del(key = "lock", namespace = "session") -->

<!-- net_summary <- readd(all_di_summary, cache = cache) -->

<!-- net_summary <- net_summary %>% -->
<!--   mutate(log_nparts = log(gmp:::as.double.bigz(nparts))) -->

<!-- example_fs <- readd(fs_s_44_n_13360, cache = cache) -->

<!-- example_di <- readd(di_fs_s_44_n_13360, cache =cache) -->

<!-- example_fs <- example_fs %>% -->
<!--   left_join(example_di) %>% -->
<!--   left_join(net_summary) -->

<!-- example_fs2 <- readd(fs_s_13_n_315, cache = cache) -->

<!-- example_di2 <- readd(di_fs_s_13_n_315, cache =cache) -->

<!-- example_fs2 <- example_fs2 %>% -->
<!--   left_join(example_di2) %>% -->
<!--   left_join(net_summary) -->

<!-- example_fs3 <- readd(fs_s_4_n_34, cache = cache) -->

<!-- example_di3 <- readd(di_fs_s_4_n_34, cache =cache) -->

<!-- example_fs3 <- example_fs3 %>% -->
<!--   left_join(example_di3) %>% -->
<!--   left_join(net_summary) -->

<!-- breadth_plots <- list( -->

<!--   ggplot(example_fs3, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .25) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Small community", subtitle = paste0("S = ", (example_fs3$s0), "; N = ", (example_fs3$n0[1]))) + -->
<!--     theme(legend.position = "right") + -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance"), -->

<!--   ggplot(example_fs3, aes(skew)) + -->
<!--   #  geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs3$skew_97p5[1], example_fs3$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle = paste0("Breadth index: ", round((example_fs3$skew_95_ratio_2t[1]), 2))) + -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count"), -->

<!--   ggplot(example_fs2, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .1) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Medium community", subtitle = paste0("S = ", (example_fs2$s0), "; N = ", (example_fs2$n0[1]))) + -->
<!--     theme(legend.position = "right")+ -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance") + -->
<!--     ylim(0, 200), # Remove 3 sads that make the axes too big to be interpretable -->

<!--   ggplot(example_fs2, aes(skew)) + -->
<!--   #  geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs2$skew_97p5[1], example_fs2$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle =  paste0("Breadth index: ", round((example_fs2$skew_95_ratio_1t[1]), 2)))+ -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count"), -->

<!--   ggplot(example_fs, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .1) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Large community", subtitle = paste0("S = ", (example_fs$s0), "; N = ", (example_fs$n0[1]))) + -->
<!--     theme(legend.position = "right") + -->
<!--     ylim(0, 4000) + # Remove a very few very very uneven SADs that make the scale too big to be interpretable -->
<!--     theme(axis.text.y = element_text(size = 6, angle = 60))+ -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance"), -->

<!--   ggplot(example_fs, aes(skew)) + -->
<!--    # geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs$skew_97p5[1], example_fs$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle =  paste0("Breadth index: ", round((example_fs$skew_95_ratio_1t[1]), 2)))+ -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count") -->

<!-- ) -->

<!-- fig_1 <- gridExtra::grid.arrange(grobs = breadth_plots, ncol = 2, top = textGrob("Figure 1", gp = gpar(fill = "white"))) -->

<!-- plot(fig_1) -->

<!-- DBI::dbDisconnect(db) -->
<!-- rm(cache) -->
<!-- rm(db) -->
<!-- ``` -->

# Dissimilarity (Supplement)

```{r, fig.dim = c(3, 6)}
plot_percentile_hist <- function(di_df, col_name, plot_name, tails = 2, facetvar = "Dataset", min_s0 = 0) {

  if(tails == 2) {
    cutoff_percentiles = c(2.5, 97.5)
    min_nparts = 40
  } else if (tails== 1) {
    cutoff_percentiles = c(95)
    min_nparts = 20
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[col_name]]),
           facetvar = di_df[[facetvar]])

  ggplot(filter(di_df, nparts > min_nparts, s0 >= min_s0), aes(response)) +
    geom_histogram(bins = 40, boundary = 100) +
    theme_bw() +
    xlab("") +
    ylab("") +
    geom_vline(xintercept = cutoff_percentiles, color = "red") +
    ggtitle( plot_name) +
    facet_wrap(vars(facetvar), ncol = 1, scales = "free_y")+ 
    theme(plot.title = element_text(size=10))

}

all_di <- all_di %>%
  mutate(`Observed \npercentile score` = ifelse(real_po_percentile_excl > 95, "High (> 95)", "Less than 95"))

```

```{r, fig.dim = c(4, 4)}
ggplot(all_di, aes(sim_pos_from_best, real_po, color = `Observed \npercentile score`)) +
  geom_point(alpha = .3) + 
  geom_point(data = filter(all_di, s0 < 0)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_color_viridis_d(end = .8, direction = -1) +
  xlab("Mean of elements from feasible set") +
  ylab("Observed value") +
 ggtitle("Dissimilarity to the central tendency") +  theme(legend.position = "bottom")#,
# 
# 
# ggplot(filter(all_di, nparts < 10 ^ 50), aes(nparts, sim_pos_from_best, color = Dataset)) +
#   geom_point(alpha = .3) + 
#   geom_point(data = filter(all_di, s0 < 0)) +
#   scale_color_viridis_d(end = .9) +
#   xlab("Number of elements in the feasible set") +
#   ylab("Mean dissimilarity of feasible set to central tendency") +
#   scale_x_log10() +
#   ggtitle("Over the size of the feasible set")
# ), ncol = 1,
#top = textGrob("Dissimilarity to central tendency"))

```

# Metric histograms by dataset (Figure 4)

```{r first hists, fig.dim = c(9,5)}

fig_2 <- gridExtra::grid.arrange(grobs = list(

plot_percentile_hist(all_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(all_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(all_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
   plot_percentile_hist(all_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset")#,
  # plot_percentile_hist(all_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
  ), ncol = 4,
top = textGrob("Figure 2", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))
```

# Results table

```{r}
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
   group_by(Dataset) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
   group_by(Dataset) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

cols2 <- c("Dataset", "Low proportion of rare species", "Low skew", "High Simpson", "High Shannon")

all_results %>%
  select(cols2)

```

# Narrowness

```{r, fig.dim = c(3, 6)}

plot_narrowness <- function(di_df, yvar, yvar_name) {

  if(grepl("sim_pos", yvar)) {
    min_nparts = 20
  } else {
    min_nparts = 40
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[yvar]])) %>%
    filter(nparts > min_nparts,
           nparts < 10 ^ 50)

  ggplot(di_df, aes(nparts, response, color = Dataset)) +
  geom_point(alpha = .3) +
  geom_point(data = filter(di_df, s0 < 0)) +
  scale_color_viridis_d(end = .9) +
  xlab("") +
  ylab("") +
  scale_x_log10() +
  ggtitle(yvar_name) + 
    theme(legend.position = "none")+ 
    theme(plot.title = element_text(size=10))
}

plot_narrowness_legend <- function(di_df) {

  legend_df <- di_df %>%
    select(Dataset) %>%
    distinct() %>%
    mutate(ymark = dplyr::row_number())

    ggplot(legend_df, aes(1, ymark, color = Dataset)) +
      geom_label(aes(x = 2, y = ymark, label = Dataset)) +
      geom_point(size = 4) +
  scale_color_viridis_d(end = .9) +
  xlab("") +
  ylab("") +
    theme(legend.position = "none") +
      xlim(1, 3) + theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
      ) +
      ggtitle("Dataset")

}
# 
# gridExtra::grid.arrange(grobs = list(
#   plot_narrowness(all_di, "sim_pos_from_best", "Dissimilarity to central tendency"),
#   plot_narrowness_legend(all_di)),
#   ncol = 1)
```

```{r, fig.dim = c(7,4)}

gridExtra::grid.arrange(grobs = list(
  plot_narrowness(all_di, "sim_pos_from_best", "Dissimilarity \nto central tendency") + theme(legend.position = "bottom", legend.direction = "vertical" , legend.margin = margin(b = .5, r = 0, t = 0, l = 0, unit = "in")
), gridExtra::grid.arrange(grobs = list(
  plot_narrowness(all_di, "skew_95_ratio_2t", "\nSkewness"),
  plot_narrowness(all_di, "simpson_95_ratio_2t", "\nSimpson evenness"),
    plot_narrowness(all_di, "shannon_95_ratio_2t", "\nShannon diversity"),
    plot_narrowness(all_di, "nsingletons_95_ratio_2t", "\nNumber of rare species")
  ), ncol = 2)), ncol = 2, widths = c(2, 4))

```

<!-- # Breadth index by dataset (Figure 3) -->

<!-- ```{r, fig.dim = c(7,7)} -->

<!-- fig_3 <- gridExtra::grid.arrange(grobs = list( -->
<!--   ggplot(filter(all_di, s0 > 2, nparts > 20), aes(x= skew_95_ratio_1t)) + -->
<!--     geom_histogram() + -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y", ncol = 1) + -->
<!--     xlab("") + -->
<!--     ylab("") + -->
<!--     ggtitle("Skewness") + -->
<!--     xlim(0, 1), -->

<!--   ggplot(filter(all_di, nparts > 20), aes(x = simpson_95_ratio_1t)) + -->
<!--     geom_histogram() + -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y", ncol = 1) + -->
<!--     xlab("") + -->
<!--     ylab("") + -->
<!--     ggtitle("Evenness") + -->
<!--     xlim(0, 1) -->
<!-- ), ncol = 2, -->
<!-- top = textGrob("Figure 3", gp = gpar(fill = "white")), -->
<!-- left = "Number of communities",  -->
<!-- bottom = "Breadth index (ranges 0-1, with 1 being very broad)") -->

<!-- plot(fig_3) -->

<!-- ``` -->

<!-- # Comparison of FIA and comparably sized communities (Figure 4) -->

<!-- ```{r, fig.dim = c(7,7)} -->

<!-- fia_max_s = max(filter(all_di, dat == "fia")$s0) -->
<!-- fia_max_n = max(filter(all_di, dat == "fia")$n0) -->

<!-- fia = filter(all_di, dat == "fia") -->

<!-- small_di <- all_di %>% -->
<!--   filter(s0 <= fia_max_s, -->
<!--          n0 <= fia_max_n) %>% -->
<!--   mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"), -->
<!--          right_corner = FALSE) -->

<!-- for(i in 1:nrow(small_di)) { -->
<!--   if(small_di$dat[i] != "fia") { -->
<!--     fia_match_s0 = filter(fia, s0 >= small_di$s0[i]) -->

<!--     max_n0_for_this_s0 = max(fia_match_s0$n0) -->

<!--     small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0 -->

<!--   } -->
<!-- } -->

<!-- small_di <- small_di %>% -->
<!--   filter(!right_corner) -->
<!-- not_fia <- filter(small_di, fia_yn != "fia") -->
<!-- fia = filter(small_di, dat == "fia") -->

<!-- small_di_s_n <- small_di %>% -->
<!--   select(s0, n0) %>% -->
<!--   distinct() %>% -->
<!--   mutate(in_fia = FALSE, -->
<!--          in_not_fia = FALSE) -->

<!-- for(i in 1:nrow(small_di_s_n)) { -->

<!--   this_s0 = small_di_s_n$s0[i] -->
<!--   this_n0 = small_di_s_n$n0[i] -->

<!--   fia_matches = filter(fia, s0 == this_s0, n0 == this_n0) -->
<!--   not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0) -->

<!--   small_di_s_n$in_fia[i] = nrow(fia_matches) > 0  -->
<!--   small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0  -->

<!-- } -->

<!-- small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia) -->

<!-- set.seed(1977) -->

<!-- subsamples <- list() -->

<!-- for(i in 1:nrow(small_di_s_n)) { -->

<!--   this_s0 = small_di_s_n$s0[i] -->
<!--   this_n0 = small_di_s_n$n0[i] -->

<!--   fia_matches = filter(fia, s0 == this_s0, n0 == this_n0) -->

<!--   not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0) -->

<!--   n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches)) -->

<!--   fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F) -->
<!--   not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F) -->

<!--   subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ])) -->

<!-- } -->

<!-- sub_di <- bind_rows(subsamples) %>% -->
<!--   mutate(Dataset = ifelse(fia_yn == "fia", "FIA", "Other datasets")) -->

<!-- simpson_ks <- filter(sub_di, nparts > 40) %>% -->
<!--   select(simpson_95_ratio_1t, Dataset, simpson_percentile, s0, n0) %>%  -->
<!--   group_by(Dataset, s0, n0) %>% -->
<!--   mutate(pairing = dplyr::row_number()) %>% -->
<!--   ungroup() %>% -->
<!--   tidyr::pivot_wider(id_cols = c(s0, n0, pairing), names_from = Dataset, values_from = c(simpson_95_ratio_1t, simpson_percentile)) -->

<!-- simpson_breadth <- ks.test(simpson_ks$simpson_95_ratio_1t_FIA, simpson_ks$`simpson_95_ratio_1t_Other datasets`) -->

<!-- simpson_percent <- ks.test(simpson_ks$simpson_percentile_FIA, simpson_ks$`simpson_percentile_Other datasets`) -->
<!-- simpson_breadth -->
<!-- simpson_percent -->

<!-- skewness_ks <- filter(sub_di, nparts > 40) %>% -->
<!--   select(skew_95_ratio_1t, Dataset, skew_percentile, s0, n0) %>%  -->
<!--   group_by(Dataset, s0, n0) %>% -->
<!--   mutate(pairing = dplyr::row_number()) %>% -->
<!--   ungroup() %>% -->
<!--   tidyr::pivot_wider(id_cols = c(s0, n0, pairing), names_from = Dataset, values_from = c(skew_95_ratio_1t, skew_percentile)) -->

<!-- skewness_breadth <- ks.test(skewness_ks$skew_95_ratio_1t_FIA, skewness_ks$`skew_95_ratio_1t_Other datasets`) -->

<!-- skewness_percent <- ks.test(skewness_ks$skew_percentile_FIA, skewness_ks$`skew_percentile_Other datasets`) -->
<!-- skewness_breadth -->
<!-- skewness_percent -->

<!-- fig_4 <- gridExtra::grid.arrange(grobs = list( -->
<!--   ggplot(filter(sub_di, nparts > 20), aes(skew_95_ratio_1t)) + -->
<!--     geom_histogram()+ -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y")  + -->
<!--     xlab("Breadth index") + -->
<!--     ylab("Number of communities") + -->
<!--     ggtitle("Breadth index", subtitle = "Skewness"), -->

<!--   ggplot(filter(sub_di, nparts > 20), aes(simpson_95_ratio_1t)) + -->
<!--     geom_histogram()+ -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y")  + -->
<!--     xlab("Breadth index") + -->
<!--     ylab("") + -->
<!--     ggtitle("", subtitle = "Evenness"), -->

<!--   ggplot(filter(sub_di, nparts > 20), aes(skew_percentile_excl)) + -->
<!--     geom_histogram()+ -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y")  + -->
<!--     xlab("Percentile rank") + -->
<!--     ylab("Number of communities") + -->
<!--     ggtitle("Percentile rank", subtitle = "Skewness"), -->

<!--   ggplot(filter(sub_di, nparts > 20), aes(simpson_percentile)) + -->
<!--     geom_histogram()+ -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y")  + -->
<!--     xlab("Percentile rank") + -->
<!--     ylab("")+ -->
<!--     ggtitle("", subtitle = "Evenness")), ncol = 2, top = textGrob("Figure 4", gp = gpar(fill = "white"))) -->

<!-- plot(fig_4) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- pdf("figure_1.pdf", title = "Figure 1", bg = "white", paper = "letter", width = 6.5, height = 6.5) -->
<!-- gridExtra::grid.arrange(grobs = breadth_plots, ncol = 2, top = textGrob("Figure 1", gp = gpar(fill = "white"))) -->
<!-- dev.off() -->

<!-- pdf("figure_2.pdf", title = "Figure 2", bg = "white", paper = "letter", width = 6.5, height = 6.5) -->
<!--  gridExtra::grid.arrange(grobs = list( -->

<!--   ggplot(filter(all_di, s0 > 2, nparts > 20), aes(skew_percentile_excl)) + -->
<!--     geom_histogram(bins = 100) + -->
<!--     geom_vline(xintercept = 95, color = "red") + -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y", ncol = 1) + ggtitle("Skewness") + -->
<!--     xlab("") + -->
<!--     ylab(""),# + -->
<!--   #  xlab("Percentile rank of observed value relative to feasible set") + -->
<!--     #ylab("Number of communities") , -->

<!--   ggplot(filter(all_di, nparts > 20), aes(simpson_percentile)) +  -->
<!--     geom_histogram(bins = 100) + -->
<!--     geom_vline(xintercept = 5, color = "red") + -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y", ncol = 1) + ggtitle("Evenness") +  -->
<!--     xlab("") + -->
<!--     ylab("")#+ -->
<!--     #xlab("Percentile rank of observed value relative to feasible set") + -->
<!--   #  ylab("Number of communities") -->
<!-- ), ncol = 2, -->
<!-- top = textGrob("Figure 2", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), -->
<!-- bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white")) -->
<!--  dev.off() -->

<!-- pdf("figure_3.pdf", title = "Figure 3", bg = "white", paper = "letter", width = 6.5, height = 6.5) -->
<!-- gridExtra::grid.arrange(grobs = list( -->
<!--   ggplot(filter(all_di, s0 > 2, nparts > 20), aes(x= skew_95_ratio_1t)) + -->
<!--     geom_histogram() + -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y", ncol = 1) + -->
<!--     xlab("") + -->
<!--     ylab("") + -->
<!--     ggtitle("Skewness") + -->
<!--     xlim(0, 1), -->

<!--   ggplot(filter(all_di, nparts > 20), aes(x = simpson_95_ratio_1t)) + -->
<!--     geom_histogram() + -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y", ncol = 1) + -->
<!--     xlab("") + -->
<!--     ylab("") + -->
<!--     ggtitle("Evenness") + -->
<!--     xlim(0, 1) -->
<!-- ), ncol = 2, -->
<!-- top = textGrob("Figure 3", gp = gpar(fill = "white")), -->
<!-- left = "Number of communities",  -->
<!-- bottom = "Breadth index (ranges 0-1, with 1 being very broad)") -->

<!-- dev.off() -->

<!-- pdf("figure_4.pdf", title = "Figure 4", bg = "white", paper = "letter", width = 6.5, height = 6.5) -->
<!--  gridExtra::grid.arrange(grobs = list( -->
<!--   ggplot(filter(sub_di, nparts > 20), aes(skew_95_ratio_1t)) + -->
<!--     geom_histogram()+ -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y")  + -->
<!--     xlab("Breadth index") + -->
<!--     ylab("Number of communities") + -->
<!--     ggtitle("Breadth index", subtitle = "Skewness"), -->

<!--   ggplot(filter(sub_di, nparts > 20), aes(simpson_95_ratio_1t)) + -->
<!--     geom_histogram()+ -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y")  + -->
<!--     xlab("Breadth index") + -->
<!--     ylab("") + -->
<!--     ggtitle("", subtitle = "Evenness"), -->

<!--   ggplot(filter(sub_di, nparts > 20), aes(skew_percentile_excl)) + -->
<!--     geom_histogram()+ -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y")  + -->
<!--     xlab("Percentile rank") + -->
<!--     ylab("Number of communities") + -->
<!--     ggtitle("Percentile rank", subtitle = "Skewness"), -->

<!--   ggplot(filter(sub_di, nparts > 20), aes(simpson_percentile)) + -->
<!--     geom_histogram()+ -->
<!--     theme_bw() + -->
<!--     facet_wrap(vars(Dataset), scales = "free_y")  + -->
<!--     xlab("Percentile rank") + -->
<!--     ylab("")+ -->
<!--     ggtitle("", subtitle = "Evenness")), ncol = 2, top = textGrob("Figure 4", gp = gpar(fill = "white"))) -->
<!-- dev.off() -->
<!-- ``` -->