### 2021-03-22 analysis/rev_prototyping/draft_reports/appendices/s7_complete_resampling_results.Rmd

---
title: "Complete results of resampling"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    # github_document:
    #    df_print: kable
    #    toc: true
    word_document:
        df_print: kable
        toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())

all_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"), stringsAsFactors = F)

all_ct <-  read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"), stringsAsFactors = F)

#fia_ct <- read.csv(here::here("fia_cts.csv"))

#all_ct <- rbind(all_ct, fia_ct)
all_ct <- all_ct %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  left_join(all_ct)

all_di <- all_di %>%
  group_by_all() %>%
  mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl),
         skew_percentile_mean = mean(skew_percentile, skew_percentile_excl),
         simpson_percentile_mean = mean(simpson_percentile,simpson_percentile_excl),
         shannon_percentile_mean = mean(shannon_percentile, shannon_percentile_excl),
         nsingletons_percentile_mean = mean(nsingletons_percentile, nsingletons_percentile_excl),) %>%
  ungroup()

all_di <- all_di %>%
  mutate(in_fia = ifelse(Dataset == "FIA", "FIA", "Other datasets"))

```

```{r, fig.dim = c(3, 6)}
plot_percentile_hist <- function(di_df, col_name, plot_name, tails = 2, facetvar = "Dataset", min_s0 = 0) {

  if(tails == 2) {
    cutoff_percentiles = c(2.5, 97.5)
    min_nparts = 40
  } else if (tails== 1) {
    cutoff_percentiles = c(95)
    min_nparts = 20
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[col_name]]),
           facetvar = di_df[[facetvar]])

  ggplot(filter(di_df, nparts > min_nparts, s0 >= min_s0), aes(response)) +
    geom_histogram(bins = 40, boundary = 100) +
    theme_bw() +
    xlab("") +
    ylab("") +
    geom_vline(xintercept = cutoff_percentiles, color = "red") +
    ggtitle( plot_name) +
    facet_wrap(vars(facetvar), ncol = 1, scales = "free_y")+
    theme(plot.title = element_text(size=10))

}

all_di <- all_di %>%
  mutate(`Observed \npercentile score` = ifelse(real_po_percentile_excl > 95, "High (> 95)", "Less than 95"))

```

```{r, fig.dim = c(3, 6)}

plot_narrowness <- function(di_df, yvar, yvar_name) {

  if(grepl("sim_pos", yvar)) {
    min_nparts = 20
    ylabel = "Mean dissimilarity"
  } else {
    min_nparts = 40
    ylabel = "Breadth index"
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[yvar]])) %>%
    filter(nparts > min_nparts,
           nparts < 10 ^ 50)

  ggplot(di_df, aes(nparts, response, color = Dataset)) +
    geom_point(data = filter(di_df, dat == "fia"), alpha = .1) +
    geom_point(data = filter(di_df, dat != "fia"), alpha = .1) +
    geom_point(data = filter(di_df, s0 < 0)) +
    scale_color_viridis_d(end = .9) +
    xlab("") +
    ylab(ylabel) +
    scale_x_log10() +
    ggtitle(yvar_name) +
    theme(legend.position = "none")+
    theme(plot.title = element_text(size=10))
}

plot_narrowness_legend <- function(di_df) {

  legend_df <- di_df %>%
    select(Dataset) %>%
    distinct() %>%
    mutate(ymark = dplyr::row_number())

  ggplot(legend_df, aes(1, ymark, color = Dataset)) +
    geom_label(aes(x = 2, y = ymark, label = Dataset)) +
    geom_point(size = 4) +
    scale_color_viridis_d(end = .9) +
    xlab("") +
    ylab("") +
    theme(legend.position = "none") +
    xlim(1, 3) + theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    ) +
    ggtitle("Dataset")

}
#
# gridExtra::grid.arrange(grobs = list(
#   plot_narrowness(all_di, "sim_pos_from_best", "Dissimilarity to central tendency"),
#   plot_narrowness_legend(all_di)),
#   ncol = 1)
```

```{r}

plot_narrowness_hist <- function(di_df, col_name, plot_name, facetvar = "Dataset", min_s0 = 0) {

  if(grepl("sim_pos", col_name)) {
    xvarname = "Mean dissimilarity"
    min_nparts = 40
  } else {
    min_nparts = 20
    xvarname = "Breadth index"
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[col_name]]),
           facetvar = di_df[[facetvar]])

  ggplot(filter(di_df, nparts > min_nparts, s0 >= min_s0), aes(response)) +
    geom_histogram(bins = 40, boundary = 100) +
    theme_bw() +
    ylab("") +
    xlab(xvarname) +
    ggtitle( plot_name) +
    facet_wrap(vars(facetvar), ncol = 1, scales = "free_y")+
    theme(plot.title = element_text(size=10))

}

```

# Resampling

```{r}

all_di_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"))

jk_di_full <- read.csv(here::here("analysis", "rev_prototyping", "all_di_jk.csv"))

all_ct_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"))

jk_ct <- read.csv(here::here("analysis", "rev_prototyping", "all_ct_jk.csv"))

jk_di_full <- left_join(jk_di_full, jk_ct)

all_di_with_resamples <- left_join(all_di_with_resamples, all_ct_with_resamples)

jk_di <- jk_di_full %>%
  group_by_all() %>%
  mutate(site_source = unlist(strsplit(site, "_")[[1]][[1]])) %>%
  ungroup() %>%
  select(dat, site, site_source, s0, n0, skew_percentile_excl,skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl, real_po_percentile, real_po_percentile_excl, sim_pos_from_best, skew_95_ratio_2t, simpson_95_ratio_2t, shannon_95_ratio_2t, nparts)

skewcols <- colnames(jk_di)[ grepl("skew", colnames(jk_di))]

jk_di[ which(jk_di$s0 < 3), skewcols] <- NA

jk_di_means <- jk_di %>%
  filter(n0 != s0,
         s0 != 1,
         n0 != (s0 + 1)) %>%
  group_by(dat, site_source) %>%
  mutate(njks = dplyr::n(),
         njks_skew = sum(!is.na(skew_percentile)))%>%
  summarise_at(vars(-group_cols(), -site), mean, na.rm =  T) %>%
  mutate(resampling = "Subsampling") %>%
  mutate(site = as.numeric(site_source)) %>%
#  select(-site_source) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20,
         njks == 10) %>%
  ungroup()

jk_di[ which(jk_di$njks_skew < 10), skewcols] <- NA

skewcols <- colnames(all_di_with_resamples[ which(grepl("skew", colnames(all_di_with_resamples)))])

all_di_with_resamples[ which(all_di_with_resamples$s0 < 3), skewcols] <- NA

all_di_with_resamples <- all_di_with_resamples %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
       #  !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  mutate(resampling = ifelse(singletons, "Rare species", "Raw")) %>%
  bind_rows(jk_di_means)

metric_rs_results <- all_di_with_resamples %>%
  group_by(Dataset, resampling) %>%
  summarize(
    skew_high = mean(skew_percentile_excl > 97.5, na.rm = T),
    skew_low = mean(skew_percentile < 2.5, na.rm = T),
    even_high = mean(simpson_percentile_excl > 97.5, na.rm = T),
    even_low = mean(simpson_percentile < 2.5, na.rm = T),
    shannon_high = mean(shannon_percentile_excl > 97.5, na.rm = T),
    shannon_low = mean(shannon_percentile < 2.5, na.rm = T),
    nsingletons_high = mean(nsingletons_percentile_excl > 97.5, na.rm = T),
    nsingletons_low = mean(nsingletons_percentile < 2.5, na.rm = T),
    nskew = sum(!is.na(skew_percentile)),
    nother = dplyr::n()
  ) %>%
  ungroup()

diss_rs_results <- all_di_with_resamples %>%
  filter(nparts > 40) %>%
  group_by(Dataset, resampling) %>%
  summarize(dissimilarity_high = mean(real_po_percentile_excl > 95, na.rm = T),
            ndiss = sum(!is.na(real_po_percentile_excl))) %>%
  ungroup()

rs_results <- left_join(metric_rs_results, diss_rs_results)

#metric_cols <- colnames(rs_results [ grepl("_", colnames(rs_results))])

as.perc <- function(x) {
  return((signif(100 * x, digits = 2)))
}

rs_results_desc <- rs_results %>%
  mutate(`High dissimilarity` = paste0(as.perc(dissimilarity_high), "%; n =", ndiss),
         `High proportion of rare species` = paste0(as.perc(nsingletons_high), "%; n =", nother),
         `High skew` = paste0(as.perc(skew_high), "%; n = ", nskew),
         `Low Simpson` = paste0(as.perc(even_low), "%; n = ", nother),
         `Low Shannon` = paste0(as.perc(shannon_low), "%; n = ", nother),
         `Low proportion of rare species` = paste0(as.perc(nsingletons_low), "%; n =", nother),
         `Low skew` = paste0(as.perc(skew_low), "%; n = ", nskew),
         `High Simpson` = paste0(as.perc(even_high), "%; n = ", nother),
         `High Shannon` = paste0(as.perc(shannon_high), "%; n = ", nother),
         `Resampling scheme` = ordered(resampling, levels = c("Raw", "Subsampling", "Rare species"))) %>%
  arrange(Dataset, `Resampling scheme`)

cols1 <- c("Dataset", "Resampling scheme", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

cols2 <- c("Dataset","Resampling scheme",  "Low proportion of rare species", "Low skew", "High Simpson", "High Shannon")

```
```{r, fig.dim = c(6,4)}
rs_results_long <- rs_results %>%
  tidyr::pivot_longer((-c("Dataset","resampling")), names_to = "metric", values_to = "value") %>%
  filter(grepl("_", metric)) %>%
  mutate(ometric = ordered(metric, levels = c("skew_high", "skew_low", "even_low", "even_high", "shannon_low", "shannon_high", "nsingletons_high", "nsingletons_low", "dissimilarity_high")),
         ometric2 = ordered(metric, levels = c("dissimilarity_high",  "nsingletons_high", "skew_high",  "even_low",  "shannon_low",  "nsingletons_low","skew_low","even_high","shannon_high"))) %>%
  mutate(weird_dir = as.numeric(ometric2) > 5,
         percentile_cutoff = ifelse(grepl("diss", metric), .05, .025),`Resampling scheme` = ordered(resampling, levels = c("Raw", "Subsampling", "Rare species")))

metric_names <- rs_results_long %>%
  select(ometric2) %>%
  distinct() %>%
  arrange(ometric2) %>%
  mutate(metric_desc =(c("High dissimilarity", "High proportion \nof rare species", "High skew", "Low Simpson", "Low Shannon", "Low proportion \nof rare species", "Low skew", "High Simpson", "High Shannon"))) %>%
  mutate(metric_desc = ordered(metric_desc))

rs_results_long <- left_join(rs_results_long, metric_names)
#
# ggplot(filter(rs_results_long, !weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
#   geom_point() +
#   facet_wrap(vars(Dataset), scales = "fixed", ncol = 1) +
#   geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
#   scale_color_viridis_d(end = .8) +
#   ylab("Proportion of extreme observed values") +
#   xlab("")
```

# Percentile histograms

## Subsampling

```{r, fig.dim = c(10,5)}

all_di_with_resamples <- all_di_with_resamples %>%
  group_by_all() %>%
  mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl, na.rm = T),
         skew_percentile_mean = mean(skew_percentile, skew_percentile_excl, na.rm = T),
         simpson_percentile_mean = mean(simpson_percentile,simpson_percentile_excl, na.rm = T),
         shannon_percentile_mean = mean(shannon_percentile, shannon_percentile_excl, na.rm = T),
         nsingletons_percentile_mean = mean(nsingletons_percentile, nsingletons_percentile_excl)) %>%
  ungroup()

fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Subsampling results", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))
```

## Adjusting for rare species

```{r, fig.dim = c(10,5)}

fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Adjusted for rare species results", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))
```

# Summary of effects on proportion of extreme values

## Usual direction

```{r}
gridExtra::grid.arrange(grobs = list(
  ggplot(filter(rs_results_long, !grepl("Misc", Dataset), !weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol =2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("Proportion of extreme observed values") +ylim(0, .7) +
  xlab("") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) ,
  ggplot(filter(rs_results_long, grepl("Misc", Dataset), !weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol = 2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("") +
  xlab("") + ylim(0, .7) +
  theme(legend.position = "bottom", legend.direction = "vertical" , legend.margin = margin(b = .3, r = 0, t = 0, l = 0, unit = "in"), axis.text.x = element_text(angle = 90))), ncol = 2, widths = c( 4,2), bottom = textGrob("Metric"))
```

## Unusual direction

```{r}
gridExtra::grid.arrange(grobs = list(
  ggplot(filter(rs_results_long, !grepl("Misc", Dataset), weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol =2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("Proportion of extreme observed values") +ylim(0, .7) +
  xlab("") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) ,
  ggplot(filter(rs_results_long, grepl("Misc", Dataset), weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol = 2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("") +
  xlab("") + ylim(0, .7) +
  theme(legend.position = "bottom", legend.direction = "vertical" , legend.margin = margin(b = .3, r = 0, t = 0, l = 0, unit = "in"), axis.text.x = element_text(angle = 90))), ncol = 2, widths = c( 4,2), bottom = textGrob("Metric"))

```

# Table of proportions of extreme values

## Usual direction

```{r}

rs_results_desc %>%
  select(cols1)

```

## Unusual direction

```{r}
rs_results_desc %>%
  select(cols2)
```