### 2021-03-22 analysis/rev_prototyping/old/revision_results.Rmd

---
title: "Updated results"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, message = FALSE)

theme_set(theme_bw())

all_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"), stringsAsFactors = F)

all_ct <-  read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"), stringsAsFactors = F)

fia_ct <- read.csv(here::here("fia_cts.csv"))

all_ct <- rbind(all_ct, fia_ct)
all_ct <- all_ct %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  left_join(all_ct)

all_di <- all_di %>%
  group_by_all() %>%
  mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl)) %>%
  ungroup()

all_di <- all_di %>%
  mutate(in_fia = ifelse(Dataset == "FIA", "FIA", "Other datasets"))

```

## Datasets

```{r, fig.dim = c(8,6)}

ggplot(all_di, aes(s0, n0, color = Dataset)) +
  geom_point(alpha = .3) +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_viridis_d(end = .8)

```

## Comparing observed to the baseline

### Central tendency

```{r}
# 
# ggplot(all_di, aes(sim_pos_from_best)) +
#   geom_histogram() +
#   facet_wrap(vars(dat), scales = "free_y") 
# 
# ggplot(all_di, aes(real_po_percentile_excl)) + geom_histogram() +
#   facet_wrap(vars(dat), scales = "free_y") + geom_vline(xintercept = 95)

ggplot(all_di, aes(real_po_percentile_mean)) + geom_histogram() +
  facet_wrap(vars(dat), scales = "free_y") + geom_vline(xintercept = 95)

all_di %>%
  group_by(dat) %>%
  summarize(prop_high_po = mean(real_po_percentile_excl > 95, na.rm = T),
            nsites = dplyr::n())

all_di %>%
  group_by(in_fia) %>%
  summarize(prop_high_po = mean(real_po_percentile_excl > 95, na.rm = T),
            nsites = dplyr::n())

ggplot(all_di, aes(sim_pos_from_best, real_po, color = real_po_percentile_excl > 95)) +
  geom_point(alpha = .3) +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "top") +
  scale_color_viridis_d(end = .8) +
  facet_wrap(vars(dat), scales = "fixed")

ggplot(all_di, aes(sim_pos_from_best, real_po, color = real_po_percentile_excl > 95)) +
  geom_point(alpha = .3) +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "top") +
  scale_color_viridis_d(end = .8)# +
#  facet_wrap(vars(dat), scales = "fixed")

filter(all_di, real_po_percentile_excl > 95) %>%
  mutate(real_to_mean_sim_po_ratio = real_po / sim_pos_from_best,
         real_to_mean_sim_po_difference = real_po - sim_pos_from_best,
         real_to_mean_sim_po_logratio = log(real_po / sim_pos_from_best)) %>%
  select(real_to_mean_sim_po_ratio, real_to_mean_sim_po_difference) %>%
  summarize_all(.funs = list(min = min,max = max))
```

### Shape metrics

```{r, fig.dim = c(12,12)}

all_di_shapes_long <- all_di  %>%
  filter(nparts > 40) %>%
  #mutate(dat = ifelse(nparts < 1000, paste0(dat, "_small"), dat)) %>%
  select(dat, skew_percentile, skew_percentile_excl, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl) %>%
  rename(skew = skew_percentile,
         skewX = skew_percentile_excl,
         simp = simpson_percentile,
         simpX = simpson_percentile_excl,
         shan = shannon_percentile,
         shanX = shannon_percentile_excl,
         nsng = nsingletons_percentile,
         nsngX = nsingletons_percentile_excl) %>%
  mutate(skewM = (skew + skewX) / 2,
         simpM = (simp+ simpX)/2,
         shanM = (shan+ shanX)/2,
         nsngM = (nsng+ nsngX)/2) %>%
  tidyr::pivot_longer(-dat, names_to = "metric", values_to = "value") %>%
  mutate(metric_name = substr(metric, 0, 4),
         metric_mod = ifelse(substr(metric, 5, 5) == "X", "excl", ifelse(substr(metric, 5,5) == "M", "mean", "incl")))

ggplot(filter(all_di_shapes_long, metric_mod == "mean"), aes(value)) +
  geom_histogram(bins = 40) +
  #facet_grid(rows = vars(dat), cols = vars(metric_name), scales = "free", switch = "y") +
  facet_wrap(vars(metric_name, dat), scales = "free", ncol = 5) +
  geom_vline(xintercept = c(2.5, 97.5))

all_di_shapes_high <- all_di_shapes_long %>%
  filter(metric_mod == "excl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_high = mean(value > 97.5),
            n = dplyr::n())

all_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = c(dat, n), names_from = metric_name, values_from = proportion_high) %>% mutate(dir = "HIGH")

all_di_shapes_low <- all_di_shapes_long %>%
  filter(metric_mod == "incl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_low = mean(value <2.5))

all_di_shapes_low %>%
  tidyr::pivot_wider(id_cols = dat, names_from = metric_name, values_from = proportion_low) %>%
  mutate(dir = "LOW")

all_di_shapes_high <- all_di_shapes_long %>%
  mutate(in_fia = ifelse(dat == "fia", "FIA", "Other")) %>%
  filter(metric_mod == "excl") %>%
  group_by(in_fia, metric_name) %>%
  summarize(proportion_high = mean(value >97.5))

all_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = in_fia, names_from = metric_name, values_from = proportion_high) %>%
  mutate(dir = "high")

all_di_shapes_low <- all_di_shapes_long %>%
  mutate(in_fia = ifelse(grepl("fia", dat), "FIA", "Other")) %>%
  filter(metric_mod == "incl") %>%
  group_by(in_fia, metric_name) %>%
  summarize(proportion_low = mean(value <2.5))

all_di_shapes_low %>%
  tidyr::pivot_wider(id_cols = in_fia, names_from = metric_name, values_from = proportion_low) %>%
  mutate(dir = "LOW")

```

### For very small communities

```{r}
all_di <- all_di %>%
  mutate(very_small = nparts < 2000)

```

#### Central tendency

```{r}

all_di %>%
  filter(nparts > 20) %>%
  group_by(dat, very_small) %>%
  summarize(prop_high_po = mean(real_po_percentile_excl > 97.5, na.rm = T),
            nsites = dplyr::n()) %>%
  tidyr::pivot_wider(id_cols = dat, names_from = very_small, values_from = c(prop_high_po, nsites))

ggplot(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund"), nparts > 20), aes(real_po_percentile_mean)) +
  geom_histogram(bins = 40) +
  facet_wrap(vars(dat, very_small), scales = "free_y", ncol = 2)
```

#### Shape metrics

```{r}

all_di_shapes_long <- all_di  %>%
  filter(nparts > 40, dat %in% c("mcdb", "misc_abund", "fia")) %>%
  #mutate(dat = ifelse(nparts < 1000, paste0(dat, "_small"), dat)) %>%
  select(dat, very_small, skew_percentile, skew_percentile_excl, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl) %>%
  rename(skew = skew_percentile,
         skewX = skew_percentile_excl,
         simp = simpson_percentile,
         simpX = simpson_percentile_excl,
         shan = shannon_percentile,
         shanX = shannon_percentile_excl,
         nsng = nsingletons_percentile,
         nsngX = nsingletons_percentile_excl) %>%
  mutate(skewM = (skew + skewX) / 2,
         simpM = (simp+ simpX)/2,
         shanM = (shan+ shanX)/2,
         nsngM = (nsng+ nsngX)/2) %>%
  tidyr::pivot_longer(c(-dat, -very_small), names_to = "metric", values_to = "value") %>%
  mutate(metric_name = substr(metric, 0, 4),
         metric_mod = ifelse(substr(metric, 5, 5) == "X", "excl", ifelse(substr(metric, 5,5) == "M", "mean", "incl")))

all_di_shapes_high <- all_di_shapes_long %>%
  filter(metric_mod == "excl") %>%
  group_by(dat, metric_name, very_small) %>%
  summarize(proportion_high = mean(value > 97.5),
            n = dplyr::n())

all_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = c(dat, n, very_small), names_from = metric_name, values_from = proportion_high) %>% mutate(dir = "HIGH")

all_di_shapes_low <- all_di_shapes_long %>%
  filter(metric_mod == "incl") %>%
  group_by(dat, metric_name, very_small) %>%
  summarize(proportion_low = mean(value <2.5),
            n = dplyr::n())

all_di_shapes_low %>%
  tidyr::pivot_wider(id_cols = c(dat, n, very_small), names_from = metric_name, values_from = proportion_low) %>%
  mutate(dir = "LOW")

```

## Narrowness of the expectation

### Central tendency

```{r}

ggplot(filter(all_di, nparts > 20), aes(nparts, sim_pos_from_best, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10()
# 
# ggplot(filter(all_di, nparts > 20, nparts < 1000), aes(nparts, sim_pos_from_best, color = log(n0/s0))) +
#   geom_point() +
#   scale_x_log10() +
#   theme(legend.position = "top")
# 
# ggplot(filter(all_di, nparts > 10, nparts < 100000), aes(nparts, sim_pos_from_best, color = log(n0/s0))) +
#   geom_point() +
#   scale_x_log10() +
#   theme(legend.position = "top")
# 
# 
# ggplot(filter(all_di, nparts > 10, nparts > 100000), aes(nparts, sim_pos_from_best, color = log(n0/s0))) +
#   geom_point() +
#   scale_x_log10() +
#   theme(legend.position = "top")
```

### Shape metrics

```{r}

ggplot(filter(all_di, nparts > 40, s0 > 2), aes(nparts, skew_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10()

ggplot(filter(all_di, nparts > 40), aes(nparts, simpson_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10()

ggplot(filter(all_di, nparts > 40), aes(nparts, shannon_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10()

ggplot(filter(all_di, nparts > 40), aes(nparts, nsingletons_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10()

```

## Resampling

### Undersampling of rare species

#### Central tendency

```{r}

all_di_singletons <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"), stringsAsFactors = F) %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  left_join(all_ct) %>%
  mutate(in_fia = ifelse(Dataset == "FIA", "FIA", "Other datasets")) %>%
  right_join(select(all_di, dat, site))

all_di_singletons %>%
  group_by(dat) %>%
  summarize(prop_high_po = mean(real_po_percentile_excl > 95, na.rm = T),
            nsites = dplyr::n()) %>%
  left_join(
    all_di %>%
       group_by(dat) %>%
  summarize(prop_high_po_real = mean(real_po_percentile_excl > 95, na.rm = T),
            nsites_real = dplyr::n())
  )

```

#### Shape metrics

```{r}

all_di_shapes_long <- all_di  %>%
  filter(nparts > 40) %>%
  #mutate(dat = ifelse(nparts < 1000, paste0(dat, "_small"), dat)) %>%
  select(dat, skew_percentile, skew_percentile_excl, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl) %>%
  rename(skew = skew_percentile,
         skewX = skew_percentile_excl,
         simp = simpson_percentile,
         simpX = simpson_percentile_excl,
         shan = shannon_percentile,
         shanX = shannon_percentile_excl,
         nsng = nsingletons_percentile,
         nsngX = nsingletons_percentile_excl) %>%
  mutate(skewM = (skew + skewX) / 2,
         simpM = (simp+ simpX)/2,
         shanM = (shan+ shanX)/2,
         nsngM = (nsng+ nsngX)/2) %>%
  tidyr::pivot_longer(-dat, names_to = "metric", values_to = "value") %>%
  mutate(metric_name = substr(metric, 0, 4),
         metric_mod = ifelse(substr(metric, 5, 5) == "X", "excl", ifelse(substr(metric, 5,5) == "M", "mean", "incl")))

all_di_shapes_high <- all_di_shapes_long %>%
  filter(metric_mod == "excl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_high = mean(value > 97.5),
            n = dplyr::n())
all_di_shapes_low <- all_di_shapes_long %>%
  filter(metric_mod == "incl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_low = mean(value <2.5),
            n = dplyr::n())

all_di_shapes_long_s <- all_di_singletons  %>%
  right_join(select(filter(all_di, nparts > 40), dat, site)) %>%
  #mutate(dat = ifelse(nparts < 1000, paste0(dat, "_small"), dat)) %>%
  select(dat, skew_percentile, skew_percentile_excl, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl) %>%
  rename(skew = skew_percentile,
         skewX = skew_percentile_excl,
         simp = simpson_percentile,
         simpX = simpson_percentile_excl,
         shan = shannon_percentile,
         shanX = shannon_percentile_excl,
         nsng = nsingletons_percentile,
         nsngX = nsingletons_percentile_excl) %>%
  mutate(skewM = (skew + skewX) / 2,
         simpM = (simp+ simpX)/2,
         shanM = (shan+ shanX)/2,
         nsngM = (nsng+ nsngX)/2) %>%
  tidyr::pivot_longer(-dat, names_to = "metric", values_to = "value") %>%
  mutate(metric_name = substr(metric, 0, 4),
         metric_mod = ifelse(substr(metric, 5, 5) == "X", "excl", ifelse(substr(metric, 5,5) == "M", "mean", "incl")))

all_di_shapes_high_s <- all_di_shapes_long_s %>%
  filter(metric_mod == "excl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_high = mean(value > 97.5),
            n = dplyr::n())
all_di_shapes_low_s <- all_di_shapes_long_s %>%
  filter(metric_mod == "incl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_low = mean(value <2.5),
            n = dplyr::n())

all_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = c(dat, n), names_from = metric_name, values_from = proportion_high) %>% mutate(dir = "HIGH", adj = F) %>%
  bind_rows(all_di_shapes_high_s %>%
  tidyr::pivot_wider(id_cols = c(dat, n), names_from = metric_name, values_from = proportion_high) %>% mutate(dir = "HIGH", adj = T)) %>%
  arrange(dat, adj)

all_di_shapes_low %>%
  tidyr::pivot_wider(id_cols = c(dat, n), names_from = metric_name, values_from = proportion_low) %>%
  mutate(dir = "LOW", adj = F) %>%
  bind_rows(all_di_shapes_low_s %>%
       tidyr::pivot_wider(id_cols = c(dat, n), names_from = metric_name, values_from = proportion_low) %>% mutate(dir = "LOW", adj = T)) %>%
  arrange(dat, adj)

```

### Jackknife resampling

```{r}

jk_di_full <- read.csv(here::here("analysis", "rev_prototyping", "all_di_jk.csv")) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20)

misc_jk_ct <- read.csv(here::here("misc_jk_cts.csv"))
mcdb_jk_ct <- read.csv(here::here("mcbd_jk_cts.csv"))

jk_cts <- read.csv(here::here("analysis", "rev_prototyping", "all_ct_jk.csv")) %>%
  bind_rows(misc_jk_ct) %>%
  bind_rows(mcdb_jk_ct) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20)

jk_di <- jk_di_full %>%
  left_join(jk_cts) %>%
  group_by_all() %>%
  mutate(site_source = unlist(strsplit(site, "_")[[1]][[1]])) %>%
  ungroup() %>%
  select(dat, site, site_source, s0, n0, skew_percentile_excl,skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile,shannon_percentile_excl,  nsingletons_percentile, nsingletons_percentile_excl, nsingletons, nsingletons_95_ratio_2t, skew_95_ratio_2t, simpson_95_ratio_2t, shannon_95_ratio_2t, sim_pos_from_best, real_po_percentile, real_po_percentile_excl, nparts)

jk_di_means <- jk_di %>%
  group_by(dat, site_source) %>%
  summarise_at(vars(-group_cols(), -site), mean)

all_di_actual <- all_di %>%
  filter(!singletons) %>%
  mutate(site_source = as.character(site)) %>%
  select(dat, site_source,  s0, n0, skew_percentile_excl,skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile,shannon_percentile_excl,  nsingletons_percentile, nsingletons_percentile_excl, nsingletons, nsingletons_95_ratio_2t, skew_95_ratio_2t, simpson_95_ratio_2t, shannon_95_ratio_2t, sim_pos_from_best, real_po_percentile, real_po_percentile_excl, nparts)

colnames(all_di_actual)[ 3:ncol(all_di_actual)] <- paste0(colnames(all_di_actual)[3:ncol(all_di_actual)], "_actual")

jk_di_results <- left_join(jk_di, all_di_actual)
jk_di_mean_results <- left_join(jk_di_means, all_di_actual)

```

#### Central tendency

```{r}
jk_di_mean_results %>%
  filter(nparts > 20) %>%
  select(dat, real_po_percentile_excl, real_po_percentile_excl_actual) %>%
  group_by(dat) %>%
  summarize(high_dissimilarity_actual = mean(real_po_percentile_excl_actual > 95, na.rm = T),
            high_dissimilarity_jk = mean(real_po_percentile_excl > 95, na.rm= T),
            n = dplyr::n())

```

#### Shape metrics

```{r}

shapecols <- c(colnames(jk_di_mean_results)[ grepl("skew", colnames(jk_di_mean_results))],
               colnames(jk_di_mean_results)[ grepl("simpson", colnames(jk_di_mean_results))],
               colnames(jk_di_mean_results)[ grepl("shannon", colnames(jk_di_mean_results))],
               colnames(jk_di_mean_results)[ grepl("nsingletons", colnames(jk_di_mean_results))])

shapecols <- shapecols[ grepl("percentile", shapecols)]

shapecols_high <- shapecols[ grepl("excl", shapecols)]

is.high <- function(x) {
  return(x > 97.5)
}

jk_shapes_high <- jk_di_mean_results %>%
  filter(nparts > 40, nparts_actual > 40) %>%
  select(dat, (shapecols_high)) %>%
  mutate_at(vars(shapecols_high), .funs = is.high) %>%
  group_by(dat) %>%
  summarize_at(vars(shapecols_high), .funs = mean) %>%
  mutate(dir = "HIGH")

shapecols_low <- shapecols[ !grepl("excl", shapecols)]
jk_shapes_high
is.low <- function(x) {
  return(x < 2.5)
}

jk_shapes_low <- jk_di_mean_results %>%
    filter(nparts > 40, nparts_actual > 40) %>%

  select(dat, (shapecols_low)) %>%
  mutate_at(vars(shapecols_low), .funs = is.low) %>%
  group_by(dat) %>%
  summarize_at(vars(shapecols_low), .funs = mean) %>%
  mutate(dir = "LOW")
jk_shapes_low
```

### Narrowness

```{r}

ggplot(jk_di_mean_results, aes(sim_pos_from_best_actual)) +
  geom_histogram(alpha = .5) +
  geom_histogram(aes(sim_pos_from_best), fill="blue", alpha = .5) +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(filter(jk_di_mean_results, s0 > 2, nparts > 40, nparts_actual > 40), aes(skew_95_ratio_2t_actual)) +
  geom_histogram(alpha = .5) +
  geom_histogram(aes(skew_95_ratio_2t), fill="blue", alpha = .5) +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(filter(jk_di_mean_results, nparts > 40, nparts_actual > 40), aes(simpson_95_ratio_2t_actual)) +
  geom_histogram(alpha = .5) +
  geom_histogram(aes(simpson_95_ratio_2t), fill="blue", alpha = .5) +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(filter(jk_di_mean_results, nparts > 40, nparts_actual > 40), aes(shannon_95_ratio_2t_actual)) +
  geom_histogram(alpha = .5) +
  geom_histogram(aes(shannon_95_ratio_2t), fill="blue", alpha = .5) +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(filter(jk_di_mean_results, nparts > 40, nparts_actual > 40), aes(nsingletons_95_ratio_2t_actual)) +
  geom_histogram(alpha = .5) +
  geom_histogram(aes(nsingletons_95_ratio_2t), fill="blue", alpha = .5) +
  facet_wrap(vars(dat), scales = "free_y")

```

## Gentry

```{r}

all_di <- all_di %>%
  mutate(low_avgn = n0/s0 < 3)

ggplot(filter(all_di, nparts > 40, dat == "gentry"), aes( nsingletons_percentile)) + 
geom_histogram() +
  facet_wrap(vars(low_avgn), scales = "free_y")

all_di %>%
  filter(nparts > 40, dat == "gentry") %>%
  group_by(low_avgn) %>%
  summarize(low_nsingles = mean(nsingletons_percentile < 2.5),
            high_even = mean(simpson_percentile_excl > 97.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            n = dplyr::n())

ggplot(filter(all_di, nparts > 40), aes(nsingletons_median / s0)) +
  geom_histogram() +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  xlab("Median proportion of rare species")

ggplot(filter(all_di, nparts > 40), aes(nsingletons_median / s0)) +
  geom_histogram() +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  xlab("Median proportion of rare species")

ggplot(filter(all_di, nparts > 40, dat == "gentry"), aes(nsingletons_median / s0)) +
  geom_histogram() +
  facet_wrap(vars(low_avgn), scales = "free_y", ncol = 1) +
  xlab("Median proportion of rare species")

ggplot(filter(all_di, nparts > 40, dat == "gentry"), aes(s0, n0, color = low_avgn)) +
  geom_point()

ggplot(filter(all_di, nparts > 40), aes(s0, n0, color = nsingletons_median / s0, shape = Dataset)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  theme(legend.position = "top")

ggplot(filter(all_di, nparts > 40, dat == "gentry"), aes((n0/s0), nsingletons_percentile)) + 
  geom_point() +
  scale_x_log10() +
  geom_hline(yintercept = 2.5)

```

## Comparison of FIA to other small communities

```{r}

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia = filter(all_di, dat == "fia")

small_di <- all_di %>%
  filter(s0 <= fia_max_s,
         n0 <= fia_max_n) %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"),
         right_corner = FALSE)

for(i in 1:nrow(small_di)) {
  if(small_di$dat[i] != "fia") {
    fia_match_s0 = filter(fia, s0 >= small_di$s0[i])

    max_n0_for_this_s0 = max(fia_match_s0$n0)

    small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0

  }
}

small_di <- small_di %>%
  filter(!right_corner)
not_fia <- filter(small_di, fia_yn != "fia")
fia = filter(small_di, dat == "fia")

small_di_s_n <- small_di %>%
  select(s0, n0) %>%
  distinct() %>%
  mutate(in_fia = FALSE,
         in_not_fia = FALSE)

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)
  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  small_di_s_n$in_fia[i] = nrow(fia_matches) > 0 
  small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0

}

small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia)

set.seed(1977)

subsamples <- list()

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)

  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches))

  fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F)
  not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F)

  subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ]))

}

sub_di <- bind_rows(subsamples) %>%
  mutate(Dataset = ifelse(fia_yn == "fia", "FIA", "Other datasets"))

fia_di <- filter(sub_di, Dataset == "FIA")
other_di <- filter(sub_di, Dataset != "FIA")

```

### Central tendency

```{r}
filter(sub_di, nparts > 20) %>%
  group_by(Dataset) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95))

ggplot(sub_di, aes(real_po_percentile_excl)) +
  geom_histogram() +
  facet_wrap(vars(Dataset))
ks.test(fia_di$real_po_percentile_excl, other_di$real_po_percentile_excl)

ggplot(sub_di, aes(sim_pos_from_best)) +
  geom_histogram() +
  facet_wrap(vars(Dataset))

ks.test(fia_di$sim_pos_from_best, other_di$sim_pos_from_best)

```

### Shape metrics

```{r, fig.dim = c(7, 4)}

sub_di_shapes_long <- sub_di  %>%
  filter(nparts > 40) %>%
  #mutate(Dataset = ifelse(nparts < 1000, paste0(Dataset, "_small"), Dataset)) %>%
  select(Dataset, skew_percentile, skew_percentile_excl, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl) %>%
  rename(skew = skew_percentile,
         skewX = skew_percentile_excl,
         simp = simpson_percentile,
         simpX = simpson_percentile_excl,
         shan = shannon_percentile,
         shanX = shannon_percentile_excl,
         nsng = nsingletons_percentile,
         nsngX = nsingletons_percentile_excl) %>%
  mutate(skewM = (skew + skewX) / 2,
         simpM = (simp+ simpX)/2,
         shanM = (shan+ shanX)/2,
         nsngM = (nsng+ nsngX)/2) %>%
  tidyr::pivot_longer(-Dataset, names_to = "metric", values_to = "value") %>%
  mutate(metric_name = substr(metric, 0, 4),
         metric_mod = ifelse(substr(metric, 5, 5) == "X", "excl", ifelse(substr(metric, 5,5) == "M", "mean", "incl")))

ggplot(filter(sub_di_shapes_long, metric_mod == "mean"), aes(value)) +
  geom_histogram(bins = 40) +
  #facet_grid(rows = vars(dat), cols = vars(metric_name), scales = "free", switch = "y") +
  facet_wrap(vars(metric_name, Dataset), scales = "free", ncol = 2) +
  geom_vline(xintercept = c(2.5, 97.5))

sub_di_shapes_high <- sub_di_shapes_long %>%
  filter(metric_mod == "excl") %>%
  group_by(Dataset, metric_name) %>%
  summarize(proportion_high = mean(value > 97.5),
            n = dplyr::n())

sub_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = c(Dataset, n), names_from = metric_name, values_from = proportion_high) %>% mutate(dir = "HIGH")

sub_di_shapes_low <- sub_di_shapes_long %>%
  filter(metric_mod == "incl") %>%
  group_by(Dataset, metric_name) %>%
  summarize(proportion_low = mean(value <2.5))

sub_di_shapes_low %>%
  tidyr::pivot_wider(id_cols = Dataset, names_from = metric_name, values_from = proportion_low) %>%
  mutate(dir = "LOW")

ks.test(fia_di$nsingletons_percentile_excl, other_di$nsingletons_percentile_excl)
ks.test(fia_di$skew_percentile_excl, other_di$skew_percentile_excl)
ks.test(fia_di$simpson_percentile, other_di$simpson_percentile)
ks.test(fia_di$shannon_percentile, other_di$shannon_percentile)

```