### 2021-04-13 analysis/reports/submission2/manuscript_main_rev.Rmd

---
title: "Figures and results for main manuscript"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)
theme_set(theme_bw())

all_di <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"), stringsAsFactors = F)

all_ct <-  read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"), stringsAsFactors = F)

#fia_ct <- read.csv(here::here("fia_cts.csv"))

#all_ct <- rbind(all_ct, fia_ct)
all_ct <- all_ct %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  left_join(all_ct)

all_di <- all_di %>%
  group_by_all() %>%
  mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl),
         skew_percentile_mean = mean(skew_percentile, skew_percentile_excl),
         simpson_percentile_mean = mean(simpson_percentile,simpson_percentile_excl),
         shannon_percentile_mean = mean(shannon_percentile, shannon_percentile_excl),
         nsingletons_percentile_mean = mean(nsingletons_percentile, nsingletons_percentile_excl),) %>%
  ungroup()

all_di <- all_di %>%
  mutate(in_fia = ifelse(Dataset == "FIA", "FIA", "Other datasets"))

```

# Datasets by S and N (Figure 1)

```{r, fig.dim = c(7,5)}

s4_fig1 <- ggplot(all_di, aes(x = s0, y = n0, color = Dataset)) +
   geom_point(data = filter(all_di, dat == "fia"), alpha = .1) +
   geom_point(data = filter(all_di, dat != "fia"), alpha = .2) +
     geom_point(data = filter(all_di, dat == "points for legend"), alpha = 1) +

  theme_bw() +
  scale_color_viridis_d(end = .9) +
  ggtitle("Communities by S and N") +
  xlab("Species richness (S); note log scale") +
  ylab("Total abundance (N); note log scale") +
  scale_x_log10() +
  scale_y_log10()

plot(s4_fig1)
```

<!-- # Illustrations of 95% interval (Figure 2) -->

<!-- To show the 95% interval, we need to load the distribution of shape metric values from the samples from the feasible set for a few communities. See rov_metric.md. -->

<!-- ```{r, fig.dim = c(7,7)} -->
<!-- library(drake) -->

<!-- db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache-net.sqlite")) -->
<!-- cache <- storr::storr_dbi("datatable", "keystable", db) -->
<!-- cache$del(key = "lock", namespace = "session") -->

<!-- net_summary <- readd(all_di_summary, cache = cache) -->

<!-- net_summary <- net_summary %>% -->
<!--   mutate(log_nparts = log(gmp:::as.double.bigz(nparts))) -->

<!-- example_fs <- readd(fs_s_44_n_13360, cache = cache) -->

<!-- example_di <- readd(di_fs_s_44_n_13360, cache =cache) -->

<!-- example_fs <- example_fs %>% -->
<!--   left_join(example_di) %>% -->
<!--   left_join(net_summary) -->

<!-- example_fs2 <- readd(fs_s_13_n_315, cache = cache) -->

<!-- example_di2 <- readd(di_fs_s_13_n_315, cache =cache) -->

<!-- example_fs2 <- example_fs2 %>% -->
<!--   left_join(example_di2) %>% -->
<!--   left_join(net_summary) -->

<!-- example_fs3 <- readd(fs_s_4_n_34, cache = cache) -->

<!-- example_di3 <- readd(di_fs_s_4_n_34, cache =cache) -->

<!-- example_fs3 <- example_fs3 %>% -->
<!--   left_join(example_di3) %>% -->
<!--   left_join(net_summary) -->

<!-- breadth_plots <- list( -->

<!--   ggplot(example_fs3, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .25) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Small community", subtitle = paste0("S = ", (example_fs3$s0), "; N = ", (example_fs3$n0[1]))) + -->
<!--     theme(legend.position = "right") + -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance"), -->

<!--   ggplot(example_fs3, aes(skew)) + -->
<!--   #  geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs3$skew_97p5[1], example_fs3$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle = paste0("Breadth index: ", round((example_fs3$skew_95_ratio_2t[1]), 2))) + -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count"), -->

<!--   ggplot(example_fs2, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .1) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Medium community", subtitle = paste0("S = ", (example_fs2$s0), "; N = ", (example_fs2$n0[1]))) + -->
<!--     theme(legend.position = "right")+ -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance") + -->
<!--     ylim(0, 200), # Remove 3 sads that make the axes too big to be interpretable -->

<!--   ggplot(example_fs2, aes(skew)) + -->
<!--   #  geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs2$skew_97p5[1], example_fs2$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle =  paste0("Breadth index: ", round((example_fs2$skew_95_ratio_1t[1]), 2)))+ -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count"), -->

<!--   ggplot(example_fs, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .1) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Large community", subtitle = paste0("S = ", (example_fs$s0), "; N = ", (example_fs$n0[1]))) + -->
<!--     theme(legend.position = "right") + -->
<!--     ylim(0, 4000) + # Remove a very few very very uneven SADs that make the scale too big to be interpretable -->
<!--     theme(axis.text.y = element_text(size = 6, angle = 60))+ -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance"), -->

<!--   ggplot(example_fs, aes(skew)) + -->
<!--    # geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs$skew_97p5[1], example_fs$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle =  paste0("Breadth index: ", round((example_fs$skew_95_ratio_1t[1]), 2)))+ -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count") -->

<!-- ) -->

<!-- fig_1 <- gridExtra::grid.arrange(grobs = breadth_plots, ncol = 2, top = textGrob("Figure 2", gp = gpar(fill = "white"))) -->

<!-- plot(fig_1) -->

<!-- signif(as.numeric(example_di$nparts), digits = 2) -->
<!-- (as.numeric(example_di2$nparts)) -->
<!-- as.numeric(example_di3$nparts) -->

<!-- DBI::dbDisconnect(db) -->
<!-- rm(cache) -->
<!-- rm(db) -->
<!-- ``` -->

# Dissimilarity (Supplement)

```{r, fig.dim = c(3, 6)}
plot_percentile_hist <- function(di_df, col_name, plot_name, tails = 2, facetvar = "Dataset", min_s0 = 0) {

  if(tails == 2) {
    cutoff_percentiles = c(2.5, 97.5)
    min_nparts = 40
  } else if (tails== 1) {
    cutoff_percentiles = c(95)
    min_nparts = 20
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[col_name]]),
           facetvar = di_df[[facetvar]])

  ggplot(filter(di_df, nparts > min_nparts, s0 >= min_s0), aes(response)) +
    geom_histogram(bins = 40, boundary = 100) +
    theme_bw() +
    xlab("") +
    ylab("") +
    geom_vline(xintercept = cutoff_percentiles, color = "red") +
    ggtitle( plot_name) +
    facet_wrap(vars(facetvar), ncol = 1, scales = "free_y")+
    theme(plot.title = element_text(size=10))

}

all_di <- all_di %>%
  mutate(`Observed \npercentile score` = ifelse(real_po_percentile_excl > 95, "High (> 95)", "Less than 95"))

```

<!-- ```{r, fig.dim = c(4, 4)} -->
<!-- ggplot(all_di, aes(sim_pos_from_best, real_po, color = `Observed \npercentile score`)) + -->
<!--   geom_point(alpha = .3) + -->
<!--   geom_point(data = filter(all_di, s0 < 0)) + -->
<!--   geom_abline(slope = 1, intercept = 0) + -->
<!--   scale_color_viridis_d(end = .8, direction = -1) + -->
<!--   xlab("Mean of elements from feasible set") + -->
<!--   ylab("Observed value") + -->
<!--   ggtitle("Dissimilarity to the central tendency") +  theme(legend.position = "bottom")#, -->
<!-- # -->
<!-- # -->
<!-- # ggplot(filter(all_di, nparts < 10 ^ 50), aes(nparts, sim_pos_from_best, color = Dataset)) + -->
<!-- #   geom_point(alpha = .3) + -->
<!-- #   geom_point(data = filter(all_di, s0 < 0)) + -->
<!-- #   scale_color_viridis_d(end = .9) + -->
<!-- #   xlab("Number of elements in the feasible set") + -->
<!-- #   ylab("Mean dissimilarity of feasible set to central tendency") + -->
<!-- #   scale_x_log10() + -->
<!-- #   ggtitle("Over the size of the feasible set") -->
<!-- # ), ncol = 1, -->
<!-- #top = textGrob("Dissimilarity to central tendency")) -->

<!-- ``` -->

# Metric histograms by dataset (Figure 4)

```{r, fig.dim = c(10,5)}

fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_percentile_hist(all_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(all_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(all_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(all_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(all_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Figure 3", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))
```

# Results table

```{r}
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

cols2 <- c("Dataset", "Low proportion of rare species", "Low skew", "High Simpson", "High Shannon")

all_results %>%
  select(cols2)

```

# Narrowness

```{r, fig.dim = c(3, 6)}

plot_narrowness <- function(di_df, yvar, yvar_name) {

  if(grepl("sim_pos", yvar)) {
    min_nparts = 20
    ylabel = "Mean dissimilarity"
  } else {
    min_nparts = 40
    ylabel = "Breadth index"
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[yvar]])) %>%
    filter(nparts > min_nparts,
           nparts < 10 ^ 50)

  ggplot(di_df, aes(nparts, response, color = Dataset)) +
    geom_point(data = filter(di_df, dat == "fia"), alpha = .1) +
    geom_point(data = filter(di_df, dat != "fia"), alpha = .1) +
    geom_point(data = filter(di_df, s0 < 0)) +
    scale_color_viridis_d(end = .9) +
    xlab("") +
    ylab(ylabel) +
    scale_x_log10() +
    ggtitle(yvar_name) +
    theme(legend.position = "none")+
    theme(plot.title = element_text(size=10))
}

plot_narrowness_legend <- function(di_df) {

  legend_df <- di_df %>%
    select(Dataset) %>%
    distinct() %>%
    mutate(ymark = dplyr::row_number())

  ggplot(legend_df, aes(1, ymark, color = Dataset)) +
    geom_label(aes(x = 2, y = ymark, label = Dataset)) +
    geom_point(size = 4) +
    scale_color_viridis_d(end = .9) +
    xlab("") +
    ylab("") +
    theme(legend.position = "none") +
    xlim(1, 3) + theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    ) +
    ggtitle("Dataset")

}
#
# gridExtra::grid.arrange(grobs = list(
#   plot_narrowness(all_di, "sim_pos_from_best", "Dissimilarity to central tendency"),
#   plot_narrowness_legend(all_di)),
#   ncol = 1)
```

```{r, fig.dim = c(7,4)}

gridExtra::grid.arrange(grobs = list(
  plot_narrowness(all_di, "sim_pos_from_best", "Dissimilarity \nto central tendency") + ylab("Mean dissimilarity") + theme(legend.position = "bottom", legend.direction = "vertical" , legend.margin = margin(b = .25, r = 0, t = 0, l = 0, unit = "in")),
  gridExtra::grid.arrange(grobs = list(
    plot_narrowness(all_di, "skew_95_ratio_2t", "\nSkewness"),
    plot_narrowness(all_di, "simpson_95_ratio_2t", "\nSimpson evenness")+ ylab(""),
    plot_narrowness(all_di, "nsingletons_95_ratio_2t", "\nNumber of rare species"),
       plot_narrowness(all_di, "shannon_95_ratio_2t", "\nShannon diversity") + ylab("")
  ), ncol = 2)), ncol = 2, widths = c( 2, 4), bottom = textGrob("Number of elements in the feasible set"))

```

```{r}

plot_narrowness_hist <- function(di_df, col_name, plot_name, facetvar = "Dataset", min_s0 = 0) {

  if(grepl("sim_pos", col_name)) {
    xvarname = "Mean dissimilarity"
    min_nparts = 40
  } else {
    min_nparts = 20
    xvarname = "Breadth index"
  }

  di_df <- di_df %>%
    mutate(response = (di_df[[col_name]]),
           facetvar = di_df[[facetvar]])

  ggplot(filter(di_df, nparts > min_nparts, s0 >= min_s0), aes(response)) +
    geom_histogram(bins = 40, boundary = 100) +
    theme_bw() +
    ylab("") +
    xlab(xvarname) +
    ggtitle( plot_name) +
    facet_wrap(vars(facetvar), ncol = 1, scales = "free_y")+
    theme(plot.title = element_text(size=10))

}

```

<!-- ```{r, fig.dim = c(10,5)} -->

<!-- fig_2 <- gridExtra::grid.arrange(grobs = list( -->

<!--   plot_narrowness_hist(all_di, "sim_pos_from_best", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(all_di, "nsingletons_95_ratio_2t", "\nNumber of rare species", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(all_di, "skew_95_ratio_2t", "\nSkewness", facetvar = "Dataset", min_s0 = 3), -->
<!--   plot_narrowness_hist(all_di, "simpson_95_ratio_2t", "\nSimpson evenness", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(all_di, "shannon_95_ratio_2t", "\nShannon diversity", facetvar = "Dataset") -->
<!-- ), ncol = 5, -->
<!--  left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), gp = gpar(fill = "white")) -->

<!-- ``` -->

<!-- # Gentry -->

<!-- ```{r} -->

<!-- gentry_di <- filter(all_di, Dataset == "Gentry") -->

<!-- gentry_di <- gentry_di %>% -->
<!--   mutate(low_avg_abund = ifelse(n0/s0 < 3, "N:S < 3", "N:S > 3")) -->

<!-- fig_2 <- gridExtra::grid.arrange(grobs = list( -->

<!--   plot_percentile_hist(gentry_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "low_avg_abund"), -->
<!--   plot_percentile_hist(gentry_di, "skew_percentile_mean", "\nSkewness", facetvar = "low_avg_abund", min_s0 = 3), -->
<!--   plot_percentile_hist(gentry_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "low_avg_abund"), -->
<!--   plot_percentile_hist(gentry_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "low_avg_abund") -->
<!-- ), ncol = 2, -->
<!--  left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), -->
<!-- bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white")) -->

<!-- ``` -->

<!-- # Comparison of FIA and comparably sized communities (Figure 4) -->

<!-- ```{r, fig.dim = c(10, 5)} -->

<!-- fia_max_s = max(filter(all_di, dat == "fia")$s0) -->
<!-- fia_max_n = max(filter(all_di, dat == "fia")$n0) -->

<!-- fia = filter(all_di, dat == "fia") -->

<!-- small_di <- all_di %>% -->
<!--   filter(s0 <= fia_max_s, -->
<!--          n0 <= fia_max_n) %>% -->
<!--   mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"), -->
<!--          right_corner = FALSE) -->

<!-- for(i in 1:nrow(small_di)) { -->
<!--   if(small_di$dat[i] != "fia") { -->
<!--     fia_match_s0 = filter(fia, s0 >= small_di$s0[i]) -->

<!--     max_n0_for_this_s0 = max(fia_match_s0$n0) -->

<!--     small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0 -->

<!--   } -->
<!-- } -->

<!-- small_di <- small_di %>% -->
<!--   filter(!right_corner) -->
<!-- not_fia <- filter(small_di, fia_yn != "fia") -->
<!-- fia = filter(small_di, dat == "fia") -->

<!-- small_di_s_n <- small_di %>% -->
<!--   select(s0, n0) %>% -->
<!--   distinct() %>% -->
<!--   mutate(in_fia = FALSE, -->
<!--          in_not_fia = FALSE) -->

<!-- for(i in 1:nrow(small_di_s_n)) { -->

<!--   this_s0 = small_di_s_n$s0[i] -->
<!--   this_n0 = small_di_s_n$n0[i] -->

<!--   fia_matches = filter(fia, s0 == this_s0, n0 == this_n0) -->
<!--   not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0) -->

<!--   small_di_s_n$in_fia[i] = nrow(fia_matches) > 0 -->
<!--   small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0 -->

<!-- } -->

<!-- small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia) -->

<!-- set.seed(1977) -->

<!-- subsamples <- list() -->

<!-- for(i in 1:nrow(small_di_s_n)) { -->

<!--   this_s0 = small_di_s_n$s0[i] -->
<!--   this_n0 = small_di_s_n$n0[i] -->

<!--   fia_matches = filter(fia, s0 == this_s0, n0 == this_n0) -->

<!--   not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0) -->

<!--   n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches)) -->

<!--   fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F) -->
<!--   not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F) -->

<!--   subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ])) -->

<!-- } -->

<!-- sub_di <- bind_rows(subsamples) %>% -->
<!--   mutate(Dataset = ifelse(fia_yn == "fia", "FIA", "Other datasets")) -->

<!-- fia_di <- filter(sub_di, fia_yn == "fia") -->
<!-- other_di <- filter(sub_di, fia_yn != "fia") -->

<!-- ks_compare <- function(fia_df, other_df, compare_var) { -->

<!--   if(grepl("skew", compare_var)) { -->
<!--     fia_df <- filter(fia_df, s0 > 2) -->
<!--     other_df <- filter(fia_df, s0 >2) -->
<!--   } -->

<!--   if(grepl("sim_pos", compare_var)) { -->
<!--     fia_df <- filter(fia_df, nparts > 20) -->
<!--     other_df <- filter(other_df, nparts > 20) -->
<!--   } else { -->
<!--       fia_df <- filter(fia_df, nparts > 40) -->
<!--     other_df <- filter(other_df, nparts > 40) -->
<!--   } -->

<!--   kstest <- ks.test(fia_df[[compare_var]], other_df[[compare_var]]) -->

<!--   return(data.frame( -->
<!--     var = compare_var, -->
<!--     d = kstest$statistic, -->
<!--     p = kstest$p.value -->
<!--   )) -->
<!-- } -->

<!-- simpson_rat_ks <- ks_compare(fia_di, other_di, "simpson_95_ratio_2t") -->
<!-- skew_rat_ks <- ks_compare(fia_di, other_di, "skew_95_ratio_2t") -->
<!-- shannon_rat_ks <- ks_compare(fia_di, other_di, "shannon_95_ratio_2t") -->
<!-- nsingletons_rat_ks <- ks_compare(fia_di, other_di, "nsingletons_95_ratio_2t") -->
<!-- diss_rat_ks <- ks_compare(fia_di, other_di, "sim_pos_from_best") -->

<!-- bind_rows(simpson_rat_ks, skew_rat_ks, shannon_rat_ks, nsingletons_rat_ks, diss_rat_ks) -->

<!-- simpson_perc_ks <- ks_compare(fia_di, other_di, "simpson_percentile") -->
<!-- skew_perc_ks <- ks_compare(fia_di, other_di, "skew_percentile_excl") -->
<!-- shannon_perc_ks <- ks_compare(fia_di, other_di, "shannon_percentile") -->
<!-- nsingletons_perc_ks <- ks_compare(fia_di, other_di, "nsingletons_percentile_excl") -->
<!-- diss_perc_ks <- ks_compare(fia_di, other_di, "real_po_percentile_excl") -->

<!-- bind_rows(simpson_perc_ks, skew_perc_ks, shannon_perc_ks, nsingletons_perc_ks, diss_perc_ks) -->

<!-- fig_2 <- gridExtra::grid.arrange(grobs = list( -->

<!--   plot_percentile_hist(sub_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1), -->
<!--   plot_percentile_hist(sub_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"), -->
<!--   plot_percentile_hist(sub_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3), -->
<!--   plot_percentile_hist(sub_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"), -->
<!--   plot_percentile_hist(sub_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset") -->
<!-- ), ncol = 5, -->
<!-- top = textGrob("Figure 2", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), -->
<!-- bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white")) -->

<!-- fig_2 <- gridExtra::grid.arrange(grobs = list( -->

<!--   plot_narrowness_hist(sub_di, "sim_pos_from_best", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(sub_di, "nsingletons_95_ratio_2t", "\nNumber of rare species", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(sub_di, "skew_95_ratio_2t", "\nSkewness", facetvar = "Dataset", min_s0 = 3), -->
<!--   plot_narrowness_hist(sub_di, "simpson_95_ratio_2t", "\nSimpson evenness", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(sub_di, "shannon_95_ratio_2t", "\nShannon diversity", facetvar = "Dataset") -->
<!-- ), ncol = 5, -->
<!--  left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), gp = gpar(fill = "white")) -->

<!-- diss_results <- sub_di %>% -->
<!--   filter(nparts > 20) %>% -->
<!--   group_by(Dataset) %>% -->
<!--   summarize(high_diss = mean(real_po_percentile_excl > 95), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- skew_results <- sub_di %>% -->
<!--   filter(nparts > 40, s0 > 2) %>% -->
<!--   group_by(Dataset) %>% -->
<!--   summarize(high_skew = mean(skew_percentile_excl > 97.5), -->
<!--             low_skew = mean(skew_percentile < 2.5), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites), -->
<!--          `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- metric_results <- sub_di %>% -->
<!--   filter(nparts > 40) %>% -->
<!--   group_by(Dataset) %>% -->
<!--   summarize(high_simpson = mean(simpson_percentile_excl > 97.5), -->
<!--             low_simpson = mean(simpson_percentile < 2.5), -->
<!--             high_shannon = mean(shannon_percentile_excl > 97.5), -->
<!--             low_shannon = mean(shannon_percentile < 2.5), -->
<!--             high_nsingletons = mean(nsingletons_percentile_excl > 97.5), -->
<!--             low_nsingletons = mean(nsingletons_percentile < 2.5), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites), -->
<!--          `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites), -->
<!--          `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites), -->
<!--          `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites), -->
<!--          `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites), -->
<!--          `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- all_results <- left_join(diss_results, skew_results) %>% -->
<!--   left_join(metric_results) -->

<!-- cols1 <- c("Dataset", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon") -->

<!-- all_results %>% -->
<!--   select(cols1) -->

<!-- ``` -->

# Very small communities

```{r, fig.dim = c(6,8)}

all_di <- all_di %>%
  mutate( `Number of elements` = ifelse(nparts < 2000, "Less than 2000", "More than 2000"))
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "Number of elements", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

all_di %>%
  filter(dat %in% c("fia", "mcdb", "misc_abund")) %>%
  group_by(Dataset) %>%
  summarize(`Proportion with nparts < 2000` = mean(nparts < 2000))

nhist <- plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "sim_pos_from_best", "Dissimilarity to the central tendency") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3) +
  xlab("Mean dissimilarity of elements of feasible set to central tendency") +
  ylab("Number of communities") +
  ggtitle("")

phist <- plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "real_po_percentile_mean", "Dissimilarity to the central tendency") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3) +
  xlab("Percentile rank of observed value relative to feasible set") +
  ylab("Number of communities") +
  ggtitle("")

gridExtra::grid.arrange(grobs = list(nhist, phist), nrow = 2, top = textGrob("Dissimilarity to the central tendency"))

# 
# plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "skew_95_ratio_2t", "Skewness") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "skew_percentile_mean", "Skewness") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

# plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_95_ratio_2t", "Number of rare species") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_percentile_mean", "Number of rare species") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_95_ratio_2t", "Simpson") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_percentile_mean", "Simpson") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# 
# plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_95_ratio_2t", "Shannon") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_percentile_mean", "Shannon") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
```

<!-- ```{r, fig.dim = c(8,3)} -->
<!-- ggplot(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), aes(s0, n0, color = `Number of elements`)) + -->
<!--   geom_point(alpha = .1) + -->
<!--   xlim(0, 30) + -->
<!--   ylim(0,150) + -->
<!--     facet_wrap(vars(Dataset), nrow = 1) -->
<!-- ggplot(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund"), nparts > 2000, nparts < 1e+20), aes(nparts)) + -->
<!--   geom_histogram() + -->
<!--   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y") + -->
<!--   scale_x_log10() -->

<!-- ggplot(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund"), nparts < 2000), aes(nparts)) + -->
<!--   geom_histogram() + -->
<!--   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y") + -->
<!--   scale_x_log10() -->

<!-- ``` -->

<!-- ```{r} -->

<!-- all_di <- all_di %>% -->
<!--   mutate( `Number of elements` = ifelse(nparts < 1000, "Less than 1000", "More than 1000")) -->
<!-- diss_results <- all_di %>% -->
<!--   filter(nparts > 20) %>% -->
<!--   group_by(Dataset,  `Number of elements`) %>% -->
<!--   summarize(high_diss = mean(real_po_percentile_excl > 95), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->
<!-- ggplot(filter(all_di, dat == "mcdb"), aes(skew_95_ratio_2t)) + -->
<!--   geom_histogram() + -->
<!--   facet_wrap(vars(`Number of elements`), scales = "free_y") -->

<!-- skew_results <- all_di %>% -->
<!--   filter(nparts > 40, s0 > 2) %>% -->
<!--   group_by(Dataset,  `Number of elements`) %>% -->
<!--   summarize(high_skew = mean(skew_percentile_excl > 97.5), -->
<!--             low_skew = mean(skew_percentile < 2.5), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites), -->
<!--          `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- metric_results <- all_di %>% -->
<!--   filter(nparts > 40) %>% -->
<!--   group_by(Dataset, `Number of elements`) %>% -->
<!--   summarize(high_simpson = mean(simpson_percentile_excl > 97.5), -->
<!--             low_simpson = mean(simpson_percentile < 2.5), -->
<!--             high_shannon = mean(shannon_percentile_excl > 97.5), -->
<!--             low_shannon = mean(shannon_percentile < 2.5), -->
<!--             high_nsingletons = mean(nsingletons_percentile_excl > 97.5), -->
<!--             low_nsingletons = mean(nsingletons_percentile < 2.5), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites), -->
<!--          `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites), -->
<!--          `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites), -->
<!--          `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites), -->
<!--          `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites), -->
<!--          `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- all_results <- left_join(diss_results, skew_results) %>% -->
<!--   left_join(metric_results) -->

<!-- cols1 <- c("Dataset", "Number of elements", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon") -->

<!-- all_results %>% -->
<!--   select(cols1) -->

<!-- ``` -->

# Resampling

```{r}

all_di_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"))

jk_di_full <- read.csv(here::here("analysis", "rev_prototyping", "all_di_jk.csv"))

all_ct_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"))

jk_ct <- read.csv(here::here("analysis", "rev_prototyping", "all_ct_jk.csv"))

jk_di_full <- left_join(jk_di_full, jk_ct)

all_di_with_resamples <- left_join(all_di_with_resamples, all_ct_with_resamples)

jk_di <- jk_di_full %>%
  group_by_all() %>%
  mutate(site_source = unlist(strsplit(site, "_")[[1]][[1]])) %>%
  ungroup() %>%
  select(dat, site, site_source, s0, n0, skew_percentile_excl,skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl, real_po_percentile, real_po_percentile_excl, sim_pos_from_best, skew_95_ratio_2t, simpson_95_ratio_2t, shannon_95_ratio_2t, nparts)

skewcols <- colnames(jk_di)[ grepl("skew", colnames(jk_di))]

jk_di[ which(jk_di$s0 < 3), skewcols] <- NA

jk_di_means <- jk_di %>%
  filter(n0 != s0,
         s0 != 1,
         n0 != (s0 + 1)) %>%
  group_by(dat, site_source) %>%
  mutate(njks = dplyr::n(),
         njks_skew = sum(!is.na(skew_percentile)))%>%
  summarise_at(vars(-group_cols(), -site), mean, na.rm =  T) %>%
  mutate(resampling = "Subsampling") %>%
  mutate(site = as.numeric(site_source)) %>%
#  select(-site_source) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20,
         njks == 10) %>%
  ungroup()

jk_di[ which(jk_di$njks_skew < 10), skewcols] <- NA

skewcols <- colnames(all_di_with_resamples[ which(grepl("skew", colnames(all_di_with_resamples)))])

all_di_with_resamples[ which(all_di_with_resamples$s0 < 3), skewcols] <- NA

all_di_with_resamples <- all_di_with_resamples %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
       #  !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  mutate(resampling = ifelse(singletons, "Rare species", "Raw")) %>%
  bind_rows(jk_di_means)

metric_rs_results <- all_di_with_resamples %>%
  group_by(Dataset, resampling) %>%
  summarize(
    skew_high = mean(skew_percentile_excl > 97.5, na.rm = T),
    skew_low = mean(skew_percentile < 2.5, na.rm = T),
    even_high = mean(simpson_percentile_excl > 97.5, na.rm = T),
    even_low = mean(simpson_percentile < 2.5, na.rm = T),
    shannon_high = mean(shannon_percentile_excl > 97.5, na.rm = T),
    shannon_low = mean(shannon_percentile < 2.5, na.rm = T),
    nsingletons_high = mean(nsingletons_percentile_excl > 97.5, na.rm = T),
    nsingletons_low = mean(nsingletons_percentile < 2.5, na.rm = T),
    nskew = sum(!is.na(skew_percentile)),
    nother = dplyr::n()
  ) %>%
  ungroup()

diss_rs_results <- all_di_with_resamples %>%
  filter(nparts > 40) %>%
  group_by(Dataset, resampling) %>%
  summarize(dissimilarity_high = mean(real_po_percentile_excl > 95, na.rm = T),
            ndiss = sum(!is.na(real_po_percentile_excl))) %>%
  ungroup()

rs_results <- left_join(metric_rs_results, diss_rs_results)

#metric_cols <- colnames(rs_results [ grepl("_", colnames(rs_results))])

as.perc <- function(x) {
  return((signif(100 * x, digits = 2)))
}

rs_results_desc <- rs_results %>%
  mutate(`High dissimilarity` = paste0(as.perc(dissimilarity_high), "%; n =", ndiss),
         `High proportion of rare species` = paste0(as.perc(nsingletons_high), "%; n =", nother),
         `High skew` = paste0(as.perc(skew_high), "%; n = ", nskew),
         `Low Simpson` = paste0(as.perc(even_low), "%; n = ", nother),
         `Low Shannon` = paste0(as.perc(shannon_low), "%; n = ", nother),
         `Low proportion of rare species` = paste0(as.perc(nsingletons_low), "%; n =", nother),
         `Low skew` = paste0(as.perc(skew_low), "%; n = ", nskew),
         `High Simpson` = paste0(as.perc(even_high), "%; n = ", nother),
         `High Shannon` = paste0(as.perc(shannon_high), "%; n = ", nother),
         `Resampling scheme` = ordered(resampling, levels = c("Raw", "Subsampling", "Rare species"))) %>%
  arrange(Dataset, `Resampling scheme`)

cols1 <- c("Dataset", "Resampling scheme", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

cols2 <- c("Dataset","Resampling scheme",  "Low proportion of rare species", "Low skew", "High Simpson", "High Shannon")

rs_results_desc %>%
  select(cols1)

rs_results_desc %>%
  select(cols2)
```
```{r, fig.dim = c(6,4)}
rs_results_long <- rs_results %>%
  tidyr::pivot_longer((-c("Dataset","resampling")), names_to = "metric", values_to = "value") %>%
  filter(grepl("_", metric)) %>%
  mutate(ometric = ordered(metric, levels = c("skew_high", "skew_low", "even_low", "even_high", "shannon_low", "shannon_high", "nsingletons_high", "nsingletons_low", "dissimilarity_high")),
         ometric2 = ordered(metric, levels = c("dissimilarity_high",  "nsingletons_high", "skew_high",  "even_low",  "shannon_low",  "nsingletons_low","skew_low","even_high","shannon_high"))) %>%
  mutate(weird_dir = as.numeric(ometric2) > 5,
         percentile_cutoff = ifelse(grepl("diss", metric), .05, .025),`Resampling scheme` = ordered(resampling, levels = c("Raw", "Subsampling", "Rare species")))

metric_names <- rs_results_long %>%
  select(ometric2) %>%
  distinct() %>%
  arrange(ometric2) %>%
  mutate(metric_desc =(c("High dissimilarity", "High proportion \nof rare species", "High skew", "Low Simpson", "Low Shannon", "Low proportion \nof rare species", "Low skew", "High Simpson", "High Shannon"))) %>%
  mutate(metric_desc = ordered(metric_desc))

rs_results_long <- left_join(rs_results_long, metric_names)
#
# ggplot(filter(rs_results_long, !weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
#   geom_point() +
#   facet_wrap(vars(Dataset), scales = "fixed", ncol = 1) +
#   geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
#   scale_color_viridis_d(end = .8) +
#   ylab("Proportion of extreme observed values") +
#   xlab("")

gridExtra::grid.arrange(grobs = list(
  ggplot(filter(rs_results_long, !grepl("Misc", Dataset), !weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol =2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("Proportion of extreme observed values") +ylim(0, .7) +
  xlab("") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) ,
  ggplot(filter(rs_results_long, grepl("Misc", Dataset), !weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol = 2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("") +
  xlab("") + ylim(0, .7) +
  theme(legend.position = "bottom", legend.direction = "vertical" , legend.margin = margin(b = .3, r = 0, t = 0, l = 0, unit = "in"), axis.text.x = element_text(angle = 90))), ncol = 2, widths = c( 4,2), bottom = textGrob("Metric"))

gridExtra::grid.arrange(grobs = list(
  ggplot(filter(rs_results_long, !grepl("Misc", Dataset), weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol =2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("Proportion of extreme observed values") +ylim(0, .7) +
  xlab("") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) ,
  ggplot(filter(rs_results_long, grepl("Misc", Dataset), weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol = 2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("") +
  xlab("") + ylim(0, .7) +
  theme(legend.position = "bottom", legend.direction = "vertical" , legend.margin = margin(b = .3, r = 0, t = 0, l = 0, unit = "in"), axis.text.x = element_text(angle = 90))), ncol = 2, widths = c( 4,2), bottom = textGrob("Metric"))

```

<!-- ```{r, fig.dim = c(10,5)} -->

<!-- all_di_with_resamples <- all_di_with_resamples %>% -->
<!--   group_by_all() %>% -->
<!--   mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl, na.rm = T), -->
<!--          skew_percentile_mean = mean(skew_percentile, skew_percentile_excl, na.rm = T), -->
<!--          simpson_percentile_mean = mean(simpson_percentile,simpson_percentile_excl, na.rm = T), -->
<!--          shannon_percentile_mean = mean(shannon_percentile, shannon_percentile_excl, na.rm = T), -->
<!--          nsingletons_percentile_mean = mean(nsingletons_percentile, nsingletons_percentile_excl)) %>% -->
<!--   ungroup() -->

<!-- fig_2 <- gridExtra::grid.arrange(grobs = list( -->

<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1), -->
<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"), -->
<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3), -->
<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"), -->
<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset") -->
<!-- ), ncol = 5, -->
<!-- top = textGrob("Subsampling results", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), -->
<!-- bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white")) -->

<!-- fig_2 <- gridExtra::grid.arrange(grobs = list( -->

<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1), -->
<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"), -->
<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3), -->
<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"), -->
<!--   plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset") -->
<!-- ), ncol = 5, -->
<!-- top = textGrob("Adjusted for rare species results", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), -->
<!-- bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white")) -->
<!-- ``` -->