### 2021-04-19 analysis/reports/submission1/accessory/compare_zenodo_data.Rmd

---
title: "Check that files from Zenodo match files from GitHub"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library(dplyr)
library(ggplot2)

```

Download from Zenodo (https://zenodo.org/record/166725)

```{r}

zip_download_dest  <- "~/Documents/sad-comparison.zip"

download.file("https://zenodo.org/record/166725/files/weecology/sad-comparison-peerj.2.zip?download=1",zip_download_dest , quiet = TRUE, mode = "wb")
unzip("~/Documents/sad-comparison.zip", exdir = "~/Documents/sad-comparison")
file.remove(zip_download_dest)

```

Load data files from Zenodo downloads

```{r}

zenodo_dir <- list.dirs("~/Documents/sad-comparison/", full.names = T)

zenodo_dir <- zenodo_dir[ which(grepl("sad-data", zenodo_dir))]
```

BBS:

```{r}
bbs_zenodo <- read.csv(file.path(zenodo_dir, "bbs_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

bbs_github <- read.csv(here::here("working-data", "abund_data", "bbs_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

any(!(bbs_github == bbs_zenodo))

```

FIA:

```{r}
fia_zenodo <- read.csv(file.path(zenodo_dir, "fia_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

fia_github <- read.csv(here::here("working-data", "abund_data", "fia_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

any(!(fia_github == fia_zenodo))

```

Gentry:

```{r}
gentry_zenodo <- read.csv(file.path(zenodo_dir, "gentry_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

gentry_github <- read.csv(here::here("working-data", "abund_data", "gentry_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

any(!(gentry_github == gentry_zenodo), na.rm = T)

any(!(which(is.na(gentry_github)) == which(is.na(gentry_zenodo))))

```
MCDB:

```{r}
mcdb_zenodo <- read.csv(file.path(zenodo_dir, "mcdb_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

mcdb_github <- read.csv(here::here("working-data", "abund_data", "mcdb_spab.csv"), stringsAsFactors = F, header = F, skip = 2)

any(!(mcdb_github == mcdb_zenodo))

```

All the data files downloaded from GitHub exactly match their counterparts on Zenodo. (This is not surprising but good to check).

The analyses for this manuscript were done using the files downloaded from GitHub. RMD doesn't think it's necessary (or really appropriate) to re-write `download_data` to download from Zenodo instead. This report demonstrates that the analyses should be replicable using the data files from Zenodo, should that be the only/the preferred option in the future.